\documentclass[reqno,11pt]{amsart}
\usepackage[letterpaper, margin=1in]{geometry}
\RequirePackage{amsmath,amssymb,amsthm,graphicx,mathrsfs,url,slashed,subcaption}
\RequirePackage[usenames,dvipsnames]{xcolor}
\RequirePackage[colorlinks=true,linkcolor=Red,citecolor=Green]{hyperref}
\RequirePackage{amsxtra}
\usepackage{cancel}
\usepackage{tikz-cd}

% \setlength{\textheight}{9.3in} \setlength{\oddsidemargin}{-0.25in}
% \setlength{\evensidemargin}{-0.25in} \setlength{\textwidth}{7in}
% \setlength{\topmargin}{-0.25in} \setlength{\headheight}{0.18in}
% \setlength{\marginparwidth}{1.0in}
% \setlength{\abovedisplayskip}{0.2in}
% \setlength{\belowdisplayskip}{0.2in}
% \setlength{\parskip}{0.05in}
%\renewcommand{\baselinestretch}{1.05}

\title{Comass minimizers}
\author{Aidan Backus}
\date{\today}

\newcommand{\NN}{\mathbf{N}}
\newcommand{\ZZ}{\mathbf{Z}}
\newcommand{\QQ}{\mathbf{Q}}
\newcommand{\RR}{\mathbf{R}}
\newcommand{\CC}{\mathbf{C}}
\newcommand{\DD}{\mathbf{D}}
\newcommand{\PP}{\mathbf P}
\newcommand{\MM}{\mathbf M}
\newcommand{\II}{\mathbf I}
\newcommand{\Hyp}{\mathbf H}
\newcommand{\Sph}{\mathbf S}
\newcommand{\Group}{\mathbf G}
\newcommand{\GL}{\mathbf{GL}}
\newcommand{\Orth}{\mathbf{O}}
\newcommand{\SpOrth}{\mathbf{SO}}
\newcommand{\Ball}{\mathbf{B}}

\newcommand*\dif{\mathop{}\!\mathrm{d}}

\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\tr}{tr}

\newcommand{\Leaves}{\mathscr L}
\newcommand{\Lagrange}{\mathcal L}
\newcommand{\Hypspace}{\mathscr H}

\newcommand{\Chain}{\underline C}

\newcommand{\Two}{\mathrm{I\!I}}

\newcommand{\normal}{\mathbf n}
\newcommand{\radial}{\mathbf r}
\newcommand{\evect}{\mathbf e}
\newcommand{\vol}{\mathrm{vol}}

\newcommand{\diam}{\mathrm{diam}}
\newcommand{\Ell}{\mathrm{Ell}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\Lip}{\mathrm{Lip}}
\newcommand{\Riem}{\mathrm{Riem}}

\newcommand{\Min}{\mathrm{Min}}
\newcommand{\Max}{\mathrm{Max}}

\newcommand{\dfn}[1]{\emph{#1}\index{#1}}

\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}

\newcommand{\loc}{\mathrm{loc}}
\newcommand{\cpt}{\mathrm{cpt}}

\def\Japan#1{\left \langle #1 \right \rangle}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{badtheorem}[theorem]{``Theorem"}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{sublemma}[theorem]{Sublemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{assumption}[theorem]{Assumption}

\newtheorem{mainthm}{Theorem}
\renewcommand{\themainthm}{\Alph{mainthm}}

\newtheorem{claim}{Claim}[theorem]
\renewcommand{\theclaim}{\thetheorem\Alph{claim}}
% \newtheorem*{claim}{Claim}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}

\newtheorem{exercise}[theorem]{Discussion topic}
\newtheorem{homework}[theorem]{Homework}
\newtheorem{problem}[theorem]{Problem}

\makeatletter
\newcommand{\proofpart}[2]{%
  \par
  \addvspace{\medskipamount}%
  \noindent\emph{Part #1: #2.}
}
\makeatother



\numberwithin{equation}{section}


% Mean
\def\Xint#1{\mathchoice
{\XXint\displaystyle\textstyle{#1}}%
{\XXint\textstyle\scriptstyle{#1}}%
{\XXint\scriptstyle\scriptscriptstyle{#1}}%
{\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
\!\int}
\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$ }
\vcenter{\hbox{$#2#3$ }}\kern-.6\wd0}}
\def\ddashint{\Xint=}
\def\dashint{\Xint-}

\usepackage[backend=bibtex,style=alphabetic,giveninits=true]{biblatex}
\renewcommand*{\bibfont}{\normalfont\footnotesize}
\addbibresource{best_curl.bib}
\renewbibmacro{in:}{}
\DeclareFieldFormat{pages}{#1}

\newcommand\todo[1]{\textcolor{red}{TODO: #1}}


\begin{document}
\begin{abstract}
	Minimizers of the comass
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
In this paper we study the problem of minimization of the \dfn{comass}
\begin{equation}\label{comass}
L(F) := \sup_{\sigma \in \Chain_2(M)} \frac{1}{|\sigma|} \int_\sigma F
\end{equation}
of a closed $2$-form in a Riemannian $3$-fold $M$, subject to a constraint on the cohomology of $F$.
Here $\Chain_2(M)$ denotes the space of oriented $2$-chains in $M$, and $|\sigma|$ is the area of $\sigma$.
This problem is the analogue on $3$-manifolds of the problem of finding a best Lipschitz map on a surface.

Sayeth Thurston \cite[Abstract]{Thurston98}:
\begin{quote}
I currently think that a characterization of minimal stretch maps should be possible in a considerably more general context ... and it should be feasible with a simpler proof based on more general principles -- in particular, the max flow min cut principle, convexity, and $L^0 \leftrightarrow L^\infty$ duality.
\end{quote}

Let $3 < p < \infty$ and $\frac{1}{p} + \frac{1}{q} = 1$.
We call a closed $2$-form \dfn{$p$-light} if it minimizes the $L^p$ norm, among all cohomologous forms.
We say that it has \dfn{best comass} if it minimizes the comass (\ref{comass}).
This problem is one of convex optimization, so it has a dual problem, namely the $q$-Laplacian.
Our first main theorem is a combination of Propositions \ref{existence infinity} and \ref{existence 1}.

\begin{mainthm}
Let $\rho \in H^2(M, \RR)$ be a cohomology class.
Let $(F_p, u_q)$ be the family of dual pairs of $p$-light forms and $q$-harmonic functions, suitably normalized, with $[F_p] = \rho$.
Then there exists a pair $(F, u)$ such that as $p \to \infty$, $F_p \to F$ weakly in $L^r$ for any $3 < r < \infty$, and $u_q \to u$ weakly in $BV$.
Moreover, $F$ has best comass in $\rho$, $u$ is $1$-harmonic, and we have the duality relation 
\begin{equation}\label{max flow mean cut}
L|\dif u| = \langle \dif u, \star F\rangle,
\end{equation}
$|\dif u|$-almost everywhere, where $L$ is the best comass constant.
\end{mainthm}

We call $F$ an \dfn{$\infty$-light form}.
Since (\ref{max flow mean cut}) asserts a form of convex duality between $\infty$-light forms and $1$-harmonic functions, we view it as the analogue of the max flow min cut principle alluded to by Thurston.

A surface $N \subseteq M$ is \dfn{calibrated} by a closed $2$-form $F$ if $\|F\|_{L^\infty} \leq 1$ and the pullback of $F$ to $N$ is the area form on $N$ \cite{Harvey82}.
In that case, the mean curvature of $N$ is 
\begin{equation}\label{calibrated surfaces are minimal}
H_N = \nabla \cdot \normal_N = \nabla \cdot (\star F)^\sharp = \star \dif F = 0,
\end{equation}
so that $N$ is minimal. 
In particular, if a lamination $\lambda$ is calibrated by $F$ (in the sense that every leaf of $\lambda$ is calibrated), then $\lambda$ is minimal.

If $F$ calibrates a surface $N$, then it is clear from the definitions that $N$ is contained in the locus on which $F$ attains its comass.
If $F$ is a continuous form of best comass, then $F/L$ calibrates a minimal lamination \cite{bangert_cui_2017}.
On the other hand, every $1$-harmonic function $u$ defines a minimal lamination $\lambda$ with Ruelle-Sullivan current $\dif u$ \todo{Cite me}.
The lamination $\lambda$ induced by the $1$-harmonic conjugate of an $\infty$-light form $F$ plays the role of Thurston's canonical lamination \cite{Thurston98}, so we call it a \dfn{Thurston lamination} associated to $[F]$.
Our next theorem, Theorem \ref{MCL contains Thurston}, is similar to \cite[Theorem 5.1]{bangert_cui_2017}, but it does not require that $F$ is continuous,shows that the lamination $\lambda$ calibrated by a form of best comass $F$ as only dependent on the cohomology class $[F]$, and identifies $\lambda$ as a familiar object that was already studied in \cite{daskalopoulos2020transverse}.

\begin{mainthm}
Let $\lambda$ be a Thurston lamination associated to $\rho \in H^2(M, \RR)$, and let $F$ be a form of best comass representing $\rho$.
Then $F/L$ calibrates $\lambda$. 
\end{mainthm}

For a measured oriented lamination $\lambda$, we let $[\lambda] \in H_2(M, \RR)$ be the homology class of the associated Ruelle-Sullivan current, and $|\lambda|$ its area -- that is, the mass of the current.
The next theorem is Theorem \ref{L equals K}, and is an analogue of the $L = K$ theorem of Thurston.

\begin{mainthm}
Let $\rho \in H^2(M, \RR)$, let $L$ be the best comass constant of $\rho$, and let $K$ be the supremum of $\langle \rho, [\lambda]\rangle/|\lambda|$, taken over all measured oriented laminations $\lambda$.
Then $L = K$, and the supremum in $K$ is attained by any Thurston lamination for $\rho$.
\end{mainthm}

%%%%%%%%%%%%%%%%%%%%%
\subsection{Outline of the paper}

%%%%%%%%%%%%%%%%%%%%%
\subsection{Notation}
We write $\Omega^\ell$, $Z^\ell$, and $B^\ell$ for the spaces of $\ell$-forms, closed $\ell$-forms ($\ell$-cocycles), and exact $\ell$-forms ($\ell$-coboundaries) respectively.
We reserve $H^\ell$ for cohomology and write $W^{s, p}$ for Sobolev spaces.
 
%%%%%%%%%%%%%%%%%%%%%%
\subsection{Acknowledgements}
George, Karen, Tom Goodwillie, Kaya Ferendo ...

This research was supported by the National Science Foundation's Graduate Research Fellowship Program under Grant No. DGE-2040433.


% \subsection{Normal trace theorem}
% We follow \cite[Chapter I, Theorem 1.2]{temam2016navier}, which handles the $p = 2$ case of the normal trace theorem.

% \begin{proposition}[normal trace theorem]
% Let $\iota: \Sph^{d - 1} \to \Ball^d$ be the inclusion map.
% The pullback is a bounded linear operator 
% $$\iota^*: L^\infty(\Ball^d, Z^{d - 1}) \to L^\infty(\Sph^{d - 1}, Z^{d - 1}).$$
% In particular, for every $\alpha \in L^\infty(\Ball^d, Z^{d - 1})$ and every $f \in W^{1, 1}(\Ball^d)$, we have integration by parts:
% \begin{equation}\label{Stokes trace}
% 	\int_{\Sph^{d - 1}} f\alpha = \int_{\Ball^d} \dif f \wedge \alpha.
% \end{equation}
% \end{proposition}
% \begin{proof}
% By the inverse trace theorem \cite[Teorema 1.II]{Gagliardo1957}, there exists a (possibly nonlinear and discontinuous) right inverse $T$ of the trace map $W^{1, 1}(\Ball^d) \to L^1(\Sph^{d - 1})$ which satisfies 
% $$\|Tf\|_{W^{1, 1}(\Ball^d)} \lesssim \|f\|_{L^1(\Sph^{d - 1})}.$$
% We use $T$ to formally define an $\ell$-current $\iota^* \alpha$ on $\Sph^{d - 1}$ by setting, for every $f \in C^\infty(\Sph^{d - 1})$,
% $$\int_{\Sph^{d - 1}} f\iota^* \alpha := \int_{\Ball^d} \dif(Tf) \wedge \alpha.$$
% To show that this current is well-defined, it suffices to show that for any $f$, 
% $$\left|\int_{\Sph^{d - 1}} f\iota^* \alpha\right| \lesssim_\alpha \|f\|_{L^1(\Sph^{d - 1})}.$$
% In fact, 
% \begin{align*}
% \left|\int_{\Sph^{d - 1}} f \iota^*\alpha\right|
% &= \left|\int_{\Ball^d} \dif(Tf) \wedge \alpha\right| \leq \|Tf\|_{W^{1, 1}(\Ball^d)} \|\alpha\|_{L^\infty(\Ball^d)} \\
% &\lesssim \|f\|_{L^1(\Sph^{d - 1})} \|\alpha\|_{L^\infty(\Ball^d)}.
% \end{align*}
% It follows that $\iota^* \alpha$ is well-defined.
% Moreover, by Stokes' theorem, that $\iota^* \alpha$ agrees with the usual definition of pullback on continuous forms $\alpha \in C^0(M, Z^{d - 1})$.
% Finally, we estimate 
% \begin{align*}
% \|\iota^* \alpha\|_{L^\infty(\Sph^{d - 1})}
% &= \sup_{\|f\|_{L^1(\Sph^{d - 1})} = 1} \int_{\Sph^{d - 1}} f \iota^*\alpha \\
% &\lesssim \sup_{\|f\|_{L^1(\Sph^{d - 1})} = 1} \|f\|_{L^1(\Sph^{d - 1})} \|\alpha\|_{L^\infty(\Ball^d)} \\
% &= \|\alpha\|_{L^\infty(\Ball^d)}. \qedhere 
% \end{align*}
% \end{proof}

% \begin{corollary}\label{trace on cycles}
% Let $N$ be a smooth embedded hypersurface in $M$ and $1 < p < \infty$.
% \begin{enumerate}
% \item \label{pullback bounded} The pullback is a bounded linear operator
% $$\iota^*_N: L^\infty(M, Z^{d - 1}) \to L^\infty(N, Z^{d - 1}).$$
% \item \label{integral continuous} For $F \in L^\infty(M, Z^{d - 1})$, and $N$ closed, the integral $\int_N F$ is well-defined, and depends continuously on $F$ for the weak topology on $L^p$.
% \item \label{cohomology exists} For $F \in L^\infty(M, Z^{d - 1})$, the cohomology class of $F$ is well-defined, and depends continuously on $F$ for the weak topology on $L^p$.
% \end{enumerate}
% \end{corollary}
% \begin{proof}
% To prove (\ref{pullback bounded}) we may use a partition of unity to work in a small ball $U$ in $N$, and then we may realize a collar neighborhood $V$ of $U$ in $M$ as a manifold-with-boundary with $U \subseteq \partial V$, and choose $\chi \in C^\infty_\cpt(M)$ in (\ref{Stokes trace}) to be a cutoff which is zero on $\partial V$ except along $U$.
% The definition of
% $$\iota^*_U: L^\infty(M, Z^{d - 1}) \to L^\infty(U, \Omega^\ell)$$
% does not depend on the choice of $V$, since by Stokes' theorem the right-hand side of (\ref{Stokes trace}) will be the same.

% To obtain (\ref{integral continuous}), we first use the fact that $N$ is closed to obtain $L^\infty(N, Z^{d - 1}) \subseteq L^1(N, Z^{d - 1})$.
% To obtain the continuity, we again use a collar neighborhood $V$ of $U$ and a cutoff $\chi$.
% Then (\ref{Stokes trace}) reads 
% $$\int_U F = \int_M F \wedge \dif \chi.$$
% Since $\chi \in C^\infty_\cpt$, $\dif \chi \in L^q$ where $\frac{1}{p} + \frac{1}{q} = 1$, so if $F_n \to F$ in the weak topology on $L^p$, then $\int_M F_n \wedge \dif \chi \to \int_M F \wedge \dif \chi$, as desired.
% Letting $N$ range over representatives of every homology class, we conclude (\ref{cohomology exists}) as a consequence of (\ref{integral continuous}).
% \end{proof}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Best comass and \texorpdfstring{$\infty$-light forms}{infinity-light forms}}
\subsection{Forms of best comass}
\begin{definition}
For a closed $2$-form $F$ and a subdomain $\Omega$, the \dfn{comass} of $F$ in $\Omega$ is
$$L_\Omega(F) := \sup_{\sigma \in \Chain_2(\Omega)} \frac{1}{|\sigma|} \int_\sigma F.$$
We let $L(F) := L_M(F)$.
\end{definition}

This definition is well-defined if $F$ is continuous, or has a trace along every $2$-simplex.
We use the following proposition to take traces of $F$ whenever $F \in L^\infty$:

\begin{proposition}
Let $\Delta$ be the standard $d$-simplex, $\tau$ a $d-1$-face of $\Delta$, and $d < p < \infty$.
Then $F \mapsto \int_\tau F$ extends to a continuous linear functional on $L^p(\Delta, Z^{d - 1})$.
Moreover,
\begin{equation}\label{integral over chain is linfinity}
	\frac{1}{|\tau|} \int_\tau F \leq \|F\|_{L^\infty}.
\end{equation}
\end{proposition}
\begin{proof}
We first show that $F \mapsto \int_\tau F$ is continuous for the $L^p$ norm on smooth closed $d-1$-forms.
Let $F, F' \in C^\infty(\Delta, Z^{d - 1})$.
Since $\Delta$ is contractible, there exist $A, A'$ such that $\dif A = F$ and $\dif A' = F'$, and such that $A, A'$ are in Coulomb gauge (that is, coclosed).
By elliptic regularity and Sobolev embedding ($p > d$),
$$\|A - A'\|_{C^0} \lesssim \|\nabla(A - A')\|_{L^p} \lesssim \|F - F'\|_{L^p}.$$
So by Stokes' theorem,
$$\left|\int_\tau F - F'\right| = \left|\int_{\partial \tau} A - A'\right| \lesssim \|A - A'\|_{C^0} \lesssim \|F - F'\|_{L^p}.$$
By a mollification argument, $F \mapsto \int_\tau F$ is a continuous linear functional on $L^p(\Delta, Z^{d - 1})$.

Suppose now that $F \in L^\infty$ and choose a sequence of mollifiers $\chi_\varepsilon$ such that $\|\chi_\varepsilon\|_{L^1} = 1$.
Let $F_\varepsilon := F * \chi_\varepsilon$, so that $F_\varepsilon \to F$ in $L^p$ for any $d < p < \infty$.
By Young's inequality, 
$$\|F_\varepsilon\|_{C^0} \leq \|F\|_{L^\infty} \|\chi_\varepsilon\|_{L^1} \leq \|F\|_{L^\infty}.$$
Taking $\varepsilon \to 0$ we see that
\begin{align*}
\frac{1}{|\tau|} \int_\tau F 
&= \lim_{\varepsilon \to 0} \frac{1}{|\tau|} \int_\tau F_\varepsilon \leq \lim_{\varepsilon \to 0} \|F_\varepsilon\|_{C^0} \leq \|F\|_{L^\infty}. \qedhere
\end{align*}
\end{proof}

In particular, since any $2$-chain $\sigma$ can be written as the sums of $2$-simplices $\tau$ which can then be viewed as faces of contractible $d$-simplices $\Delta$, the integral in the definition of comass is well-defined as long as $F \in L^p$, $p > 3$, and each such integral depends continuously on $F$ in $L^p$.

\begin{definition}
	Let $\rho \in H^2(M, \RR)$ and let $F$ be a closed $2$-form representing $\rho$.
\begin{enumerate}
	\item We say that $F$ has \dfn{best comass} if $F$ is a minimizer of $L$ among all $F$ representing $\rho$.
	\item We say that $F$ has \dfn{absolutely best comass} if it has best comass, and for every small open ball $\Omega$, $F$ is a minimizer of $L_\Omega$ among all closed $2$-forms $B$ with $B|_{\partial \Omega} = F|_{\partial \Omega}$.
\end{enumerate}
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Local comass}
We will be interested in the points at which $F$ attains its comass.
One could pose this problem as the problem of computing the locus $\{|F| = \|F\|_{L^\infty}\}$.
However, $F$ is both only defined almost everywhere, and not norm-approximable by smooth functions.
So as a proxy for $|F|$, which may fail to be defined on a null set, we use the local comass, which is defined everywhere.

\begin{definition}
The \dfn{local comass} of a closed $2$-form $F$ at $x \in M$ is 
$$L(F, x) := \limsup_{\varepsilon \to 0} \sup_{\sigma \in \Chain_2(B_\varepsilon(x))} \frac{1}{|\sigma|} \int_\sigma F.$$
\end{definition}

One has
$$L(F, x) = \limsup_{\varepsilon \to 0} L_{B_\varepsilon(x)}(F)$$
but $L_{B_\varepsilon(x)}(F)$ is increasing in $\varepsilon$ (since it's a supremum over a set which grows in $\varepsilon$).
So the limit superior is actually a limit and an infimum:
$$L(F, x) = \lim_{\varepsilon \to 0} L_{B_\varepsilon(x)}(F) = \inf_{\varepsilon > 0} L_{B_\varepsilon(x)}(F)$$
and in particular $L(F, x) \leq L(F)$. Thus we have the following analogue of \cite[Lemma 4.3(a)]{Crandall2008}:

\begin{proposition}
For $F \in L^\infty(M, Z^2)$, the local comass $L(F, \cdot)$ is upper semicontinuous. \label{crandall usc}
\end{proposition}
\begin{proof}
Let $x^n \to x$ and $r > 0$. Then eventually $x^n \in B_r(x)$, hence $L(F, x^n) \leq L_{B_r(x)}(F)$ and so
\begin{align*}
\limsup_{n \to \infty} L(F, x^n) &\leq \inf_{r > 0} L_{B_r(x)}(F) = L(F, x). \qedhere 
\end{align*}
\end{proof}

It's convenient to test the local comass against a more restrictive class of $2$-chains.
For a tangent plane $S \subseteq T_x M$ and $\varepsilon > 0$, let
$$P_{S, \varepsilon} := (\exp_x)_*(S) \cap B_\varepsilon(x).$$
We refer to $P_{S, \varepsilon}$ as a \dfn{diskette} at scale $\varepsilon$.\footnote{The terminology serves to remind ourselves to how we want to test the $2$-form $F$ against small disks, analogous to how in lattice gauge theory one tests $2$-forms against small squares (or plaques) called plaquettes \cite{Gupta98}.}

The utility of diskettes is that, for a covector $\xi \in T_x' M$, we obtain a foliation of the tangent space $T_x M$ into planes, namely $(\ker \xi + t\xi^\sharp)_{t \in \RR}$.
Pushing forward this foliation by the exponential map, we obtain a foliation
$$\lambda_{\xi, \varepsilon} := (P_{\ker \xi + t\xi^\sharp, \varepsilon})_{t \in \RR}$$
of $B_\varepsilon(x)$ into diskettes which are approximately conormal to $\xi$ in the limit $\varepsilon \to 0$.
Using this foliation and Fubini's theorem, we can turn integrals of $F$ over surfaces, which are Lebesgue null, into integrals over balls, which satisfy the Lebesgue differentiation theorem.
From these considerations and the Hardy-Littlewood maximal inequality, we deduce that diskettes satisfy their own version of the Lebesgue differentiation theorem, at least on average.

\begin{proposition}[diskette Lebesgue differentiation theorem]\label{PLDT}
Let $F \in L^\infty(M, Z^2)$.
Then for almost every $x \in M$ there exists a unit covector $\xi \in T_x' M$, which depends measurably on $x$, such that
\begin{equation}\label{diskette LDT}
|F(x)| = \lim_{\varepsilon \to 0} \frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \left[\int_{P_{\ker \xi + t\xi^\sharp, \varepsilon}} F\right] \dif t
\end{equation}
where $\dif t$ is the disintegration of the Riemannian measure with respect to $\lambda_{\xi, \varepsilon}$.
\end{proposition}

To set up the proof, we define for any closed $2$-form $G$, covector $\eta \in T'_x M$, and $\varepsilon > 0$,
$$A(G, \eta, \varepsilon) := \frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \left[\int_{P_{\ker \eta + t\eta^\sharp, \varepsilon}} G\right] \dif t.$$
Also let 
$$\Pi_\eta: T' B_\varepsilon(x) \wedge T' B_\varepsilon(x) \to T' \lambda_{\eta, \varepsilon} \wedge T' \lambda_{\eta, \varepsilon}$$
be the projection from the bundle of antisymmetric $2$-tensors on $B_\varepsilon(x)$ to those which are cotangent to $\lambda_{\eta, \varepsilon}$.
Thus, if $G$ and $P := P_{\ker \eta + t\eta^\sharp, \varepsilon}$ are cooriented, then
\begin{equation}\label{cooriented integral is area integral of norm}
	\int_P G = \int_P |\Pi_\eta G| \dif A
\end{equation}
where $\dif A$ is the area element on the leaves of the foliation.

\begin{lemma}\label{PLDT nonzero}
Proposition \ref{PLDT} holds on almost all of $\{F = 0\}$ for any measurable section $\xi$ of the cosphere bundle of $M$.
\end{lemma}
\begin{proof}
We bound using (\ref{cooriented integral is area integral of norm}) and Fubini's theorem
$$|A(F, \xi(x), \varepsilon)| \leq \frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \int_{P_{\ker \xi + t\xi^\sharp, \varepsilon}} |F| \dif A \dif t = \dashint_{B_\varepsilon(x)} \star |F|.$$
Almost every $x$ such that $F(x) = 0$ is a Lebesgue point of $|F|$, and for such $x$,
$$0 \leq \limsup_{\varepsilon \to 0} |A(F, \xi(x), \varepsilon)| \leq \lim_{\varepsilon \to 0} \dashint_{B_\varepsilon(x)} \star |F| = |F(x)| = 0.$$
This implies (\ref{diskette LDT}) on almost all of $\{F \neq 0\}$.
\end{proof}

\begin{lemma}\label{PLDT continuous}
Proposition \ref{PLDT} holds if $F$ is continuous and everywhere nonzero, and $\xi := \star F/F$.
\end{lemma}
\begin{proof}
For $\varepsilon > 0$, $\xi$ is conormal to $P_{\ker \xi, \varepsilon}$, and hence $\lambda_{\xi, \varepsilon}$, at $x$.
Thus, by definition of $\xi$, $\Pi_\xi F(x) = F(x)$, so by continuity of $F$,
\begin{equation}\label{continuous forms are almost their projections}
	\|\Pi_\xi F - F\|_{C^0(B_\varepsilon(x))} \ll 1
\end{equation}
as $\varepsilon \to 0$.
By (\ref{cooriented integral is area integral of norm}), the fact that the Hodge star coorients $F$ and $\xi$, and Fubini's theorem,
\begin{align*}
A(F, \xi, \varepsilon)
&= \frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \int_{P_{\ker \xi + t\xi^\sharp, \varepsilon}} |\Pi_\xi F| \dif A \dif t
= \dashint_{B_\varepsilon(x)} \star |\Pi_\xi F|.
\end{align*}
By (\ref{continuous forms are almost their projections}) and the fact that $|F|$ is continuous (so $x$ is a Lebesgue point of $|F|$),
\begin{align*}
\dashint_{B_\varepsilon(x)} \star |\Pi_\xi F|
&= \dashint_{B_\varepsilon(x)} \star |F| + o(1) = |F(x)| + o(1).
\end{align*}
Taking the limit $\varepsilon \to 0$, we deduce (\ref{diskette LDT}) for $x$.
\end{proof}

\begin{lemma}\label{PLDT bound mass}
Suppose that $F \neq 0$ almost everywhere.
Let $\xi := \star F/|F|$, and let $G$ be a continuous, nowhere zero, closed $2$-form.
For every $\alpha > 0$,
$$\left|\left\{\liminf_{\varepsilon \to 0} |A(F, \xi, \varepsilon) - |F|| > 4\alpha\right\} \cap \{F \neq 0\}\right| \lesssim \frac{\|F - G\|_{L^1}}{\alpha}.$$
\end{lemma}
\begin{proof}
Let $\eta := \star G/|G|$.
We split up
\begin{align*}
|A(F, \xi(x), \varepsilon) - |F(x)||
&\leq |A(F, \xi(x), \varepsilon) - A(G, \xi(x), \varepsilon)| + |A(G, \xi(x), \varepsilon) - A(G, \eta(x), \varepsilon)| \\
&\qquad + |A(G, \eta(x), \varepsilon) - |G(x)|| + ||F(x)| - |G(x)||\\
&=: I_1 + I_2 + I_3 + I_4.
\end{align*}

To treat $I_1$, we apply (\ref{cooriented integral is area integral of norm}):
\begin{align*}
I_1
&= \left|\frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \int_{P_{\ker \xi + t\xi^\sharp, \varepsilon}} F - G \dif t\right| 
\leq \frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \int_{P_{\ker \xi + t\xi^\sharp, \varepsilon}} |\Pi_\xi(F - G)| \dif A \dif t.
\end{align*}
By Fubini's theorem, 
\begin{align*}
\frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \int_{P_{\ker \xi + t\xi^\sharp, \varepsilon}} |\Pi_\xi(F - G)| \dif A \dif t
&= \dashint_{B_\varepsilon(x)} \star|\Pi_\xi(F - G)| \leq \dashint_{B_\varepsilon(x)} \star |F - G|.
\end{align*}
If we let $\mathcal M$ denote the Hardy-Littlewood maximal operator, then we have just proven $I_1 \leq \mathcal M|F - G|(x)$.
So by the Hardy-Littlewood maximal inequality, for any $\varepsilon > 0$,
$$|\{I_1 > \alpha\}| \leq |\{\mathcal M |F - G| > \alpha\}| \lesssim \frac{\|F - G\|_{L^1}}{\alpha}.$$

Next, to bound $I_2$, we pull back everything into the tangent space at $x$.
We put tildes over balls or Hodge stars to indicate that they are taken with respect to the euclidean metric on $T_x M$.
After accepting a correction of size $O(\varepsilon)$ arising from the metric on $M$, 
$$I_2 = \frac{1}{\varepsilon^d |\Ball^d|} \left|\int_{-\varepsilon}^\varepsilon \left[\int_{\ker \xi + t\xi^\sharp \cap \tilde B_\varepsilon} - \int_{\ker \eta + t\eta^\sharp \cap \tilde B_\varepsilon}\right] G \dif t\right| + O(\varepsilon)$$
which we can rewrite as
$$I_2 = \left|\dashint_{\tilde B_\varepsilon(0)} \tilde \star [(G, \xi(x)^\sharp) - (G, \eta(x)^\sharp)]\right| + O(\varepsilon) \leq \dashint_{\tilde B_\varepsilon(0)} \tilde \star |G| \cdot |\xi(x) - \eta(x)| + O(\varepsilon).$$
Since $G$ is continuous, on $\tilde B_\varepsilon(0)$, $|G| = |G(x)| + o(1)$ as $\varepsilon \to 0$.
Therefore by the reverse triangle inequality,
\begin{align*}
I_2
&\leq |G(x)| |\xi(x) - \eta(x)| + o(1) \\
&= |G(x)| \left|\frac{G(x)}{|G(x)|} - \frac{F(x)}{|F(x)|}\right| + o(1) \\
&= \left|G(x) - \frac{|G(x)|}{|F(x)|} F(x)\right| + o(1) \\
&= |G(x) - F(x)| + \left|1 - \frac{|G(x)|}{|F(x)|}\right| |F(x)| + o(1)\\
&= |G(x) - F(x)| + ||F(x)| - |G(x)|| + o(1) \\
&\leq 2|G(x) - F(x)| + o(1).
\end{align*}
By Markov's inequality, we conclude that 
$$\left|\left\{\liminf_{\varepsilon \to 0} I_2 > \alpha\right\}\right| \leq \left|\left\{|F - G| > \frac{\alpha}{2}\right\}\right| \leq \frac{2\|F - G\|_{L^1}}{\alpha}.$$

By Lemma \ref{PLDT continuous}, $I_3 = o(1)$, or in other words,
$$\left|\left\{\liminf_{\varepsilon \to 0} I_3 > \alpha\right\}\right| = 0.$$
We bound $I_4$ for any $\varepsilon > 0$ using the reverse triangle inequality and Markov's inequality as 
$$|\{I_4 > \alpha\}| \leq |\{|F(x) - G(x)| > \alpha\}| \leq \frac{\|F - G\|_{L^1}}{\alpha}.$$
Adding up $I_1, \dots, I_4$, we conclude that 
\begin{align*}
\left|\left\{\liminf_{\varepsilon \to 0} |A(F, \xi, \varepsilon) - |F|| > 4\alpha\right\}\right| \leq \sum_{i=1}^4 \left|\left\{\liminf_{\varepsilon \to 0} I_i > \alpha\right\}\right| &\lesssim \frac{\|F - G\|_{L^1}}{\alpha}. \qedhere 
\end{align*}
\end{proof}

\begin{proof}[Proof of Proposition \ref{PLDT}]
We write $M = \{F = 0\} \sqcup \{F \neq 0\}$.
By Lemma \ref{PLDT nonzero}, the claim holds on almost all of $\{F = 0\}$, so we may assume that $F \neq 0$ almost everywhere.
In particular, $\xi := \star F/|F|$ is well-defined almost everywhere.
We choose a continuous, everywhere nonzero, closed $2$-form $G$ so that $\|F - G\|_{L^1} \leq \delta \alpha^3$ for any $\delta > 0$, so that, after applying Lemma \ref{PLDT bound mass} with $\alpha = \frac{1}{m}$,
$$\left|\left\{\lim_{\varepsilon \to 0} |A(F, \xi, \varepsilon) - |F|| > 0\right\}\right| \leq \sum_{m=1}^\infty \left|\left\{\lim_{\varepsilon \to 0} |A(F, \xi, \varepsilon) - |F|| > \frac{1}{m}\right\}\right| \lesssim \sum_{m=1}^\infty \frac{\delta}{m^2} \lesssim \delta.$$
The claim follows when we take $\delta \to 0$.
\end{proof}

As a consequence of the diskette Lebesgue differentiation theorem, we obtain an analogue for the comass of a familiar result about the Lipschitz seminorm \cite[Lemma 4.3]{Crandall2008}.

\begin{proposition}\label{crandall}
Let $F \in L^\infty(M, Z^2)$. Then:
\begin{enumerate}
% \item If $F(x)$ exists then $L(F, x) \geq |\dif A(x)|$. \label{crandall dA bounds LA}
% \item If $L(A, x) = 0$, then $\dif A(x)$ exists and $\dif A(x) = 0$. \label{crandall zero LA implies diffble}
\item For almost every $x \in M$,
$$|F(x)| \leq L(F, x).$$
\label{crandall LDT}
\item The local comass is bounded, and \label{crandall linfinity}
$$L(F) = \max_{x \in M} L(F, x) = \|F\|_{L^\infty}.$$
% \item Let $\iota^* F$ be the pullback of $F$ to a $2$-chain $\sigma$. Then
% $$\|\iota^* F\|_{L^\infty(\sigma)} \leq \|F\|_{L^\infty}.$$
% \label{crandall normal trace is contraction}
% \item If $\sigma \in \Chain_2$ then \label{crandall best curl is ABC}
% $$\frac{1}{|\sigma|} \int_\sigma F \leq \max_{x \in \sigma} L(F, x).$$
\end{enumerate}
\end{proposition}
\begin{proof}
% We now bound for a net of diskettes $R_{ij}^\varepsilon(x)$ using (\ref{riemann diskette})
% \begin{equation}\label{difference quotients}
% 	\limsup_{\varepsilon \to 0} \frac{|A_j(x + \varepsilon \partial_i) - A_j(x) - A_i(x + \varepsilon \partial_j) + A_i(x)|}{\varepsilon}
% = \limsup_{\varepsilon \to 0} \frac{1}{|R_{ij}^\varepsilon(x)|} \left|\int_{\partial R_{ij}^\varepsilon(x)} A\right| \leq L(A, x).
% \end{equation}
% If the first limit superior is actually a limit, then it is the definition of $|\dif A_{ij}(x)|$.
% So if $\dif A_{ij}(x)$ exists we conclude $|\dif A_{ij}(x)| \leq L(A, x)$, which proves (\ref{crandall dA bounds LA}).
% On the other hand, the corresponding limit \emph{inferior} must be nonnegative, so if $L(A, x) = 0$, the first limit superior in (\ref{difference quotients}) is actually a limit and we obtain $\dif A_{ij}(x) = 0$, proving (\ref{crandall zero LA implies diffble}).

% Now let $\sigma$ be a $2$-chain and $\varepsilon > 0$.
% Then we may write $\sigma = \sum_{n=1}^N \sigma_n(\varepsilon)$ where $\sigma_n(\varepsilon)$ is a $2$-simplex such that $\sigma_n(\varepsilon) \in \Chain_2(B_\varepsilon(x_n(\varepsilon)))$ for some $x_n(\varepsilon) \in \sigma$, and the $\sigma_n(\varepsilon)$ are almost disjoint.
% Then 
% $$\frac{1}{|\sigma|} \left|\int_\sigma F\right| \leq \sum_{n=1}^{N_\varepsilon} \frac{|\sigma_n(\varepsilon)|}{|\sigma|} \frac{1}{|\sigma_n(\varepsilon)|} \left|\int_{\sigma_n(\varepsilon)} F\right| \leq \sum_{n=1}^{N_\varepsilon} \frac{|\sigma_n(\varepsilon)|}{|\sigma|} L_{B_{\varepsilon}(x_n(\varepsilon))}(F).$$
% Since this inequality is true for every $\varepsilon$, it remains true in the limit superior as $\varepsilon \to 0$:
% $$\frac{1}{|\sigma|} \left|\int_\sigma F\right| \leq \limsup_{\varepsilon \to 0} \sum_{n=1}^{N_\varepsilon} \frac{|\sigma_n(\varepsilon)|}{|\sigma|} L_{B_{\varepsilon}(x_n(\varepsilon))}(F).$$
% Let $(y_m(\varepsilon))_{m \in \NN}$ be a maximizing sequence for $L_{B_\varepsilon(\cdot)}(F)$ subject to $y_m(\varepsilon) \in \sigma$.
% Such a sequence exists, since we trivially have
% $$\sup_{x \in \sigma} L_{B_\varepsilon(x)}(F) \leq \|F\|_{L^\infty} < \infty.$$
% After passing to a subsequence in $m$, we may assume that $y_m(\varepsilon) \to y(\varepsilon)$ for some $y(\varepsilon) \in \sigma$.
% Then for $m$ large depending on $\varepsilon$, $y_m(\varepsilon) \in B_{2\varepsilon}(y(\varepsilon))$, and it follows that
% \begin{align*}
% \limsup_{\varepsilon \to 0} \sum_{n=1}^{N_\varepsilon} \frac{|\sigma_n(\varepsilon)|}{|\sigma|} L_{B_{\varepsilon}(x_n(\varepsilon))}(F)
% &\leq \limsup_{\varepsilon \to 0} \sum_{n=1}^{N_\varepsilon} \frac{|\sigma_n(\varepsilon)|}{|\sigma|} \lim_{m \to \infty} L_{B_\varepsilon(y_m(\varepsilon))}(F) \\
% &\leq \limsup_{\varepsilon \to 0} \lim_{m \to \infty} L_{B_\varepsilon(y_m(\varepsilon))}(F) \\
% &\leq \limsup_{\varepsilon \to 0} L_{B_{2\varepsilon}(y(\varepsilon))}(F).
% \end{align*}
% We then pass to a subsequence in $\varepsilon$ along which the limit superior is obtained, and then a further subsequence along which $y(\varepsilon) \to y$ for some $y \in \sigma$.
% Let $\delta > 0$; then $B_{2\varepsilon}(y(\varepsilon)) \subseteq B_\delta(y)$ for any sufficiently small $\varepsilon$, hence 
% $$\limsup_{\varepsilon \to 0} L_{B_{2\varepsilon}(y(\varepsilon))}(F) \leq L_{B_\delta(y)}(F).$$
% Plugging everything in and taking a limit superior in $\delta$,
% $$\frac{1}{|\sigma|} \left|\int_\sigma F\right| \leq \limsup_{\delta \to 0} L_{B_\delta(y)}(F) = L(F, y),$$
% implying (\ref{crandall best curl is ABC}).

We bound for almost every $x \in M$ using the diskette Lebesgue differentiation theorem
\begin{align*}
|F(x)|
&= \lim_{\varepsilon \to 0} \frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty \int_{P_{\ker \xi + t\xi^\sharp, \varepsilon}} F \dif t \\
&\leq \lim_{\varepsilon \to 0} L_{B_\varepsilon(x)}(F) \frac{1}{|B_\varepsilon(x)|} \int_{-\infty}^\infty |P_{\ker \xi + t\xi^\sharp, \varepsilon}| \dif t.
\end{align*}
Since $\dif t$ is the disintegration of the Riemannian measure with respect to the foliation $\lambda_{\xi, \varepsilon}$ of $B_\varepsilon(x)$,  it follows that 
$$\int_{-\infty}^\infty |P_{\ker \xi + t\xi^\sharp, \varepsilon}| \dif t = |B_\varepsilon(x)|.$$
Therefore
$$|F(x)| \leq \lim_{\varepsilon \to 0} L_{B_\varepsilon(x)}(F) = L(F, x).$$
This proves (\ref{crandall LDT}). In particular, if we combine this estimate with (\ref{integral over chain is linfinity}), then
$$\sup_{x \in M} L(F, x) \leq L(F) \leq \|F\|_{L^\infty} \leq \sup_{x \in M} L(F, x).$$
The inequalities collapse; moreover, $L(F, \cdot)$ attains its supremum since it is upper semicontinuous.
This proves (\ref{crandall linfinity}).
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis of the \texorpdfstring{$q$-Laplacian as $q \to 1$}{q-Laplacian}}
It is well-known that one can obtain a $1$-harmonic function $u$ by taking limits of $q$-harmonic functions $u_q$ as $q \to 1$, as in \cite{Mazon14}.
However, it does not immediately follow that $\|\dif u_q\|_{L^q}$ converges to the total variation of $u$, because it is not clear that $|\dif u_q|^q - |\dif u_q|$ is small in $L^1$ as $q \to 1$; essentially, this is because we know that $\dif u_q \in L^q$ but not that $\dif u_q$ lies in the Orcliz space $L^q \log L$.
In this section, we regularize the $q$-harmonic function $u_q$ using a truncated heat kernel, and use this to estimate the total variation of $u$ in terms of $\|\dif u_q\|_{L^q}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A truncated heat kernel}
Let
$$\tilde H_t(r, \theta) = (4\pi t)^{-\frac{d}{2}} \exp\left(-\frac{r^2}{4t}\right)$$
be the heat kernel on $\RR^d$.
Choose a smooth radial function $H_t$ on $\RR^d$ such that $H_t = \tilde H_t$ on $r \leq t^{1/4}$, $0 \leq H_t \leq \tilde H_t$ for $r \in [t^{1/4}, 2t^{1/4}]$, and $H_t = 0$ for $r \geq 2 t^{1/4}$.
For $f \in C^0_\cpt(\RR^d)$, we let $f(t) := H_t * f$, so that $t \mapsto f(t)$ is an approximate solution of the heat equation.
However, this truncated heat kernel preserves compact support: if $\supp f \subseteq B_\delta(0)$, then
$$\supp f(t) \subseteq B_{\delta + 2t^{1/4}}(0).$$
On the other hand, since $H_t$ is close to the true heat kernel $\tilde H_t$ in every $L^p$ space, many of the same estimates apply.

\begin{lemma}\label{approximation by heat kernels}
Uniformly in $q \in [1, \infty]$ and $0 < t \ll 1$,
$$\|H_t - \tilde H_t\|_{L^q} \lesssim e^{-d/\sqrt t}.$$
\end{lemma}
\begin{proof}
From the definitions,
$$|H_t - \tilde H_t| \leq 1_{r^4 > t} \tilde H_t,$$
so we integrate 
\begin{align*}
\|H_t - \tilde H_t\|_{L^q}^q
&\leq \int_{\RR^d} 1_{|x|^4 > t} \tilde H_t(x)^q \dif x \lesssim t^{-\frac{d}{2}} \int_{\RR^d} 1_{|x|^4 > t} \exp\left(-\frac{q|x|^2}{4t}\right) \dif x.
\end{align*}
With the change of variables $y = x \sqrt{\frac{q}{4t}}$, we obtain 
\begin{align*}
t^{-\frac{d}{2}} \int_{\RR^d} 1_{|x|^4 > t} \exp\left(-\frac{q|x|^2}{4t}\right) \dif x
&= q^{\frac{d}{2}} \int_{\RR^d} 1_{|y|^2 > \frac{q}{4\sqrt t}} e^{-|y|^2} \dif y \lesssim q^{\frac{d}{2}} \exp\left(-\frac{dq}{\sqrt t}\right).
\end{align*}
\todo{Write down the details of the last step, it's very important that this error term is controlled}
Taking $q$th roots, and recalling that $q^{a/q} \sim_a 1$, we conclude the lemma.
\end{proof}

\begin{lemma}
For $\frac{1}{p} + \frac{1}{q} = 1$, $q \in [1, \infty]$, and $0 < t \ll 1$,
$$\|H_t\|_{L^q} \sim t^{-\frac{d}{2p}}.$$
\end{lemma}
\begin{proof}
We have $t^{-\frac{d}{2p}} \gg e^{-d/\sqrt t}$, so by Lemma \ref{approximation by heat kernels}, it suffices to prove the same estimate for $\tilde H_t$. But this is standard:
\begin{align*}
\|\tilde H_t\|_{L^q} &\sim t^{-\frac{d}{2}} \left[\int_{\RR^d} e^{-\frac{q|x|^2}{4t}} \dif x\right]^{\frac{1}{q}} \sim \frac{t^{d/2q}}{q^{d/2q}t^{d/2}} \sim t^{\frac{d}{2q} - \frac{d}{2}} = t^{-\frac{d}{2p}}. \qedhere
\end{align*}
\end{proof}

\begin{lemma}
One has $f(t) \to f$ in $C^0$ as $t \to 0$.
\end{lemma}
\begin{proof}
By Lemma \ref{approximation by heat kernels}, Young's inequality, and the standard fact that $\tilde H_t * f = f + o(1)$ in $C^0$ as $t \to 0$,
\begin{align*}
\|f - f(t)\|_{C^0}
& \leq \|f - \tilde H_t * f\|_{C^0} + \|\tilde H_t * f - f(t)\|_{C^0} \leq o(1) + \|\tilde H_t - H_t\|_{L^1} \|f\|_{C^0} = o(1). \qedhere
\end{align*}
\end{proof}

\todo{Follow the following outline}
\begin{enumerate}
	\item $f - f(t)$ is small in $L^q$ because it's controlled by $\|(H_{s_1} - H_{s_2}) * f\|_{L^q}$ where $s_1, s_2 \in (0, t]$
	\item By a partition of unity argument, define $f(t)$ for $f$ a $1$-form on $M$
	\item New subsection - Use Young's inequality to estimate $\|\dif u_q(t)\|_{L^q \log_+ L}$ in terms of $\|\dif u_q\|_{L^q}$
	\item Use the Fubini trick to estimate $\||\dif u_q(t)|^q - |\dif u_q(t)|\|_{L^1}$ in $L^1$ 
	\item Since $u_q(t_q)$ and $u_q$ are co-Cauchy in $BV$ for suitable $t_q \to 0$, conclude that $\|\dif u_q\|_{L^q}$ is a Cauchy sequence.
\end{enumerate}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Taking the limit}
\begin{lemma}\label{L1 convergence preserves pi1}
Let $\tilde M \to M$ be the universal cover, and let $(u_q)$ be a sequence of $\pi_1(M)$-equivariant functions on $\tilde M$ which converge in $L^1_\loc(\tilde M)$ to a function $u$ as $q \to 1$.
Then $u$ is $\pi_1(M)$-equivariant, and $[u_q] \to [u]$.
\end{lemma}
\begin{proof}
Since $u_q$ is $\pi_1(M)$-equivariant, there exists $\alpha_q \in H^1(M, \RR)$ such that for every $\gamma \in \pi_1(M)$,
\begin{equation}\label{equivariance q}
	\gamma^* u_q = u_q + \langle \alpha_q, \gamma\rangle.
\end{equation}
Let $M_{\rm fun}$ be a fundamental domain and $U_\gamma := M_{\rm fun} \cup \gamma_* (M_{\rm fun})$.

We claim that $(\alpha_q)$ has a convergent subsequence.
To see this, we first recall that $M$ has finite Betti numbers, so $H^1(M, \RR)$ is locally compact.
Therefore, if no convergent subsequence exists, there exists a $\gamma \in \pi_1(M)$ and a subsequence along which $\langle \alpha_q, \gamma\rangle \to \infty$.
Moreover, since $u_q \to u$ in $L^1_\loc$, $\|u_q\|_{L^1(M_{\rm fun})} \leq 2\|u\|_{L^1(M_{\rm fun})}$ if $q - 1$ is small enough.
But then 
$$\|u_q\|_{L^1(\gamma_* M_{\rm fun})} = \|\gamma^* u_q\|_{L^1(M_{\rm fun})} \geq \langle \alpha_q, \gamma\rangle - \|u_q\|_{L^1(M_{\rm fun})} \geq \langle \alpha_q, \gamma\rangle - 2\|u\|_{L^1(M_{\rm fun})}$$
and taking $q \to 1$ we conclude that $(u_q)$ is not compact in $L^1(\gamma_* M_{\rm fun})$, contradicting the convergence in $L^1_\loc(\tilde M)$.
So $\alpha_q \to \alpha$ for some $\alpha \in H^1(M, \RR)$.

For any $q > 1$,
\begin{align*}
\dashint_{M_{\rm fun}} \star |\gamma^* u - u - \langle \alpha, \gamma\rangle| 
&\leq \dashint_{M_{\rm fun}} \star (|\gamma^* u_q - u_q - \langle \alpha_q, \gamma\rangle| + |\gamma^* u_q - u_q| + |\gamma^* u - u|) \\
&\qquad + |\langle \alpha_q - \alpha, \gamma\rangle|.
\end{align*}
Taking $q \to 1$ and applying (\ref{equivariance q}), we conclude that $\|\gamma^* u - u - \langle \alpha, \gamma\rangle\|_{L^1} = 0$.
The claim follows when we recall $\alpha = [u]$, $\alpha_q = [u_q]$.
\end{proof}

\begin{proposition}\label{qharmonics converge}
For each $1 < q < 2$, let $u_q$ be a $q$-harmonic, $\pi_1(M)$-equivariant function on $\tilde M$, so that
$$\|u_q\|_{W^{1, q}} \lesssim 1.$$
Then there exists a $\pi_1(M)$-equivariant function $u \in BV_\loc(M)$ such that, as $q \to 1$ along a subsequence, $u_q \to u$ weakly in $BV_\loc(\tilde M)$ and strongly in $L^r$ for $1 \leq r < \frac{d}{d - 1}$.
Moreover, $[u_q] \to [u]$ and
\begin{equation}\label{convergence of Lq norms to TV}
\lim_{q \to 1} \frac{1}{q} \int_M \star |\dif u_q|^q = \int_M \star |\dif u|.
\end{equation}
\end{proposition}
\begin{proof}
We first apply the assumption that $(u_q)$ is bounded in $W^{1, q}$.
By H\"older's inequality,
$$\lim_{q \to 1} \int_M \star |\dif u_q| \leq \lim_{q \to 1} |M|^{\frac{1}{p}} \int_M \star |\dif u_q|^q \lesssim 1.$$
So by Rellich's theorem, $(u_q)$ is weakly compact in $BV$ and strongly compact in $L^r$ for $1 \leq r < \frac{d}{d - 1}$.
In particular, $\dif u_q \to \dif u$ in the weak topology of measures and $u_q \to u$ in $BV$ and $L^r$.
As the limit of $\pi_1(M)$-equivariant functions, $u$ is also $\pi_1(M)$-equivariant by Lemma \ref{L1 convergence preserves pi1}.
In particular, $\dif u$ drops to a current on $M$.

It remains to show (\ref{convergence of Lq norms to TV})
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\texorpdfstring{$\infty$-light forms}{Infinity-light forms}}
Let $M$ be a closed Riemannian threefold and fix a cohomology class $\rho$.

\subsection{\texorpdfstring{$p$-light forms and the $q$-Laplacian}{p-light forms and the q-Laplacian}}
\begin{definition}
Let $1 < p < \infty$ and let $F_p$ be a closed $2$-form with $[F_p] = \rho$.
We call $F_p$ a \dfn{$p$-light form} if it is a minimizer of $\|F_p\|_{L^p}$ among all $2$-forms representing $\rho$.
\end{definition}

We call these forms ``light'' because as $p \to \infty$ they become ``not massive'', in the sense that, as we shall later show, they will define a minimizing sequence for the comass.

Taking $p$th powers of the $L^p$ norm, we see that a closed $2$-form is $p$-light iff it is a minimizer of
$$J_p(F) := \frac{1}{p} \int_M \star |F|^p.$$

\begin{proposition}
A $2$-form $F \in L^p(M, \Omega^2)$ is $p$-light iff it solves the system
\begin{equation}\label{pMaxwell}
\begin{cases}
	\dif F = 0, \\
	\dif^*(|F|^{p - 2} F) = 0
\end{cases}
\end{equation}
in the distributional sense. Moreover, the restriction of $J_p$ to any cohomology class in $L^p(M, Z^2)$ is uniformly strictly convex.
\end{proposition}
\begin{proof}
Let us take variations of $J_p$ in the space $L^p(M, Z^2)$ of closed $2$-forms of finite $J_p$ energy.
Let $G$ be an exact $2$-form (so $F + tG$ is cohomologous to $F$), so that
$$\frac{\dif}{\dif t} J_p(F + tG) = \frac{1}{p} \int_M \star \frac{\partial}{\partial t} |F + tG|^p = \int_M \star |F + tG|^{p - 2} \langle F + tG, G\rangle.$$
Setting $t = 0$, and writing $G = \dif B$, we obtain 
$$0 = \int_M \star |F|^{p - 2} \langle F, \dif B\rangle = \int_M \star \langle \dif^*(|F|^{p - 2} F), B\rangle.$$
Thus the Euler-Lagrange equations for $J_p$ are (\ref{pMaxwell}).
The second variation is 
$$\frac{\dif^2}{\dif t^2} J_p(F + tG)\bigg|_{t = 0} = (p - 2) \int_M \star |F|^{p - 4} \langle F, G\rangle^2 + \int_M \star |F|^{p - 2} |G|^2.$$
By the Cauchy-Schwarz inequality,
$$(p - 2) \int_M \star |F|^{p - 4} \langle F, G\rangle^2 + \int_M \star |F|^{p - 2} |G|^2 \geq (p - 1) \int_M \star |F|^{p - 2} |G|^2.$$
This implies uniform strict convexity for $p > 1$.
The restriction to a cohomology class remains strictly convex, since cohomology classes are closed affine subspaces of $L^p(M, Z^2)$ and hence are themselves closed and convex.
\end{proof}

Arguing as in the proof of existence of $p$-harmonic function (see \cite[\S8.2]{evans2010partial} for the general theory, or \cite[\S2.1]{daskalopoulos2020transverse} for $p$-harmonic maps $M \to \Sph^1$ which is closely analogous to our case), one can easily check that there exists a unique $p$-light form in every cohomology class on the closed manifold $M$.
We omit the details, both because the argument is standard, and because we shall give a separate proof in Appendix \ref{Max Flow Min Cut}.

By \cite{Uhlenbeck77}, every $p$-light form is H\"older continuous. However, the continuity degenerates as $p \to \infty$, so we shall not be able to apply this fact in the sequel.

\begin{definition}
Let $F$ be a $p$-light form, let
\begin{equation}
\dif u := (-1)^{d - 1} |F|^{p - 2} \star F, \label{inverse extremality}
\end{equation}
and let $u$ be the primitive of $\dif u$ on the universal cover $\tilde M$, which is normalized to have zero mean on a fundamental domain $M_{\rm fun}$.
Then $u$ is called the \dfn{$q$-harmonic conjugate} of the $p$-light form $F$, where $\frac{1}{p} + \frac{1}{q} = 1$.
\end{definition}

If $u$ is the $q$-harmonic conjugate of a $p$-light form $F$, then $u$ is $\pi_1(M)$-equivariant, since $\dif u$ drops to a $1$-form on $M$ and hence is $\pi_1(M)$-invariant.
To be more precise, there exists a representation $\alpha: \pi_1(M) \to \RR$ of the fundamental group such that for any $\gamma \in \pi_1(M)$,
$$\gamma^* u = u + \alpha(\gamma).$$
Throughout this paper, we use the Hurcewiz theorem to identify real-valued representations of $\pi_1(M)$ with cohomology classes in $H^1(M, \RR)$, and write $[u] = \alpha$ to mean that $[u]$ is $\alpha$-equivariant.
The assertion that $[u] = \alpha$ is equivalent to the assertion that $\dif u$ is $\pi_1(M)$-invariant and drops to a form of class $\alpha$ on $M$.

Let $u$ be the $q$-harmonic conjugate of $F$.
By Poincar\'e's inequality,
$$\|u\|_{W^{1, q}(M_{\rm fun})}^q \lesssim \int_M \star |\dif u|^q = \int_M \star |F|^{(p - 1)q} = \int_M \star |F|^p < \infty$$
since $F$ is $p$-light; that is, we have $F \in L^p$ and $u \in W^{1, q}$.
From this, and H\"older's inequality, it follows that all expressions in the next proposition are well-defined.

\begin{proposition}
Let $1 < p, q < \infty$ and $\frac{1}{p} + \frac{1}{q} = 1$.
Let $F$ be a $p$-light form, and let $u$ be its $q$-harmonic conjugate. Then:
\begin{enumerate}
\item $u$ is $q$-harmonic.
\item One has 
\begin{equation}
F = |\dif u|^{q - 2} \star \dif u. \label{extremality} \\
\end{equation}
\item One has the \dfn{max flow min cut principle}
\begin{equation}\label{strong duality}
\frac{1}{q} \int_M \star |\dif u|^q + \frac{1}{p} \int_M \star |F|^p = \int_M \dif u \wedge F.
\end{equation}
\end{enumerate}
\end{proposition}

We give a more highbrow proof of the same proposition in Appendix \ref{Max Flow Min Cut}, which also justifies why (\ref{strong duality}) should be called the ``max flow min cut principle.''
However, we think there is value in providing a straightforward proof from first principles.

\begin{proof}
We first compute 
$$(q - 2)(p - 1) + p = 2,$$
so that
$$|\dif u|^{q - 2} \star \dif u = (-1)^{d - 1} |F|^{(q - 2)(p - 1)} \star \star |F|^{p - 2} F = |F|^{(q - 2)(p - 1) - (p - 2)} F = F.$$
Thus we have (\ref{extremality}), and moreover,
$$\dif \star (|\dif u|^{q - 2} \dif u) = \dif F = 0$$
so that $u$ is $q$-harmonic.
We also compute 
\begin{align*}
\int_M \star |F|^p &= \int_M \star |\dif u|^{(q - 1)p} = \int_M \star |\dif u|^q = \int_M \dif u \wedge |\dif u|^{q - 2} \star \dif u = \int_M \dif u \wedge F. \qedhere 
\end{align*}
\end{proof}


% We put any norm on $H^2(M, \RR)$ (as all are equivalent), hence $|\alpha|$ makes sense for $\alpha \in H^2(M, \RR)$.

% \begin{proposition}
% Assume $p > 2$ and $\rho \in H^2(M, \RR)$.
% Then:
% \begin{enumerate}
% \item There exists a unique $p$-light form $F_p$ whose cohomology class is $\rho$.
% \item One has
% \begin{equation}\label{Sobolev bounds for p}
% 	\|F_p\|_{L^p} \lesssim |\rho|.
% \end{equation}
% The constant is independent of $p$.
% \item $F_p$ is H\"older continuous.
% \end{enumerate}
% \end{proposition}
% \begin{proof}
% The existence and uniqueness follows from Proposition \ref{convex duality} and the fact that the Euler-Lagrange equations for $p$-light forms are given by (\ref{pMaxwell}).
% From (\ref{pMaxwell}), the fact that $p > 2$, and \cite{Uhlenbeck77}, we deduce that $F_p$ is H\"older continuous.

% Now select a basis $\xi_1, \dots, \xi_r$ of $H^2(M, \RR)$, and apply the above argument to obtain $p$-light forms $G_i$ representing $\xi_i$ for each $i \in \{1, \dots, r\}$.
% Then $\|G_i\|_{L^p}$ is bounded indepdendently of $p$ by (\ref{infinity magnetic rules p magnetic}).
% Decompose
% $$\alpha = \sum_{i=1}^r \alpha_i \xi_i.$$
% Using the fact that $F_p$ is $p$-light and cohomologous to $\sum_i \alpha_i G_i$,
% $$\|F_p\|_{L^p} \leq \left\|\sum_{i=1}^r \alpha_i G_i\right\|_{L^p} \leq \sum_{i=1}^r |\alpha_i| \|G_i\|_{L^p} \lesssim |\alpha|$$
% which proves (\ref{Sobolev bounds for p}).
% \end{proof}

% \todo{We do not get $C^\infty$ regularity on sets $\Subset \{|F_p| \neq 0\}$, or so it seems.}
% Why?
% Expanding out (\ref{pMaxwell}) with $F_{ij} = \partial_i A_j - \partial_j A_i$,
% $$0 = \partial^j(|F|^{p - 2} \partial_j A_i) - |\dif A|^{p - 2} \partial_i (\dif^* A) - |\dif A|^{p - 2} [\partial^j, \partial_i] A_j - \partial^j(|\dif A|^{p - 2}) \partial_i A_j.$$
% The first term is an elliptic operator with H\"older coefficients, the second can be gauged away, and the third is H\"older since $[\partial^j, \partial_i]$ is a connection coefficient of the metric.
% But the last term is bad.
% \todo{Maybe we can get rid of it using paraproducts?}

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\texorpdfstring{Existence of $\infty$-light forms}{Existence of infinity-light forms}}
\begin{lemma}
Let $F_p$ be a $p$-light form, and let $B$ range over closed $2$-forms cohomologous to $F_p$. Then
\begin{equation}\label{infinity magnetic rules p magnetic}
	\|F_p\|_{L^p} \leq |M|^{1/p} \inf_B \|B\|_{L^\infty}.
\end{equation}
\end{lemma}
\begin{proof}
By H\"older's inequality and the fact that $F_p$ is $p$-light,
$$\|F_p\|_{L^p} \leq \|B\|_{L^p} \leq |M|^{1/p} \|B\|_{L^\infty},$$
hence the same holds for the infimum.
\end{proof}

\begin{proposition}\label{existence infinity}
Let $\rho \in H^2(M, \RR)$.
For each $p \geq 2$, let $F_p$ be the $p$-light form representing $\rho$. Then there exists a closed $2$-form $F$ such that:
\begin{enumerate}
\item $F_p \to F$ weakly in $L^r$ along a subsequence, for any $3 < r < \infty$.
\item $F$ is a best comass representative of $\rho$.
% \item One has 
% \begin{equation}\label{Sobolev bounds for infinity}
% 	\|F\|_{L^\infty} \lesssim |\rho|.
% \end{equation}
\end{enumerate}
\end{proposition}
\begin{proof}
We roughly follow \cite[\S3]{Lindqvist14}.
Let $r > 3$, and let $B$ be an $L^\infty$ representative of $\rho$.
By H\"older's inequality and (\ref{infinity magnetic rules p magnetic}),
\begin{equation}\label{uniform bounds in p by best curl}
	\|F_p\|_{L^r} \leq |M|^{\frac{1}{r} - \frac{1}{p}} \|F_p\|_{L^p} \leq |M|^{\frac{1}{r}} \|B\|_{L^\infty}.
\end{equation}
Thus a compactness argument gives $F_p \to F$ for some $2$-form $F$, weakly in $L^r$, and 
$$\|F\|_{L^r} \leq \liminf_{p \to \infty} \|F_p\|_{L^r} \leq |M|^{\frac{1}{r}} \|B\|_{L^\infty}.$$
Diagonalizing, we may assume that $F_p \to F$ weakly in $L^r$ for every such $r$, and taking $r \to \infty$, we conclude 
\begin{equation}\label{infinity magnetics have best curl}
	\|F\|_{L^\infty} \leq \|B\|_{L^\infty}.
\end{equation}
Moreover, $[F] = \lim_{p \to \infty} [F_p] = \rho$.
So by Proposition \ref{crandall}(\ref{crandall linfinity}) and the fact that $B$ was arbitrary in (\ref{infinity magnetics have best curl}), $F$ has best comass.
%  Moreover, taking the limit as $p \to \infty$ in (\ref{Sobolev bounds for p}), we obtain (\ref{Sobolev bounds for infinity}).
\end{proof}

\begin{definition}
The $2$-form $F$ of best comass in Proposition \ref{existence infinity} is called an \dfn{$\infty$-light form}.
\end{definition}

It seems likely that the $\infty$-light form representing a cohomology class is unique.
However, this is not obvious, and we do not prove it, since this will not cause us any difficulties later.

\begin{lemma}\label{p lights approximate L}
Let $F_p$ be the $p$-light representative of $\rho$, and $L$ the best comass of $\rho$. Then 
$$\lim_{p \to \infty} \|F_p\|_{L^p} = L.$$
\end{lemma}
\begin{proof}
We follow \cite[Lemma 2.7]{daskalopoulos2020transverse}.
Let $F$ be an $\infty$-light representative of $\rho$, so by Proposition \ref{crandall}(\ref{crandall linfinity}), $\|F\|_{L^\infty} = L$.
Since $F_p$ is $p$-light, H\"older's inequality implies 
$$\|F_p\|_{L^p} \leq \|F\|_{L^p} \leq |M|^{\frac{1}{p}} L.$$
Therefore 
$$\limsup_{p \to \infty} \|F_p\|_{L^p} \leq L.$$
To prove the converse, suppose that 
$$\liminf_{p \to \infty} \|F_p\|_{L^p} \leq \tilde L < L.$$
Along a subsequence which attains the limit inferior, $F_p$ converges weakly in every $L^r$, $3 < r < \infty$, to an $\infty$-light form $\tilde F$ such that (by H\"older's inequality)
$$\|\tilde F\|_{L^r} \leq \liminf_{p \to \infty} \|F_p\|_{L^r} \leq \liminf_{p \to \infty} |M|^{\frac{1}{r}} \|\tilde F\|_{L^\infty} \leq |M|^{\frac{1}{r}} \tilde L.$$
Taking $r \to \infty$, we obtain $L(\tilde F) < L$, which contradicts the fact that $L$ is the best comass.
\end{proof}


%%%%%%%%%%%%%%%%%%%%
\subsection{\texorpdfstring{$1$-harmonic conjugates of $\infty$-light forms}{One-harmonic conjugates of infinity-light forms}}
We now construct the $1$-harmonic conjugate of an $\infty$-light form.
We cannot naively take limits in, because (\ref{inverse extremality}) may blow up as $p \to \infty$.
Instead, we have to renormalize the $q$-harmonic conjugates of $p$-light forms, as in \cite[\S3.2]{daskalopoulos2020transverse}.

To this end, fix a class $\rho \in H^2(M, \RR)$ and denote by $L$ the comass of a best comass representative of $\rho$.
Also let $k_p$ be defined by 
$$k_p^{1 - p} = \int_M \star |F_p|^p$$
where $F_p$ is the $p$-light representative of $\rho$.

\begin{definition}
The \dfn{renormalized $q$-harmonic conjugate} of a $p$-light form $F_p$ is the function $u_q: \tilde M \to \RR$ which has mean zero on $M_{\rm fun}$ and solves
$$\dif u_q = (-1)^{d - 1} k_p^{p - 1} |F_p|^{p - 2} \star F_p.$$
\end{definition}

\begin{lemma}\label{normalizations converge}
As $p \to \infty$, $k_p \to 1/L$.
\end{lemma}
\begin{proof}
We follow \cite[Lemma 3.4]{daskalopoulos2020transverse}.
By Lemma \ref{p lights approximate L},
$$\lim_{p \to \infty} k_p^{-\frac{1}{q}} = \lim_{p \to \infty} \|F_p\|_{L^p} = L.$$
Taking logarithms we see that $q^{-1} \log k_p \to -\log L$, and since $q \to 1$ the claim follows.
\end{proof}

\begin{proposition}\label{existence 1}
Let $\rho \in H^2(M, \RR)$ and let $\tilde M \to M$ be the universal cover.
For $2 < p < \infty$ and $\frac{1}{p} + \frac{1}{q} = 1$, let $u_q$ be the renormalized $q$-harmonic conjugate of the $p$-light representative of $\rho$.
Then there exists a $\pi_1(M)$-equivariant function $u \in BV_\loc(\tilde M)$ such that:
\begin{enumerate}
\item $u$ is $1$-harmonic.
\item As $q \to 1$ along a subsequence, $u_q \to u$ weakly in $BV_\loc(\tilde M)$ and strongly in $L^r$ for $1 \leq r < \infty$.
\item Let $F$ be the $\infty$-light representative of $\rho$, with best comass $L$. We have the \dfn{max flow min cut principle} that, as Radon measures,
\begin{equation}\label{1 extremality}
\dif u \wedge F = L \star |\dif u|.
\end{equation}
\end{enumerate}
\end{proposition}
\begin{proof}
We first compute using Lemma \ref{normalizations converge}
\begin{equation}\label{Lqs of qLaplace converge}
\lim_{q \to 1} \int_M \star |\dif u_q|^q = \lim_{p \to \infty} k_p^p \int_M \star |F_p|^p = \lim_{p \to \infty} k_p = \frac{1}{L}.
\end{equation}

Renormalizing (\ref{strong duality}), we obtain 
$$\frac{k_p^{-p}}{q} \int_M \star |\dif u_q|^q + \frac{1}{p} \int_M \star |F_p|^p = k_p^{1 - p} \int_M \dif u_q \wedge F_p.$$
Multiplying by $k_p^p$, we have 
\begin{equation}\label{1 strong duality before limits}
	\frac{1}{q} \int_M \star |\dif u_q|^q + \frac{k_p}{p} \int_M \star |F_p|^p = k_p \int_M \dif u_q \wedge F_p.
\end{equation}
We want to take the limit as $p \to \infty$ in (\ref{1 strong duality before limits}) to deduce
\begin{equation}\label{1 strong duality}
	L \int_M \star |\dif u| = \int_M \dif u \wedge F.
\end{equation}

First,
$$\left|\frac{1}{q} \int_M \star |\dif u_q|^q - \int_M \star |\dif u|\right| \leq \left|\frac{\|\dif u_q\|_{L^q}^q}{q} - \|\dif u_q\|_{L^1}\right| + \left|\int_M \star |\dif u_q| - \int_M \star |\dif u|\right| =: I_1 + I_2.$$
By H\"older's inequality and the fact that $\|\dif u_q\|_{L^q}$ remains bounded as $q \to 1$ (by (\ref{Lqs of qLaplace converge})), $I_1 \to 0$.
By the portmanteau theorem \todo{cite it} and the fact that $\dif u_q \to u_q$ in the weak topology of measures, $\int_M \star |\dif u_q| \to \int_M \star |\dif u|$, hence $I_2 \to 0$.
Therefore 
$$\lim_{q \to 1} \frac{1}{q} \int_M \star |\dif u_q|^q = \int_M \star |\dif u|.$$
Second, by Lemma \ref{normalizations converge},
$$\lim_{p \to \infty} \frac{k_p}{p} \int_M \star |F_p|^p = \lim_{p \to \infty} \frac{k_p}{p} = \frac{0}{L} = 0.$$
Third, let $\alpha_q := [u_q]$, so that $\alpha_q$ is the cohomology class of $\dif u_q$.
Then, by Lemma \ref{normalizations converge},
$$\lim_{p \to \infty} k_p \int_M \dif u_q \wedge F_p = \frac{1}{L} \lim_{p \to \infty} \int_M \alpha_q \wedge \rho.$$
By Lemma \ref{L1 convergence preserves pi1}, $\alpha_q \to [u] =: \alpha$, so 
$$\lim_{p \to \infty} \int_M \alpha_q \wedge \rho = \int_M \alpha \wedge \rho = \int_M \dif u \wedge F,$$
as desired.
Taking the limit in (\ref{1 strong duality before limits}) completes the proof of (\ref{1 strong duality}).

We next claim that as Radon measures, 
\begin{equation}\label{one sided extremality}
\dif u \wedge F \leq L \star |\dif u|.
\end{equation}
To see this, fix a smooth test function $f$, integrate by parts, and use H\"older's inequality:
$$\int_M f \dif u \wedge F = -\int_M u \dif f \wedge F \leq L \|u \dif f\|_{L^1}.$$
But, if $\psi$ ranges over smooth $d-1$-forms such that $\|\psi\|_{C^0} = 1$, then another application of integration by parts and H\"older's inequality gives
\begin{align*}
\|u \dif f\|_{L^1} &= \sup_\psi \int_M u \dif f \wedge \psi = \sup_\psi \int_M f \dif u \wedge \psi = \int_M \star |f \dif u| \\
&\leq \|f\|_{C^0} \int_M \star |\dif u|.
\end{align*}
Thus we have 
$$\int_M f \dif u \wedge F \leq L \|f\|_{C^0} \int_M \star |\dif u|,$$
and since $f$ is arbitrary we conclude (\ref{one sided extremality}).

Next we deduce (\ref{1 extremality}).
We reason by contradiction: if (\ref{1 extremality}) is false, then there exists an open set $U \subseteq M$ such that 
$$\int_U \dif u \wedge F < L \int_U \star |\dif u|.$$
(Indeed, strict inequality cannot point in the other direction, by (\ref{one sided extremality}).)
However, by (\ref{one sided extremality}), 
$$\int_{M \setminus U} \dif u \wedge F \leq L \int_{M \setminus U} \star |\dif u|.$$
Adding up the integrals of $\dif u \wedge F$ over $U$ and $M \setminus U$, we conclude 
$$\int_M \dif u \wedge F < L \int_M \star |\dif u|,$$
but this contradicts (\ref{1 strong duality}); thus (\ref{1 extremality}) must be true.

To round out the proof, let $X := (\star F/L)^\sharp$ be the Poincar\'e dual vector field to $F/L$. Then
$$\nabla \cdot X = \star \frac{\dif F}{L} = 0,$$
and $\|X\|_{L^\infty} \leq 1$.
Moreover, by (\ref{1 extremality}), $X$ is normal to the level sets of $u$, and hence is a witness that $u$ is $1$-harmonic \cite{Mazon14}.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A gauge-theoretic interpretation}
In order to understand the analogy between best comass $2$-forms on threefolds and best Lipschitz functions on surfaces, it is convenient to introduce the formalism of gauge theory.
We shall not need this in the sequel, but include it for the reader's interest and intuition.

Fix an \emph{integral} class $\rho \in H^2(M, \ZZ)$.
Let $F$ be a best comass $2$-form representing $\rho$.
Then there locally exists a $1$-form $A$ such that $\dif A = F$.
If $F$ is $\infty$-light, then we can find approximations $A_p$ which solve the $p$-Maxwell equation 
$$\dif^* (|\dif A_p|^{p - 2} \dif A_p) = 0$$
and so could be called \dfn{$p$-magnetic potentials}.
The $p$-magnetic potentials, and their limit, the \dfn{$\infty$-magnetic potential} $A$, are naturally viewed as (the Christoffel symbols of globally defined) connections on the line bundle $\mathscr L$ whose Chern class is $c_1(\mathscr L) = \rho$.
In particular, $A$ is only defined up to a gauge transformation.
We could call $A$ a \dfn{best curvature} connection, since it minimizes the $L^\infty$ norm of the curvature among all connections on $\mathscr L$.

\todo{Do we want to write down the equivariance relations? They're the same as for a connection, except that it's not true that the differences in the transition functions are integers.}

To compare with the case of best Lipschitz functions on surfaces, now suppose that $M$ is a surface and $\rho \in H^1(M, \ZZ)$.
As in \cite[\S2.1]{daskalopoulos2020transverse}, if $F$ is a closed $1$-form which minimizes its $L^\infty$ norm, then there exists a map $A: M \to \Sph^1$ such that $F = \dif A$.
The map $A$ is called \dfn{best Lipschitz} because it minimizes the Lipschitz norm among all maps $M \to \Sph^1$ of the same degree.
Again there is a gauge ambiguity: \cite{daskalopoulos2020transverse} treats two maps $A, A'$ as the same if $A - A'$ is a constant angle.
Of especial interest among the best Lipschitz maps are the $\infty$-harmonic maps $M \to \Sph^1$, which are approximated by solutions of the $p$-Laplace equation.

Since the best comass problem can be viewed as the best curvature problem on a line bundle, it is natural to make the following conjecture, which we shall not address in this paper.

\begin{conjecture}
For any compact gauge group $\mathbf G$, and $\mathbf G$-bundle $\mathscr E$, there exists a minimizer of the $L^\infty$ norm of the curvature among all connections on $\mathscr E$.
Moreover, one can find a minimizer by taking suitable weak limits of solutions of the \dfn{$p$-Yang-Mills equation}
$$\dif_A^* (|\dif A|^{p - 2} \dif A) = 0$$
where $\dif_A$ is the covariant derivative whose Christoffel symbols are given by $A$. 
\end{conjecture}

%%%%%%%%%%%%%%%%%%%%

\section{The maximum comass locus}
Throughout this section, let $M$ be a closed space form of dimension $3$.

\begin{definition}
Let $F$ be a form of best comass.
The \dfn{maximum comass locus} is the set $\{L(F, \cdot) = L(F)\}$.
\end{definition}

By Proposition \ref{crandall usc} and the compactness of $M$, the maximum comass locus is a nonempty closed subset of $M$.

By \todo{Cite laminations paper}, to each $1$-harmonic function $u$, we may associated a measured oriented minimal lamination $\lambda_u$, whose Ruelle-Sullivan current is $\dif u$, and whose leaves are the level sets of $u$.

\begin{definition}
Let $\rho \in H^2(M, \RR)$, let $F$ be an $\infty$-light representative of $\rho$, and let $u$ be a $1$-harmonic conjugate of $F$.
A \dfn{Thurston lamination} $\lambda$ associated to $\rho$ is $\lambda := \lambda_u$.
\end{definition}

\subsection{Calibration of Thurston laminations}
\begin{theorem}\label{MCL contains Thurston}
Let $F$ be a best comass representative of $\rho \in H^2(M, \RR)$.
Then the maximum comass locus of $F$ contains any Thurston lamination $\lambda$ associated to $\rho$.
In particular, if $L$ is the best comass of $\rho$, then $F/L$ calibrates $\lambda$.
\end{theorem}

On the first reading, the reader may wish to take $F$ to be $\infty$-light, as this already captures the essential ideas.
Throughout the proof we fix the $p$-light representative $F_p$ of $\rho$, its conjugate $q$-harmonic $u_q$, and the limiting $1$-harmonic $u$.

\begin{lemma}
Let
$$f_p := 2\langle k_p F_p, k_p F_p - L^{-1} F\rangle.$$
Then 
\begin{equation}\label{MCL contains Thurston 1}
	\lim_{p \to \infty} \int_{\{f_p \geq 0\}} \star |k_p F_p|^{p - 2} f_p = 0.
\end{equation}
\end{lemma}
\begin{proof}
The proof is very similar to \cite[Lemma 6.3]{daskalopoulos2020transverse}.
Since $F_p, F$ are cohomologous, there exists a $1$-form $\xi$ such that $\dif \xi = F_p - F$.
On the other hand, since $F_p$ is $p$-light, an integration by parts gives
\begin{align*}
\int_M |F_p|^{p - 2} \langle F_p, F_p - F \rangle = \int_M \langle \dif^*(|F_p|^{p - 2} F_p), \xi\rangle = 0.
\end{align*}
In particular,
\begin{equation}\label{MCL contains Thurston 2}
	I_p := \int_M \star |k_p F_p|^{p - 2} \langle k_p F_p, k_p F_p - k_p F\rangle = 0.
\end{equation}
On the other hand, by the Cauchy-Schwarz inequality, Lemma \ref{normalizations converge}, and the fact that 
$$\lim_{p \to \infty} \|F_p\|_{L^p} = \|F\|_{L^\infty} = L,$$
we have
\begin{align*}
\lim_{p \to \infty} I_p - \frac{1}{2} \int_M \star |k_p F_p|^{p - 2} f_p 
&= \lim_{p \to \infty} \int_M \star |k_p F_p|^{p - 2} \left\langle k_p F_p, \left(\frac{1}{L} - k_p\right) F\right\rangle \\
&\leq \lim_{p \to \infty} k_p^{p - 1} \left(\frac{1}{L} - k_p\right) \int_M \star |F_p|^{p - 1} |F| \\
&= \lim_{p \to \infty} \frac{1}{L^{p - 1}} \left(\frac{1}{L} - \frac{1}{L}\right)L^p \\
&= \lim_{p \to \infty} 1 - 1 = 0.
\end{align*}
If we plug this into (\ref{MCL contains Thurston 2}), we obtain 
\begin{equation}\label{MCL contains Thurston 3}
	\lim_{p \to \infty} \int_M \star |k_p F_p|^{p - 2} f_p = 0.
\end{equation}
If $f_p(x) \leq 0$, then by the Cauchy-Schwarz inequality,
$$|L^{-1} F|(x) \geq |k_p F_p|(x),$$
so by the Cauchy-Schwarz and Peter-Paul inequalities,
\begin{align*}
f_p(x) &\leq 2|L^{-1} F|(x) |k_p F_p|(x) - 2|k_p F_p|^2(x) \\
&\leq |L^{-1} F|(x)^2 + |k_p F_p|(x)^2 - 2|k_p F_p|^2(x) \\
&= |L^{-1} F|(x)^2 - |k_p F_p|(x).
\end{align*}
So by \cite[Lemma 6.2]{daskalopoulos2020transverse},
$$|k_p F_p|(x)^{p - 2} f_p(x) < \frac{2}{p - 2}.$$
Integrating this inequality, 
$$0 \leq \lim_{p \to \infty} \int_{\{f_p \leq 0\}} |k_p F_p|^{p - 2} f_p \leq \lim_{p \to \infty} \frac{2|M|}{p - 2} = 0.$$
Therefore, by (\ref{MCL contains Thurston 3}), we deduce (\ref{MCL contains Thurston 1}).
\end{proof}

\begin{lemma}\label{MCL contains Thurston lemma}
The set $\{L(F, \cdot) < L\}$ satisfies
$$\lim_{p \to \infty} \int_{\{L(F, \cdot) < L\}} \star |k_p F_p|^p = 0.$$
\end{lemma}
\begin{proof}
We decompose
$$\{L(F, \cdot) < L\} = \bigcup_{0 < \theta < 1} \{L(F, \cdot) < \theta L\}$$
and then it suffices to fix $\theta$ and show that the integral over $\{L(F, \cdot) < \theta L\}$ is zero.
We then use the fact that $|F|(x) \leq L(F, x)$ almost everywhere, by Proposition \ref{crandall}(\ref{crandall LDT}), to obtain
$$\{L(F, \cdot) < \theta L\} \subseteq A \cup B \cup Z$$
where 
\begin{align*}
A &:= \{|L^{-1} F|^2 \leq \theta |k_p F_p|^2 + |k_p F_p - L^{-1} F|^2\}, \\
B &:= \{|L^{-1} F|^2 \geq \theta |k_p F_p|^2 + |k_p F_p - L^{-1} F|^2\} \cap \{L(F, \cdot) < \theta L\},
\end{align*}
and $Z$ is null. 
In particular,
$$0 \leq \lim_{p \to \infty} \int_{\{L(F, \cdot) < L\}} \star |k_p F_p|^p \leq \lim_{p \to \infty} \int_A \star |k_p F_p|^p + \lim_{p \to \infty} \int_B \star |k_p F_p|^p.$$
By (\ref{MCL contains Thurston 1}),
\begin{align*}
0 &= \lim_{p \to \infty} \int_{\{f_p \geq 0\}} |k_p F_p|^{p - 2} f_p \\
&= \lim_{p \to \infty} \int_{\{f_p \geq 0\}} |k_p F_p|^{p - 2}((1 - \theta) |k_p F_p|^2 + \theta |k_p F_p|^2 + |k_p F_p - L^{-1} F|^2 - |L^{-1} F|^2).
\end{align*}
But on $A$,
\begin{align*}
|L^{-1} F|^2 &\leq \theta |k_p F_p|^2 + |k_p F_p - L^{-1} F|^2 \\
&\leq |k_p F_p|^2 + |k_p F_p - L^{-1} F|^2 \\
&= f_p + |L^{-1} F|^2
\end{align*}
which implies $f_p \geq 0$ and 
$$\lim_{p \to \infty} \int_A \star |k_p F_p|^p \leq \lim_{p \to \infty} \frac{1}{\theta} \int_{\{f_p \geq 0\}} \star |F_p|^{p - 2} f_p = 0.$$
Meanwhile, on $B$,
$$|k_p F_p|^2 \leq |k_p F_p|^2 + \theta^{-1} |k_p F_p - L^{-1} F|^2 \leq \theta^{-1} |L^{-1} F|^2 < \theta.$$
Therefore
\begin{align*}
	\lim_{p \to \infty} \int_B |k_p F_p|^p &\leq \lim_{p \to \infty} \theta^{p/2} |M| = 0. \qedhere
\end{align*}
\end{proof}

\begin{proof}[Proof of Theorem \ref{MCL contains Thurston}]
Let $U := \{L(F, \cdot) < L\}$, which is open since it is the complement of the best comass locus of $F$.
So by the portmanteau theorem \todo{Cite it}
\begin{align*}
\int_U \star |\dif u|
&\leq \lim_{q \to 1} \int_U \star |\dif u_q|
= \lim_{p \to \infty} \int_U \star |k_p F_p|^{p - 1}.
\end{align*}
Then by H\"older's inequality and Lemma \ref{MCL contains Thurston lemma},
\begin{align*}
\lim_{p \to \infty} \int_U \star |k_p F_p|^{p - 1}
&\leq \lim_{p \to \infty} |M|^{-\frac{1}{p}} \left[\int_U \star |k_p F_p|^p\right]^{\frac{1}{q}} 
= \lim_{p \to \infty} \int_U \star |k_p F_p|^p = 0.
\end{align*}
Since $\dif u$ is the Ruelle-Sullivan current for the Thurston lamination $\lambda$, $\lambda$ is contained in the best comass locus of $F$. 
Moreover, if $N$ is a leaf of $\lambda$, then the restriction of $F/L$ is, by (\ref{1 extremality}), equal to $\star \normal_N^\flat$, which is the area form of $N$.
In other words, $F/L$ calibrates $N$.
\end{proof}



%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thurston's \texorpdfstring{$L = K$}{L equals K} theorem}
If $\lambda$ is a measured oriented lamination, then we denote by $[\lambda] \in H_2(M, \RR)$ its homology class, and by $|\lambda|$ its area; namely, the mass of its Ruelle-Sullivan current $T_\lambda$.

\begin{theorem}\label{L equals K}
	Let $\rho \in H^2(M, \RR)$, and let 
	$$K := \sup_\lambda \frac{\langle \rho, [\lambda]\rangle}{|\lambda|},$$
	where $\lambda$ ranges over measured oriented laminations. Then:
\begin{enumerate}
	\item The supremum in $K$ is attained by the Thurston lamination associated to $\rho$.
	\item Let $L$ be the best comass of $\rho$. Then $L = K$.
\end{enumerate}
\end{theorem}
\begin{proof}
Fix the $\infty$-light form $F$ representing $\rho$, and let $u$ be its $1$-harmonic conjugate.

We first prove $K \leq L$.
Let $\lambda$ be a measured oriented lamination; then, since $F$ represents $\rho$ and the Ruelle-Sullivan current $T_\lambda$ represents $[\lambda]$,
$$\langle \rho, [\lambda]\rangle = \int_M F \wedge T_\lambda.$$
Let $\mu_\lambda$ be the transverse measure to $\lambda$ and $\mathscr L_\lambda$ the space of leaves of $\lambda$; then, by definition of $T_\lambda$ and the definition of comass,
$$\int_M F \wedge T_\lambda = \int_{\mathscr L_\lambda} \int_N F \dif \mu_\lambda(N) \leq L(F) \int_{\mathscr L_\lambda} |N| \dif \mu_\lambda(N) = L(F) |\lambda|.$$
By Proposition \ref{crandall}(\ref{crandall linfinity}), $F$ has best comass $L(F) = L$.
Since $\lambda$ was arbitrary, it holds that $K \leq L$.

Let $\lambda$ be the Thurston lamination, so $T_\lambda = \dif u$.
Then by the max flow min cut principle (\ref{1 extremality}),
$$\langle \rho, [\lambda]\rangle = \int_M F \wedge \dif u = L \int_M \star |\dif u| = L|\lambda|.$$
Dividing both sides by $|\lambda|$ and applying the direction we already proved,
$$K \leq L \leq \frac{\langle \rho, [\lambda]\rangle}{|\lambda|} \leq K$$
which is only possible if $L = K$ and $\lambda$ is a maximizer.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix 
\section{Max flow, min cut}\label{Max Flow Min Cut}
Let us interpret the duality between the $q$-Laplacian and $p$-light forms as a version of the max flow min cut principle. 
For a reflexive Banach space $X$, we denote by $\hat X$ its dual.
If $J: X \to \RR \cup \{+\infty\}$ is a convex function, we introduce its \dfn{Legendre transform}
\begin{align*}
	\hat J: \hat X &\to \RR \cup \{+\infty\}\\
	\xi &\mapsto \sup_{x \in X} \langle \xi, x\rangle - f(x).
\end{align*}

\begin{definition}
Let $Y$ be a reflexive Banach space equipped with a mapping $\Lambda: X \to Y$.
A function $J: Y \to \RR \cup \{+\infty\}$ is said to be a \dfn{suitable convex function} for our purposes if:
\begin{enumerate}
\item $J$ is strictly convex,
\item $J$ is lower semicontinuous and not identically $+\infty$,
\item if $|y| \to \infty$ in $Y$, then $J(y) \to +\infty$, and 
\item there exists a point $x \in X$ such that $J$ is continuous and finite at $\Lambda(x)$.
\end{enumerate}
\end{definition}

\begin{proposition}\label{abstract convex analysis}
Let $X, Y$ be reflexive Banach spaces equipped with a bounded linear operator $\Lambda: X \to Y$ of trivial kernel.
Let $J: Y \to \RR \cup \{+\infty\}$ be a suitable convex function.
Then:
\begin{enumerate}
\item There exists a unique minimizer $\underline x \in X$ of $J(\Lambda(x))$, and a maximizer $\overline \eta \in \hat Y$ of $-\hat J(-\eta)$.
\item If $\hat J$ is strictly convex, then $\overline \eta$ is the unique maximizer of $-\hat J(-\eta)$.
\item We have \dfn{strong duality}
\begin{equation}\label{abstract strong duality}
J(\Lambda(\underline x)) = -\hat J(-\overline \eta).
\end{equation}
\end{enumerate}
\end{proposition}
\begin{proof}
This is the conjunction of several standard results in convex analysis.
Let $\Gamma_0(Y)$ denote the set introduced in \cite[Chapter I, Definition 3.1]{Ekeland99}.
Then by \cite[Chapter I, Proposition 3.1]{Ekeland99}, $J \in \Gamma_0(Y)$, so by \cite[Chapter III, Theorem 4.2]{Ekeland99} and \cite[Chapter III, (4.23)]{Ekeland99}, there exist minimizers which satisfy (\ref{abstract strong duality}).
By \cite[Chapter II, Proposition 1.2]{Ekeland99}, the minimizers of $J$ and $\hat J$ are unique.
Since $\Lambda$ has trivial kernel, minimizers of $J \circ \Lambda$ are also unique.
\end{proof}

\todo{Talk about the max flow min cut principle}

We now specialize to the problem at hand.
Let $\alpha \in H^1(M, \RR)$ be a cohomology class.
Recall that we write $[u] = \alpha$ to mean that $u$ is an $\alpha$-equivariant function on $\tilde M$.

We here consider the problem
\begin{equation}\label{preprimal problem}
	\Min\{\|\dif u\|_{L^q(M_{\rm fun})}: u \in W^{1, q}(M_{\rm fun}), [u] = \alpha\},
\end{equation}
where $1 < q < \infty$; we identify two solutions if they agree by a constant.
Taking Euler-Lagrange equations, we see that (\ref{preprimal problem}) is equivalent to the $q$-Laplacian 
\begin{equation}\label{qLaplace}
\begin{cases}
	\dif^*(|\dif u|^{q - 2} \dif u) = 0 \\
	[u] = \alpha.
\end{cases}
\end{equation}

To put (\ref{preprimal problem}) in the framework of Proposition \ref{abstract convex analysis}, we shall fix a point $0 \in M_{\rm fun}$, choose a representative $1$-form (which we also call $\alpha$), and solve 
$$\begin{cases}
\dif f = \Pi^* \alpha \\
v(0) = 0.
\end{cases}$$
Thus (\ref{preprimal problem}) is equivalent to
\begin{equation}\label{primal problem}
	\Min\left\{\frac{1}{q} \int_{M_{\rm fun}} \star|\dif v + \Pi^* \alpha|^q: v \in W^{1, q}_0(M_{\rm fun})\right\}
\end{equation}
where we set $u = v + f$.

Let
$$J(\xi) := \frac{1}{q} \int_M \star|\xi + \alpha|^q,$$
defined for $\xi \in L^q(M, \Omega^1)$.
Since $v$ is traceless in (\ref{primal problem}), it is invariant, so $\dif v$ drops to a $1$-form on $M$, and (\ref{primal problem}) is the problem of minimizing $J(\dif v)$.
It is clear that $J$ is a suitable convex function on $L^q(M, \Omega^1)$ when it is equipped with the map
\begin{equation}\label{derivative on traceless}
\dif: W^{1, q}_0(M_{\rm fun}) \to L^q(M, \Omega^1),
\end{equation}
and moreover the kernel of (\ref{derivative on traceless}) is trivial.
By \cite[Chapter I, (4.9)]{Ekeland99} and \cite[Chapter I, Remark 4.1]{Ekeland99}, the Legendre transform
$$\hat J: L^p(M, \Omega^{d - 1}) \to \RR$$
of $J$, where $\frac{1}{p} + \frac{1}{q} = 1$, satisfies
\begin{equation}\label{Legendre transform}
\hat J(F) = \frac{1}{p} \int_M \star |F|^p + \int_M \alpha \wedge F
\end{equation}
and in particular is strictly convex.
Thus the convex dual problem of (\ref{primal problem}), namely the problem of maximizing $-\hat J(-F)$, is the problem
\begin{equation}\label{predual problem}
\Max\left\{- \frac{1}{p} \int_M \star |F|^p + \int_M \alpha \wedge F: F \in L^p(M, \Omega^{d - 1})\right\}.
\end{equation}

\begin{proposition}\label{convex duality}
Let $1 < p, q < \infty$ satisfy $\frac{1}{p} + \frac{1}{q} = 1$.
Then we have an isomorphism of cohomology groups
\begin{align*}
H^1(M, \RR) &\to H^{d - 1}(M, \RR) \\
\alpha &\mapsto \rho,
\end{align*}
so that the convex dual problem of the $\alpha$-equivariant $q$-Laplacian (\ref{qLaplace}) with solution $u: \tilde M \to \RR$ is the $p$-light problem (\ref{pMaxwell}) subject to $[F] = \rho$.
The problems (\ref{qLaplace}) and (\ref{pMaxwell}) both have unique solutions, and we have the relations (\ref{extremality}) and (\ref{inverse extremality}), and the strong duality theorem (\ref{strong duality}).
\end{proposition}
\begin{proof}
Since $J$ is a suitable convex function, $\hat J$ is strictly convex, and (\ref{derivative on traceless}) has trivial kernel, we may apply Proposition \ref{abstract convex analysis} to see that (\ref{qLaplace}) and (\ref{predual problem}) both have unique solutions $u, F$ which are related by strong duality (\ref{abstract strong duality}) as (\ref{strong duality}).
By uniqueness, if we exhibit $\tilde F$ defined in terms of $u$ which satisfies (\ref{strong duality}) and is a maximizer of $-\hat J$, then $F = \tilde F$.
Following \cite[Chapter IV, (2.12)]{Ekeland99}, we define $\tilde F$ to satisfy (\ref{extremality}), so that 
\begin{align*}
\int_M \dif u \wedge \tilde F 
&= \int_M \star |\dif u|^q \\
&= \frac{1}{q} \int_M \star |\dif u|^q + \frac{1}{p} \int_M \star |\dif u|^p \\
&= \frac{1}{q} \int_M \star |\dif u|^q + \frac{1}{p} \int_M \dif u \wedge \tilde F.
\end{align*}
Moreover, if we decompose 
$$\dif u = \dif v + \alpha$$
then
\begin{align*}
-\hat J(\tilde F)
&= -\frac{1}{p} \int_M \star |\dif u|^q + \int_M \alpha \wedge |\dif u|^{q - 2} \star \dif u \\
&= -\frac{1}{p} \int_M \star |\dif u|^q + \int_M \star |\dif u|^q - \int_M \dif v \wedge |\dif u|^{q - 2} \star \dif u \\
&= \frac{1}{q} \int_M \star |\dif u|^q + \int_M v \dif(|\dif u|^{q - 2} \star \dif u) \\
&= \frac{1}{q} \int_M \star |\dif u|^q = J(\dif u)
\end{align*}
where we used the fact that $\frac{1}{q} = 1 - \frac{1}{p}$, and the fact that $u$ solves the $q$-Laplace equation.
By strong duality (\ref{abstract strong duality}), it follows that $\tilde F$ is a maximizer of $-\hat J$, and therefore $F = \tilde F$; in particular, $F$ satisfies (\ref{extremality}).
Since $u$ solves the $q$-Laplace equation, (\ref{extremality}) implies that $\dif F = 0$.
We then define $\rho := [F]$.
On the other hand, 
$$(p - 2)(q - 1) + q - 2 = 0,$$
so following \cite[Lemma 3.2]{daskalopoulos2020transverse},
$$|F|^{p - 2} F = |\dif u|^{(p - 2)(q - 1) + q - 2} \star \star \dif u = (-1)^{d - 1} \dif u$$
which gives (\ref{inverse extremality}) and the equation
$$\dif^*(|F|^{p - 2} F) = 0.$$
By reflexivity, the above argument may also be reversed, so that the map $\alpha \mapsto \rho$ is invertible on the level of cohomology.
\end{proof}

\todo{Cite those glaciology papers which already include this duality}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The \texorpdfstring{$\infty$-light equation}{infinity-light equation}}
We have the following Euler-Lagrange equation for $\infty$-light forms.
Because of the lack of a good analogue for viscosity solutions for $\infty$-elliptic systems, \todo{and because we did not show that $\infty$-light forms have absolutely best comass}, the equation can only really be interpreted in a formal sense, at least as far as we are aware.
As such, we did not use it in the main body of this paper, but only include it as a curiosity item.

\todo{If we knew that $p$-Maxwell had good quantitative uniqueness, then we would have}
It remains to show that $A$ has absolutely best curl, so let $\Omega$ be a small ball and $B$ a $1$-form with $B|_{\partial \Omega} = A|_{\partial \Omega}$.
By a straightforward modification of the existence theorem, there exists a $p$-magnetic potential $B_p$ in Coulomb gauge with $B_p|_{\partial \Omega} = A|_{\partial \Omega}$ and $B \in C^{1 + \alpha}$.
By quantitative uniqueness
$$\|B_p - A\|_{C^0(\Omega)} \leq \|B_p - A_p\|_{C^0(\Omega)} + o(1) \lesssim \|A - A_p\|_{C^0(\partial \Omega)} + o(1) \ll 1.$$
Therefore $B_p \to A$ uniformly, and for $3 < q < p < \infty$ with $p$ dyadic,
$$\|\dif B_p\|_{L^q(\Omega)} \leq |\Omega|^{\frac{1}{q} -\frac{1}{p}} \|\dif B_p\|_{L^p(\Omega)} \leq |\Omega|^{\frac{1}{q} -\frac{1}{p}} \|\dif B\|_{L^p(\Omega)} \leq |\Omega|^{\frac{1}{q}} \|\dif B\|_{L^\infty(\Omega)}.$$
Then along a subsequence, $\dif B_p \to \dif A$ in $L^q(\Omega)$, so 
$$\|\dif A\|_{L^q(\Omega)} \leq |\Omega|^{\frac{1}{q}} \|\dif B\|_{L^\infty(\Omega)}.$$
Taking $q \to \infty$ we arrive at the conclusion that $F$ has absolutely best comass.

\begin{proposition}
Suppose that $F$ has absolutely best comass, regularity $C^1$, and no points with $F = 0$. Then
\begin{equation}\label{infinityMaxwell}
	F^{ij} \partial_i |F| = 0.
\end{equation}
\end{proposition}
\begin{proof}
For a covariant $2$-tensor $T$, let $T^{\rm as}$ be its antisymmetrization, and let
$$f(x, T) := |T^{\rm as}|_{g(x)}.$$
Working locally, we may write $F = \dif A$ for some $A$, which we may assume to be Coulomb gauge and therefore $C^2$.
Since $A$ has absolutely best curl and $(\nabla A)^{\rm as} = \dif A$, $A$ is an absolute minimizer (see \cite[Definition 5.1]{Barron2001}) of the essential supremum of $f(\cdot, \nabla A)$.
By the Euler-Lagrange-Aronsson formula \cite[Theorem 5.2]{Barron2001},
\begin{equation}\label{ELA}
	\left\langle \frac{\partial f}{\partial T}(x, \nabla A(x)), \dif (f(x, \nabla A(x))) \right\rangle = 0.
\end{equation}
Now
$$\dif(f(x, \nabla A(x))) = \dif |\dif A(x)|$$
and 
$$\frac{\partial f}{\partial T}(x, \nabla A(x)) = \frac{\nabla A(x)^{\rm as}}{|\nabla A(x)^{\rm as}|} = \frac{\dif A(x)}{|\dif A(x)|}.$$
We conclude the claim after multiplying both sides of (\ref{ELA}) by $|\dif A|$.
\end{proof}

\begin{corollary}
Suppose that $F$ has absolutely best comass, regularity $C^1$, and no points with $F = 0$, and $N$ is a surface whose normal vector field is annihilated by $F$.
Then $N$ is a minimal surface.
\end{corollary}
\begin{proof}
Let $V$ be a tangent vector field to $N$. Then $V(|F|) = 0$, by (\ref{infinityMaxwell}).
Therefore $|F|$ is constant along $N$, but $F$ is a continuous section of the area bundle of $N$, which is a real line bundle.
It follows that $F$ is constant along $N$, and $F/|F|$ is the area form on $N$.
In other words, $N$ is calibrated by $F$, and the claim follows from (\ref{calibrated surfaces are minimal}).
\end{proof}

\printbibliography

\end{document}
