\documentclass[reqno,11pt]{amsart}
\usepackage[letterpaper, margin=1in]{geometry}
\RequirePackage{amsmath,amssymb,amsthm,graphicx,mathrsfs,url,slashed,subcaption}
\RequirePackage[usenames,dvipsnames]{xcolor}
\RequirePackage[colorlinks=true,linkcolor=Red,citecolor=Green]{hyperref}
\RequirePackage{amsxtra}
\usepackage{cancel}
\usepackage{tikz-cd}

% \setlength{\textheight}{9.3in} \setlength{\oddsidemargin}{-0.25in}
% \setlength{\evensidemargin}{-0.25in} \setlength{\textwidth}{7in}
% \setlength{\topmargin}{-0.25in} \setlength{\headheight}{0.18in}
% \setlength{\marginparwidth}{1.0in}
% \setlength{\abovedisplayskip}{0.2in}
% \setlength{\belowdisplayskip}{0.2in}
% \setlength{\parskip}{0.05in}
%\renewcommand{\baselinestretch}{1.05}

\title{1-harmonic functions and sequences of minimal laminations}
\author{Aidan Backus}
\date{October 2022}

\newcommand{\NN}{\mathbf{N}}
\newcommand{\ZZ}{\mathbf{Z}}
\newcommand{\QQ}{\mathbf{Q}}
\newcommand{\RR}{\mathbf{R}}
\newcommand{\CC}{\mathbf{C}}
\newcommand{\DD}{\mathbf{D}}
\newcommand{\PP}{\mathbf P}
\newcommand{\MM}{\mathbf M}
\newcommand{\II}{\mathbf I}
\newcommand{\Hyp}{\mathbf H}
\newcommand{\Sph}{\mathbf S}
\newcommand{\Group}{\mathbf G}
\newcommand{\GL}{\mathbf{GL}}
\newcommand{\Orth}{\mathbf{O}}
\newcommand{\SpOrth}{\mathbf{SO}}
\newcommand{\Ball}{\mathbf{B}}

\newcommand*\dif{\mathop{}\!\mathrm{d}}

\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\MeasLam}{MeasLam}
\DeclareMathOperator{\MinLam}{MinLam}
\DeclareMathOperator{\Lam}{Lam}
\DeclareMathOperator{\supp}{supp}

\newcommand{\Leaves}{\mathscr L}
\newcommand{\Hypspace}{\mathscr H}

\newcommand{\Two}{\mathrm{I\!I}}


\newcommand{\Hilb}{\mathcal H}
\newcommand{\Homology}{\mathrm H}
\newcommand{\normal}{\mathbf n}
\newcommand{\radial}{\mathbf r}
\newcommand{\evect}{\mathbf e}
\newcommand{\vol}{\mathrm{vol}}

\newcommand{\Bmu}{\boldsymbol \mu}
\newcommand{\Bnu}{\boldsymbol \nu}
\newcommand{\Blambda}{\boldsymbol \lambda}

\newcommand{\pic}{\vspace{30mm}}
\newcommand{\dfn}[1]{\emph{#1}\index{#1}}

\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}

\newcommand{\loc}{\mathrm{loc}}
\newcommand{\cpt}{\mathrm{cpt}}

\def\Japan#1{\left \langle #1 \right \rangle}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{badtheorem}[theorem]{``Theorem"}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{sublemma}[theorem]{Sublemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{assumption}[theorem]{Assumption}

\newtheorem{mainthm}{Theorem}
\renewcommand{\themainthm}{\Alph{mainthm}}

% \newtheorem{claim}{Claim}[theorem]
% \renewcommand{\theclaim}{\thetheorem\Alph{claim}}
\newtheorem*{claim}{Claim}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}

\newtheorem{exercise}[theorem]{Discussion topic}
\newtheorem{homework}[theorem]{Homework}
\newtheorem{problem}[theorem]{Problem}

\makeatletter
\newcommand{\proofpart}[2]{%
  \par
  \addvspace{\medskipamount}%
  \noindent\emph{Part #1: #2.}
}
\makeatother



\numberwithin{equation}{section}


% Mean
\def\Xint#1{\mathchoice
{\XXint\displaystyle\textstyle{#1}}%
{\XXint\textstyle\scriptstyle{#1}}%
{\XXint\scriptstyle\scriptscriptstyle{#1}}%
{\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
\!\int}
\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$ }
\vcenter{\hbox{$#2#3$ }}\kern-.6\wd0}}
\def\ddashint{\Xint=}
\def\dashint{\Xint-}

\usepackage[backend=bibtex,style=numeric]{biblatex}
\renewcommand*{\bibfont}{\normalfont\footnotesize}
\addbibresource{topics.bib}
\renewbibmacro{in:}{}
\DeclareFieldFormat{pages}{#1}


\begin{document}
\begin{abstract}
We collect several results relating different notions of convergence of laminations, especially minimal laminations.
We also give a condition for a partition of a closed set into minimal hypersurfaces to be a lamination.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \tableofcontents

\section{Introduction}
The space of codimension-$1$ minimal laminations on a Riemannian manifold has been topologized in several different ways.
Thurston \cite[Chapter 8]{thurston1979geometry} introduced both his geometric topology as well as the weak topology of measures on the space of measured geodesic laminations.
Seemingly independently of Thurston, Colding--Minicozzi \cite[Appendix B]{ColdingMinicozziIV} introduced a topology that emphasized not the laminations themselves, but rather the coordinate charts which flatten them.
One goal of this paper is to explain how these three modes of convergence are related, and fill in the details of the compactness theorems available for each mode of convergence.
Our results will consider compactness of sequences of laminations on manifolds that are not necessarily closed and will explicitly rule out the possibility that a sequence of laminations converges to a lamination of lower regularity.

Our second goal is to give a sufficient condition under which a set of disjoint minimal hypersurfaces is actually a minimal lamination.
In dimension $d = 2$, any such set is a lamination \cite[Proposition 7.3]{daskalopoulos2020transverse} but in general one needs uniform bounds on the curvature of the leaves in order to apply the compactness theory and prevent the formation of a singularity in the limit.
In dimensions $d = 3, 4$, if one in addition knows that the leaves are stable, then the curvature bounds are instead provided by the stable Bernstein theorem of Schoen \cite{Schoen2016} and Chodosh--Li \cite{Chodosh2021}.

The application of the stable Bernstein theorem completes the third and final goal of this paper, and the main goal of this series of papers, which also includes the prequel paper \cite{BackusFLG}.
We show that any $1$-harmonic function gives rise to a Ruelle-Sullivan current for a minimal lamination, and conversely.
This generalizes a theorem of Daskalopoulos--Uhlenbeck \cite[Theorem 6.1]{daskalopoulos2020transverse} and resolves \cite[Problem 9.4]{daskalopoulos2020transverse} and \cite[Conjecture 9.5]{daskalopoulos2020transverse}.

%%%%%%%%%%%%%%%%%
\subsection{Minimal laminations}\label{Lams sections}
Let $I \subseteq \RR$ be an interval, and $M$ a Riemannian manifold of dimension $d \geq 2$.
A (codimension-$1$) \dfn{laminar flow box} is a $C^0$ coordinate chart $F: I \times \RR^{d - 1} \to M$ and a compact set $K \subseteq I$ such that each \dfn{leaf} $F(\{k\} \times \RR^{d - 1})$ is $C^2$.
A \dfn{laminar transition map} between two laminar flow boxes $(F_\alpha, K_\alpha), (F_\beta, K_\beta)$ is a $C^0$ map
$$\psi_{\alpha \beta}: I \times \RR^{d - 1} \to I \times \RR^{d - 1}$$
satisfying the usual transition relation
\begin{equation}\label{transition relation}
F_\alpha = F_\beta \circ \psi_{\alpha \beta},
\end{equation}
which maps each leaf $\{k\} \times \RR^{d - 1}$, $k \in K_\alpha$, to a leaf $\{\psi_{\alpha \beta}(k)\} \times \RR^{d - 1}$, so that $\psi_{\alpha \beta}$ is a homeomorphism $K_\alpha \to K_\beta$.
By a \dfn{laminar atlas} we shall mean an atlas for $M$, such that the coordinate charts are all laminar flow boxes and the transition maps are also laminar.

\begin{definition}
A \dfn{lamination} $\lambda$ consists of a nonempty closed set $S \subseteq M$, called its \dfn{support}, and a maximal laminar atlas $\{(F_\alpha, K_\alpha): \alpha \in A\}$ such that in the image $U_\alpha$ of each flow box $F_\alpha$,
$$S \cap U_\alpha = F_\alpha(K \times \RR^{d - 1}).$$
If $\lambda$ is a lamination in the image of a flow box $F$, and $N := F(\{k\} \times \RR^{d - 1})$ is a leaf of $\lambda$, we call $k$ the \dfn{label} of $\lambda$.
\end{definition}

Summarizing the above definitions, a lamination is a nonempty closed set $S$ with a $C^0$ local product structure which realizes it as $K \times \RR^{d - 1}$ for some compact set $K \subset \RR$.

We assume that the leaves are $C^2$ in order to ensure that the normal vectors to each leaves are well-defined in $C^1$, and in particular the second fundamental form and mean curvature of each leaf is well-defined.
Such laminations are sometimes called $C^2$ \dfn{along leaves} \cite{Morgan88}.
This is not the same thing as assuming that the lamination admits a $C^2$ atlas, as it may not be able to extend the normal vectors to each leaf to a $C^1$ vector field on $M$ even locally; see \S\ref{RegularitySec} for more precise assertions about regularity.

\begin{definition}
A lamination $\lambda$ is \dfn{minimal} if its leaves $F_\alpha(\{k\} \times \RR^{d - 1})$ have zero mean curvature, and is \dfn{geodesic} if, in addition, $d = 2$.
\end{definition}

Geodesic laminations are of great interest to the Thurston school of geometric topology \cite[Chapter 8]{thurston1979geometry}.
Later Thurston introduced \dfn{best Lipschitz maps}, namely maps $v: M \to N$ between closed manifolds which minimize their Lipschitz constant $L$ subject to a constraint on their homotopy class.
These maps define a geodesic lamination whose support is the set of points $x$ so that the local Lipschitz constant of $v$ at $x$ is equal to $L$ \cite{thurston1998minimal}. 
If $M, N$ are hyperbolic surfaces of the same genus $g$, then the best Lipschitz constant $L$ is the distance between $M$ and $N$ in \dfn{Thurston's asymmetric metric} on Teichm\"uller space.
This circle of ideas has been developed by the Thurston school \cite{papadopoulos:hal-00129729} but has recently also made contact with geometric PDE through the work of Daskalopoulos--Uhlenbeck \cite{daskalopoulos2020transverse,daskalopoulosPrep1,DaskalopoulosPrep2}, as we recall in \S\ref{FLG section}.

We prove a sufficient condition for a collection of minimal hypersurfaces to define a minimal lamination.

\begin{theorem}\label{building a minimal lamination}
Let $2 \leq d \leq 7$ and suppose that $M$ has constant sectional curvature.
If $\mathcal N$ is a set of disjoint embedded minimal hypersurfaces in $M$, and
\begin{equation}\label{bounding Two}
\sup_{N \in \mathcal N} ||\Two_N||_{C^0} < \infty,
\end{equation}
then $\mathcal N$ is the set of leaves of a minimal lamination.
\end{theorem}

If $d = 2$ or $d \in \{3, 4\}$ and the leaves are stable, then the curvature bound (\ref{bounding Two}) holds.
The stability assumption here is due to the stable Bernstein theorem.

%%%%%%%%%%%%%%%%%%
\subsection{Spaces of minimal laminations}\label{LamSpace section}
In the literature there are at least three different topologies on the space of laminations on $M$.
The first is Thurston's geometric topology \cite[Chapter 8]{thurston1979geometry}, which says that a lamination $\lambda'$ is close to a lamination $\lambda$ if every leaf of $\lambda$ is close to a leaf of $\lambda'$ at least locally, and the same holds for their normal vectors $\normal$.

\begin{definition}
A sequence of laminations $\lambda_i$ converges to a lamination $\lambda$ in \dfn{Thurston's geometric topology} if, for every leaf $N$ of $\lambda$, every $x \in N$, and every $\varepsilon > 0$, there exists $i_\varepsilon \in \NN$ such that for every $i \geq i_\varepsilon$, $\supp \lambda_i$ intersects $B(x, \varepsilon)$, and for $x_i \in B(x, \varepsilon) \cap \supp \lambda_i$,
\begin{equation}\label{convergence of normals}
\dist_{SM}(\normal_{\lambda_i}(x_i), \normal_\lambda(x)) < 2\varepsilon.
\end{equation}
\end{definition}

It is straightforward to show that Thurston's geometric topology does not depend on the choice of Riemannian metric on $M$, or the choice of extension of the distance function on $M$ to its sphere bundle $SM$, which are implicit in the statement thereof.
However, the limiting lamination is not unique, as if $\lambda_i \to \lambda$ and $\lambda'$ is a sublamination of $\lambda$, then $\lambda_i \to \lambda'$.
In particular, Thurston's topology is not Hausdorff, and we say that $\lambda$ is a \dfn{maximal limit} of a sequence $(\lambda_i)$ if $\lambda_i \to \lambda$ and for every $\lambda'$ such that $\lambda_i \to \lambda'$, $\lambda'$ is a sublamination of $\lambda$.

Independently of Thurston, Colding--Minicozzi \cite[Appendix B]{ColdingMinicozziIV} defined a sequence of laminations to converge ``if the corresponding coordinate maps converge;'' that is, if the laminar atlases converge.
This of course says nothing about the limiting set of leaves and in the sequel paper \cite{ColdingMinicozziV} they additionally impose that the sets of leaves converge (say, in Hausdorff distance).
The following is equivalent to requiring that the sets of leaves converge.

\begin{definition}
A sequence $(\lambda_i)$ of laminations \dfn{flow-box converges} in a function space $X$ to $\lambda$ if it converges in Thurston's geometric topology, and there exists a laminar atlas $(F_\alpha)$ for $\lambda$ such that for each $\alpha$, $F_\alpha$ and $(F_\alpha)^{-1}$ are limits in $X$ of flow boxes $F_\alpha^i$, $(F_\alpha^i)^{-1}$ in laminar atlases for $\lambda_i$.
\end{definition}

We now define convergence of laminations equipped with transverse measures.

\begin{definition}
Let $\lambda$ be a lamination with atlas $A$.
A \dfn{transverse measure} to $\lambda$ consists of Radon measures $\mu_\alpha$ with $\supp \mu_\alpha = K_\alpha$, $\alpha \in A$, such that each transition map $\psi_{\alpha \beta}$ is measure-preserving:
$$\mu_\alpha|_{K_\alpha \cap K_\beta} = \psi_{\alpha \beta}^* (\mu_\beta|_{K_\alpha \cap K_\beta}).$$
The pair $(\lambda, \mu)$ is called a \dfn{measured lamination}.
\end{definition}

Caveat lector: we assume that $\supp \mu_\alpha = K_\alpha$, but in \cite{daskalopoulos2020transverse}, it is only assumed that $\supp \mu_\alpha \subseteq K_\alpha$.
In particular, not every lamination admits a transverse measure. 

The definition of transverse measure in terms of Radon measures on $K_\alpha$ is convenient because $K_\alpha$ is compact.
However, the definition is not intrinsic, and this causes problems when considering questions of convergence: the fact that the flow boxes of a convergent sequence of measured laminations converge should be a consequence of, not a part of, the definition!

To rectify this, we first observe that in the definition of a transverse measure, we cannot define a transverse measure to be one on the underlying manifold $M$ itself.
Indeed, Lebesgue measure is ``transverse'' to all foliations; thus such a definition forgets the ``direction'' the measure points in.
However, the notion of Ruelle-Sullivan current allows us to speak of a measure-theoretic object on $M$ which has a well-defined local product structure.

\begin{definition}
A lamination is \dfn{oriented} if one can choose its transition maps to all be orientation-preserving.
\end{definition}

It is clear that a lamination $\lambda$ is locally orientable, since if one replaces $M$ by a small open set, then $\lambda$ has a global flow box.

\begin{definition}
Let $(\lambda, \mu)$ be a measured oriented lamination and let $(\chi_\alpha)_{\alpha \in A}$ be a subordinate partition of unity.
The \dfn{Ruelle-Sullivan current} $T_\mu$ associated to $(\lambda, \mu)$ is defined for all compactly supported $d-1$-forms $\varphi$ by
\begin{equation}\label{RS current}
\int_M T_\mu \wedge \varphi := \sum_{\alpha \in A} \int_{K_\alpha} \left[\int_{\{k\} \times \RR^{d - 1}} (F_\alpha^{-1})^* (\chi_\alpha \varphi) \right] \dif \mu_\alpha(k).
\end{equation}
\end{definition}

It is clear that any lamination is locally orientable, so the next definition makes sense.

\begin{definition}
A sequence of measured laminations $(\lambda_i, \mu_i)$ \dfn{converges} to $(\lambda, \mu)$ if locally, their Ruelle-Sullivan currents $T_{\mu_i} \to T_\mu$ converge in the weak topology of measures.
\end{definition}

The convergence of Ruelle-Sullivan currents, which is very convenient to work with analytically, is equivalent to a definition of measure convergence that may be more familiar to topologists, namely convergence of the transverse measure along each transverse curve.

It is clear from the definitions that flow-box convergence implies Thurston convergence, and it is well-known that measure convergence implies Thurston convergence \cite[Proposition 8.10.3]{thurston1979geometry}.
We show that flow-box convergence actually sits in the middle of the chain of implications:

\begin{theorem}\label{implication theorem}
Suppose that $M$ has constant sectional curvature and $2 \leq d \leq 7$.
Let $(\lambda_i, \mu_i)$ be measured minimal laminations, and $(\lambda_i, \mu_i) \to (\lambda, \mu)$. Then $\lambda_i \to \lambda$ as flow boxes.
\end{theorem}

We also prove some compactness results for the above modes of convergence.
In general a sequence of laminations may escape to infinity or blow up in curvature, and we must rule these possibilities out.

\begin{definition}
A sequence $(\lambda_i)$ of laminations is \dfn{tight} if there exists a compact set $K \subseteq M$ such that for every $i$, $\supp \lambda_i$ intersects $K$.
The sequence has \dfn{bounded curvature} if there exists $C > 0$ such that for any $i$ and any leaf $N$ of $\lambda_i$, the second fundamental form satisfies $||\Two_N||_{C^0} \leq C$.
\end{definition}

We write $C^{1-}$ for the Fr\'echet space $\bigcap_{\alpha < 1} C^\alpha$, where $C^\alpha$ are H\"older spaces.

\begin{theorem}\label{compactness theorem}
Suppose that $M$ has constant sectional curvature and $2 \leq d \leq 7$.
Let $(\lambda_i)$ be a tight sequence of minimal laminations of bounded curvature.
Then a subsequence converges as flow boxes in $C^{1-}$, and in particular in Thurston's topology.
If $\lambda_i$ are measured, then a further subsequence converges as measures.
\end{theorem}



%%%%%%%%%%%%%%%%%%
\subsection{Best Lipschitz and least gradient maps}\label{FLG section}
If $M$ is a closed hyperbolic manifold, then the Euler-Lagrange equation for best Lipschitz maps $v: M \to \Sph^1$ is the $\infty$-Laplace equation \cite{daskalopoulos2020transverse}
\begin{equation}\label{infinity laplacian}
(\nabla^\mu \partial^\nu v) \partial_\mu v \partial_\nu v = 0.
\end{equation}
This equation is invariant under translations $v \mapsto v + y$, so by Noether's theorem, it has a conserved flux $\dif u$.
If $d = 2$, the associated conservation law is the $1$-Laplace equation.
We studied the $1$-Laplacian in the prequel paper \cite{BackusFLG}; here we recall the main result of that paper.

\begin{definition}
A function $u \in BV_\loc(M)$ has \dfn{least gradient}, or is \dfn{$1$-harmonic}, if for every $w \in BV_\cpt(M)$,
\begin{equation}\label{least gradient functional}
\int_M \star |\dif u| \leq \int_M \star |\dif u + \dif w|.
\end{equation}
\end{definition}

Here $\star |\dif u|$ is the total variation of the current $\dif u$; we refer to \S\ref{MeasurePrelims} for the precise definition.
The formal Euler-Lagrange equation for (\ref{least gradient functional}) is the $1$-Laplace equation
\begin{equation}\label{1Laplacian}
\dif^* \left(\frac{\dif u}{|\dif u|}\right) = 0.
\end{equation}
Formally, the $1$-Laplace equation (\ref{1Laplacian}) asserts that the level sets are indeed minimal, but since the derivation of the Euler-Lagrange equation for (\ref{least gradient functional}) is only formal, and the precise definition of weak solution \cite{Mazon14} does not directly imply that the level sets have zero mean curvature, this has to be checked separately.
This is the main result of the prequel paper \cite{BackusFLG}:

\begin{theorem}\label{main thm of old paper}
Let $M$ be a manifold of constant sectional curvature and $2 \leq d \leq 7$.
Then for every $1$-harmonic function $u: M \to \RR$ and $y \in \RR$, the level set $\partial \{u > y\}$ is an analytic embedded stable minimal hypersurface in $M$.
\end{theorem}
\begin{proof}[Proof sketch]
By a straightfoward modification of \cite[Theorem 1]{BOMBIERI1969}, the superlevel sets $\{u > y\}$ have least perimeter, that is their indicator functions have least gradient.
The regularity of boundaries of sets of least perimeter was established for $M = \RR^d$ by the classical work of de Giorgi and Miranda \cite{deGiorgi61, Miranda66} but their proof does not generalize nicely because it relies on the invariance of tangent vectors under parallel transport in order to define averages of normal vectors to sets of least perimeter.
In \cite[\S3]{BackusFLG} we establish a suitable method of taking averages of the normal vector provided that $M = \Hyp^d$ or $M = \Sph^d$.
This allows us to modify the relevant parts of \cite{Miranda66}.
See \cite[\S1]{BackusFLG} for a more detailed summary of \cite{BackusFLG}.
\end{proof}

We use Theorem \ref{main thm of old paper} as a regularity result at various points in this paper, which allows us to show that the limiting laminations have smooth leaves (rather than currents for leaves, say).
However, Theorem \ref{main thm of old paper} is of more interest to this paper rather than just as a regularity lemma; to motivate why, we recall the  relationship between best Lipschitz maps, maps of least gradient, and geodesic laminations:

\begin{theorem}[Daskalopoulos--Uhlenbeck]\label{DU theorem}
Let $M$ be a closed hyperbolic surface and $v: M \to \Sph^1$ an $\infty$-harmonic function with conserved flux $\dif u$ and maximal stretch geodesic lamination $\lambda$.
Then any local primitive $u$ of $\dif u$ is a $1$-harmonic function, and $\dif u$ is Ruelle-Sullivan for $\lambda$.
Moreover, the level sets of $u$ are geodesics in $\lambda$.
\end{theorem}

We refer to the original paper \cite{daskalopoulos2020transverse} for a more precise statement.
Inspired by this theorem, Daskalopoulos--Uhlenbeck conjectured that for any $1$-harmonic function on $\Hyp^2$, $\dif u$ should be Ruelle-Sullivan for some (possibly not maximum-stretch) geodesic lamination \cite[Problem 9.4]{daskalopoulos2020transverse}, and conversely that if $T$ is a Ruelle-Sullivan current for some geodesic lamination, then local primitives of $T$ are $1$-harmonic \cite[Conjecture 9.5]{daskalopoulos2020transverse}.
Of course, if $d \geq 3$, then the level sets will be minimal hypersurfaces rather than geodesics.

Using Theorem \ref{building a minimal lamination}, the stable Bernstein theorem \cite{Schoen2016, Chodosh2021}, and Theorem \ref{main thm of old paper}, we prove the conjectures of Daskalopoulos--Uhlenbeck:

\begin{theorem}\label{main thm}
Let $2 \leq d \leq 4$ and suppose that $M$ has constant sectional curvature.
\begin{enumerate}
\item Let $u$ be a $1$-harmonic function.
Then $\bigcup_{y \in \RR} \partial \{u > y\}$ is the support of a minimal lamination $\lambda$ whose leaves are the connected components of the hypersurfaces $\partial \{u > y\}$, and $\dif u$ is a Ruelle-Sullivan current for $\lambda$.
\item Conversely, if $\lambda$ is a minimal lamination and $\dif u$ is a Ruelle-Sullivan current for $\lambda$, then $u$ is $1$-harmonic.
\end{enumerate}
\end{theorem}

However, Theorem \ref{main thm} leaves a key point -- the r\^ole of the $\infty$-Laplacian -- open, and we have not attempted to address this point here.
The Daskalopoulos--Uhlenbeck theorem was our main motivation for this series of papers; however, our work works in dimension $d = 3$ while Theorem \ref{DU theorem} is purely a statement about $d = 2$.
Indeed, if $d \geq 3$, then the $\infty$-Laplacian gives geodesic laminations, but the $1$-Laplacian gives codimension-$1$ laminations, which are necessarily not geodesic, so there is not an obvious generalization of Theorem \ref{DU theorem} to $d = 3$.

It is tantalizing to think that a suitable system of coupled $\infty$-Laplacians will satisfy a generalized maximum-stretch condition on a codimension-$1$ minimal lamination $\lambda$, and that the conservation law for this conjectural system will be the $1$-Laplacian.
However, even if one was to derive such a system of $\infty$-Laplacians formally, the analysis for studying such a system would likely not be in place.
Indeed, the study of $\infty$-harmonic functions is almost entirely carried out in the language of viscosity solutions and comparison-with-cones, which only make sense when the target is $\RR$ rather than $\RR^m$ or a vector bundle.
Even so, we hope to return to this question in a later work: what is the correct generalization of the Daskalopoulos--Uhlenbeck theorem to $d = 3$?

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Outline of the paper}
In \S\ref{Prelims} we recall preliminaries.

In \S3 we prove the compactness Theorem \ref{compactness theorem} and the convergence implications Theorem \ref{implication theorem} simultaneously.

In \S4 we prove Theorem \ref{building a minimal lamination} establishing which sets of minimal hypersurfaces are minimal laminations, using Theorem \ref{compactness theorem}.
We then use Theorem \ref{building a minimal lamination} to prove Theorem \ref{main thm}.






%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Acknowledgements}
I would like to thank Georgios Daskalopoulos for suggesting this project and for many helpful discussions.
I would also like to thank Chao Li for helpful comments.
This work was supported by an NSF Graduate Resarch Fellowship TODO.

%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Preliminaries}\label{Prelims}
\subsection{Laminations}\label{RegularitySec}
\begin{definition}
Let $\lambda$ be a lamination in $M$.
\begin{enumerate}
\item $\lambda$ is \dfn{discrete} if every leaf space of $\lambda$ is finite.
\item $\lambda$ is \dfn{finite} if $\lambda$ is discrete and admits a finite laminar atlas.
\item $\lambda$ is a \dfn{foliation} if $\supp \lambda = M$.
\end{enumerate}
\end{definition}

Every finite lamination is discrete, and conversely, a discrete lamination in a closed manifold is finite.
However, a lamination with finitely many leaves need not be finite, even if it is geodesic:

\begin{example}\label{two geodesics}
Let $M = \RR \times \Sph^1$ be a cylinder.
We may choose a metric on $M$ so that there is a geodesic $\gamma_1$ which is a simple closed curve looping around $M$, and that there is another geodesic $\gamma_2$ which winds around $M$ infinitely many times, converging to $\gamma_1$.
Then $\gamma_1, \gamma_2$ form a geodesic lamination with two leaves which is not finite.
This example, and slight modifications thereof, will be frequently useful as a counterexample throughout this paper.
\end{example}

Though we impose that laminations are $C^0$ (in the sense that their flow boxes are $C^0$) we can improve this regularity in the case of minimal laminations.

\begin{theorem}\label{regularity theorem}
Let $\lambda$ be a minimal lamination. Then $\lambda$ admits a Lipschitz laminar atlas, and $\normal_\lambda$ is locally Lipschitz.
\end{theorem}
\begin{proof}
The proof is identical to \cite[Theorem 1.1]{Solomon86}, which proves this if $\lambda$ is a foliation.
\end{proof}

\begin{example}
Lipschitz regularity is optimal, even in the nicest possible case of a geodesic foliation of a complete Riemannian manifold.
Indeed, let $M = B((2, 0), 1)$ in $\RR^2$. Then the chords $\{y = ax\}$ for $a > 0$ and $\{y = a\}$ for $a \leq 0$ define a foliation $\lambda$ of $M$.
If we equip $M$ with the Beltrami-Klein metric, then, since the leaves of $\lambda$ are chords, $\lambda$ defines a geodesic foliation of $M \cong \Hyp^2$.
Moreover the conormal to the foliation is
$$\normal = \frac{1_{y > 0}}{\sqrt{1 + y^2/x^2}} \left[\dif y - \frac{y}{x} \dif x\right] + 1_{y \leq 0} \dif y$$
which is not $C^1$.
\end{example}


%%%%%%%%%%%%%%%%
\subsection{Thurston's geometric topology}

\begin{definition}
Let $X$ be a compact metric space. The \dfn{Hausdorff distance} between two closed sets $A, B \subset X$ is
$$\dist(A, B) := \max\left(\max_{a \in A} \min_{b \in B} \dist(a, b), \max_{b \in B} \min_{a \in A} \dist(a, b)\right).$$
The space of closed subsets of $X$ is the \dfn{hyperspace} $\Hypspace X$.
\end{definition}

For compact $X$, however, one can give a homeomorphism-invariant definition of the topology of $\Hypspace X$ \cite[Chapter 4]{nadler2017continuum}, and from that definition it follows that $\Hypspace X$ is a compact metric space.
In particular, $\Hypspace$ is a self-map of the class of compact metrizable spaces.
Moreover, if $K_i \to K$ in $\Hypspace X$, then $K$ is the set of limits of sequences $(k_i)$ such that $k_i \in K_i$.


%%%%%%%%%%%%%%%%%%%%%
\subsection{Measure theory}\label{MeasurePrelims}
Let $X$ be a Polish space, and let $C_\cpt(X)$ be the LF-space of compactly supported continuous functions $f: X \to \RR$.
Its dual $C_\cpt(X)'$ is canonically isomorphic to the space of signed Radon measures on $X$, where the bilinear pairing is given by integration.
The weak topology on $C_\cpt(X)'$ is known as the \dfn{weak topology of measures}.
Unpacking the definitions, a sequence $(\mu_i)$ of Radon measures converges to $\mu$ in the weak topology of measures iff for every compact $Y \subseteq X$ and every continuous function $f: Y \to \RR$,
$$\lim_{i \to \infty} \int_Y f \dif \mu_i = \int_Y f \dif \mu.$$

If $X = M$ is a manifold, then we can consider instead the space $C_\cpt(M, \Omega^\ell)$ of compactly supported continuous $\ell$-forms.
An $\ell$-\dfn{current} with locally finite total variation is an element of the dual space $C_\cpt(M, \Omega^\ell)'$.
(We shall never need to consider $\ell$-currents whose total variation is not locally finite.)
We denote the pairing of an $\ell$-current $T$ and an $\ell$-form $\varphi$ by $\int_M T \wedge \varphi$.
Again we have the weak topology on the space of $\ell$-currents; we also have the \dfn{derivative}
$$\int_M \dif T \wedge \psi := -\int_M T \wedge \dif \psi$$
defined for any $\ell$-current $T$ such that $\psi \mapsto \int_M T \wedge \dif \psi$ extends from $C^1_\cpt(M, \Omega^{\ell - 1})$ to a continuous linear function on $C_\cpt(M, \Omega^{\ell - 1})$.

With this machinery in place, we can talk about Ruelle-Sullivan currents and their limits.

\begin{lemma}
Let $(\lambda, \mu)$ be a measured oriented lamination.
Then the Ruelle-Sullivan current $T_\mu$ is well-defined; it is honestly a $d-1$-current, and does not depend on the choice of partition of unity.
Moreover, $\dif T_\mu = 0$.
\end{lemma}
\begin{proof}
We first claim that the right-hand side of (\ref{RS current}) is always finite, and is continuous in $\varphi$.
In fact, possibly after refining $(\chi_\alpha)$, we may assume that it is a locally finite partition of unity.
In particular, we just need to check the continuity in a single flow box:
$$\left|\int_{K_\alpha} \left[\int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^* (\chi_\alpha \varphi) \right] \dif \mu_\alpha(k)\right| \leq \int_{K_\alpha} \int_{\RR^{d - 1} \times \{k\}} |(F_\alpha^{-1})^* (\chi_\alpha \varphi)| \dif \mu_\alpha(k).$$
The inner integral is controlled by $||\varphi||_{C^0(U_\alpha)} \cdot |U_\alpha|$ where $U_\alpha$ is the image of $F_\alpha$.
The outer integral is then well-defined because it is against a Radon measure.

We next observe that the choice of partition of unity is irrelevant, thus if $\varphi$ has compact support in $U_\alpha \cap U_\beta$, then
\begin{equation}\label{well-defined of Ruelle-Sullivan}
\int_{K_\alpha} \int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^* \varphi \dif \mu_\alpha(k) = \int_{K_\beta} \int_{\RR^{d - 1} \times \{k\}} (F_\beta^{-1})^* \varphi \dif \mu_\beta(k).
\end{equation}
Indeed,
\begin{align*}
\int_{K_\alpha} \int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^* \varphi \dif \mu_\alpha(k)
&= \int_{K_\beta} (F_\alpha F_\beta^{-1})^* \left[\int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^* \varphi \dif \mu_\alpha(k)\right] \\
&= \int_{K_\beta} \left[\int_{\RR^{d - 1} \times \{k\}} (F_\beta^{-1})^* \varphi\right] (F_\alpha F_\beta^{-1})^* \dif \mu_\beta(k) \\
&= \int_{K_\beta} \int_{\RR^{d - 1} \times \{k\}} (F_\beta^{-1})^* \varphi \dif \mu_\beta(k)
\end{align*}
where the last equation is because of the measure-preserving nature of the transition maps; this proves (\ref{well-defined of Ruelle-Sullivan}).

Finally, if a $d-2$-form $\psi$ has compact support in a single flow box, then
$$\int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^* \dif \psi = \int_{\RR^{d - 1} \times \{k\}} \dif((F_\alpha^{-1})^* \psi) = 0$$
by Stokes' theorem, so
\begin{align*}
\int_M \dif T_\mu \wedge \psi &= -\int_M T_\mu \wedge \dif \psi \\
&= -\int_{K_\alpha} \int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^* \dif \psi \dif \mu_\alpha(k) = 0. \qedhere
\end{align*}
\end{proof}

Though (\ref{RS current}) is the more traditional way of stating the definition of a Ruelle-Sullivan current, there is a more intrinsic way as well.
We first observe that if $\mu$ is a transverse measure, then $\mu$ defines a measure on $\supp \lambda$: in each flow box $F_\alpha$, an open set $U$ has measure
\begin{equation}\label{transverse measure of an open set}
\mu(U) := \int_{K_\alpha} |F_\alpha(\RR^{d - 1} \times \{k\}) \cap U| \dif \mu_\alpha(k).
\end{equation}

\begin{lemma}
For an oriented measured lamination $(\lambda, \mu)$, the polar decomposition of $T_\mu$ is
\begin{equation}\label{polar ruelle sullivan}
T_\mu = \normal_\lambda \mu.
\end{equation}
\end{lemma}
\begin{proof}
For an open set $U \subseteq M$ in a flow box $F_\alpha$, the total variation measure $|T_\mu|$ satisfies
$$|T_\mu|(U) = \sup_{||\varphi||_{C^0} \leq 1} \int_{K_\alpha} \int_{\RR^{d - 1} \times \{k\}} \varphi \dif \mu_\alpha(k)$$
where the supremum ranges over $d-1$-forms $\varphi$ with compact support in $U$.
However, $\star \normal_\lambda$ is the Riemannian measure on $F_\alpha(\RR^{d - 1} \times \{k\})$, so
$$\int_{\RR^{d - 1} \times \{k\}} \varphi \leq \int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^*(\star \normal_\lambda).$$
Since $||\normal^\lambda||_{C^0} = 1$, it follows that a sequence of cutoffs of $\star \normal_\lambda$ to more and more of $U$ is a maximizing sequence.
Therefore $\normal_\lambda$ is the polar part of (\ref{polar ruelle sullivan}), and
$$|T_\mu|(U) = \int_{K_\alpha} \int_{\RR^{d - 1} \times \{k\}} (F_\alpha^{-1})^*(1_U \star \normal_\lambda) \dif \mu_\alpha(k).$$
The inner integral is the Riemannian measure of $F_\alpha(\RR^{d - 1} \times \{k\}) \cap U$, so by (\ref{transverse measure of an open set}), $|T_\mu| = \mu$.
\end{proof}

% \begin{proposition}
% If $\lambda$ is a minimal lamination which admits a transverse measure, then every leaf of $\lambda$ is closed.
% \end{proposition}
% \begin{proof}
% If $N$ is a leaf of $\lambda$ which is not closed, but $\mu$ is a transverse measure, then since $\supp \lambda$ is closed, we can find a sequence $(x_i)$ which converges to a point $x \in \supp \lambda \setminus N$.
% Considering flow box coordinates $F_\alpha$ at $x$, we see that there exist infinitely many different labels $(k_i)$ of $N$.
% Indeed, if it is possible to express $N$ locally near $x$ as $J \times \RR^{d - 1}$ where $J \subset I$ is finite, then $N$ would be locally closed in a neighborhood of $x$, a contradiction.

% In the image $U_\alpha$ of $F_\alpha$, $\lambda$ is orientable and so we can write $T_\mu = \dif u$ for some function $u$ of least gradient.
% In particular, $N \cap U_\alpha$ is a level set of $u$ and hence bounds a set of least perimeter, which then must have infinitely many connected components in any neighborhood of $x$.
% However, sets of least perimeter only have locally finitely many connected components, so this case is still a contradiction.
% \end{proof}

% In particular, the pathological lamination of Example \ref{two geodesics} does not admit a transverse measure.

Here is another characterization of the weak topology of measures in terms of transverse curves.

\begin{definition}
For a flow box $F: I \times \RR^{d - 1} \to M$, we define its \dfn{labelling map}
$$\Pi_F: F(I \cap \RR^{d - 1}) \to I$$
to be the projection of $F^{-1}$ onto $I$:
% https://q.uiver.app/?q=WzAsMyxbMCwwLCJGKEkgXFxjYXAgXFxtYXRoYmIgUl57ZCAtIDF9KSJdLFsyLDAsIkkgXFxjYXAgXFxtYXRoYmIgUl57ZCAtIDF9Il0sWzIsMiwiSSJdLFswLDEsIkZeey0xfSJdLFsxLDIsIiIsMCx7InN0eWxlIjp7ImhlYWQiOnsibmFtZSI6ImVwaSJ9fX1dLFswLDIsIlxcUGlfRiIsMl1d
\[\begin{tikzcd}
	{F(I \cap \mathbb R^{d - 1})} && {I \cap \mathbb R^{d - 1}} \\
	\\
	&& I
	\arrow["{F^{-1}}", from=1-1, to=1-3]
	\arrow[two heads, from=1-3, to=3-3]
	\arrow["{\Pi_F}"', from=1-1, to=3-3]
\end{tikzcd}\]
\end{definition}

\begin{definition}
For a smooth path $\gamma: I \to M$ and a lamination $\lambda$, we say that $\gamma$ is \dfn{transverse} to $\lambda$ if, for every $t \in I$ such that $\gamma(t) \in \supp \lambda$, there exists a flow box $F$ in a neighborhood of $\gamma(t)$ and a neighborhood $J \subseteq I$ of $t$ such that
$$\Pi_F \circ \gamma|_J: J \to I$$
is a topological embedding, i.e. a homeomorphism onto its image.
If $\Pi_F \circ \gamma|_J$ is in addition increasing, we say that $\gamma$ is \dfn{positively transverse}.
% By a \dfn{transverse homotopy} of transverse curves to $\lambda$ we mean a homotopy of curves $H: I^2 \to M$ such that every curve $\gamma_s := H(s, \cdot)$ is transverse to $\lambda$, and if $H(s, t)$ is an element of some leaf $N$ of $\lambda$, then so is $H(s', t)$ for any $s'$.
\end{definition}

\begin{definition}
For a transverse curve $\gamma$ to a measured lamination $(\lambda, \mu)$, we get a Radon measure $\gamma^! \mu$ on $I$, as follows.
Near $t \in I$, choose a flow box $F$ near $\gamma(t)$ and a neighborhood $J$ of $t$, such that $\Pi_F \circ \gamma|_J$ is a topological embedding.
Then
$$(\gamma^! \mu)|_J := ((\Pi_F \circ \gamma|_J)^{-1})_* \mu_F.$$
\end{definition}

The measure $\gamma^! \mu$ is well-defined, essentially since the transition maps are measure-preserving (so the only thing that actually matters is the weight assigned to each leaf that $\gamma$ passes through).

\begin{lemma}
Let $(\lambda, \mu)$ be a measured oriented lamination, $\gamma$ a curve, and $\gamma(t) \in \supp \lambda$.
Then $\gamma$ is positively transverse to $\lambda$ at $\gamma(t)$ iff
\begin{equation}\label{transverse means rs}
\lim_{\varepsilon \to 0} \frac{1}{\mu(B(\gamma(t), \varepsilon))} \int_{B(\gamma(t), \varepsilon)} T_\mu \wedge \star (\gamma')^\flat > 0.
\end{equation}
\end{lemma}
\begin{proof}
We work in flow box coordinates $(k, y)$, so that the labelling map $\Pi$ is $\Pi(k, y) = k$.
In such coordinates, leaves $N$ take the form $\{k = k_N\}$ for some $k_N$, so $\dif k$ is conormal to each leaf.
In particular, $\normal_\lambda = (g^{kk})^{-1/2} \dif k$.
Also $(\Pi \circ \gamma)' = (\dif k, \gamma')$, so (since $(g^{kk})^{-1/2}$ is positive), $\gamma$ is positively transverse at $\gamma(t)$ iff
\begin{equation}\label{transverse means dk}
(\normal_\lambda, \gamma')(\gamma(t)) > 0.
\end{equation}
But $\gamma(t) \in \supp \mu$, and $\mu$ is a Radon measure.
So by the Lebesgue differentiation theorem, (\ref{transverse means dk}) happens iff
$$\lim_{\varepsilon \to 0} \frac{1}{\mu(B(\gamma(t), \varepsilon))} \int_{B(\gamma(t), \varepsilon)} (\normal_\lambda, \gamma') \dif \mu > 0.$$
Applying the polar decomposition (\ref{polar ruelle sullivan}) of $T_\mu$ we conclude that this last condition is equivalent to (\ref{transverse means rs}).
\end{proof}

\begin{proposition}
Let $(\lambda, \mu)$ be a measured oriented lamination and $\gamma$ a transverse curve to $\lambda$.
Then there exists an open set $U \supseteq \supp \lambda$ in $M$ and an open set $\mathcal U \ni (\lambda, \mu)$ in the space of measured oriented laminations, such that if $(\kappa, \nu) \in \mathcal U$ has support in $U$, then $\gamma$ is transverse to $\kappa$.
\end{proposition}
\begin{proof}
Suppose not. Then there exist laminations $\lambda_i$ with
$$\supp \lambda_i \subseteq \{x \in M: \dist(x, \supp \lambda) < i^{-1}\},$$
and transverse measures $\mu_i$ to $\lambda_i$, such that $(\lambda_i, \mu_i) \to (\lambda, \mu)$ in the weak topology of measures, but with $\gamma$ not transverse to all the $\lambda_i$, say at $\gamma(t_i) \in \supp \lambda_i$.
By taking a subsequence, we may assume that $\gamma(t_i) \to \gamma(t)$ for some $t \in I$.
Since $\dist(\gamma(t_i), \supp \lambda) < i^{-1}$, $\gamma(t) \in \supp \lambda$.
Since $\gamma$ is transverse to $\lambda$, there exist $\varepsilon_*, \delta > 0$ such that for each $0 < \varepsilon \leq \varepsilon_*$,
$$\frac{1}{\mu(B(\gamma(t), \varepsilon))} \int_{B(\gamma(t), \varepsilon)} T_\mu \wedge \psi \geq \delta.$$
From the definition of the weak topology of measures,
$$\lim_{i \to \infty} \frac{1}{\mu(B(\gamma(t_i), \varepsilon))} \int_{B(\gamma(t_i), \varepsilon)} T_{\mu_i} \wedge \psi \geq \delta.$$
By the portmanteau theorem, this implies
$$\liminf_{i \to \infty} \frac{1}{\mu_i(B(\gamma(t_i), \varepsilon))} \int_{B(\gamma(t_i), \varepsilon)} T_{\mu_i} \wedge \psi \geq \delta.$$
In particular, for every $i$ large enough,
$$\frac{1}{\mu_i(B(\gamma(t_i), \varepsilon))} \int_{B(\gamma(t_i), \varepsilon)} T_{\mu_i} \wedge \psi > 0,$$
which contradicts the fact that $\gamma$ is not transverse to $\lambda_i$ at $t_i$.
\end{proof}

\begin{example}
There exist measured oriented laminations $(\lambda, \mu)$ and $(\kappa, \nu)$ which are arbitrarily close in the weak topology of measures, such that $\lambda$ admits a transverse curve $\gamma$ which is not transverse to $\kappa$.
Working in $\RR^2$, let $\lambda$ consist of the horizontal axis, $\gamma$ the vertical axis, and $\kappa$ the horizontal axis plus a curve $\rho$ in the upper half-plane which meets $\gamma$ tangentially.
Then take $\mu$ to be the usual measure on $\lambda$, $\nu$ to be $\mu$ plus a small measure on $\rho$.
We can take $\nu$ to be arbitrarily close to $\mu$ while preserving the fact that $\supp \nu = \supp \kappa$.
\end{example}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weak topology of measures versus Thurston's geometric topology}
We now show that for a geodesic lamination, convergence in the weak topology of measures implies convergence in Thurston's geometric topology.
Thurston claimed this fact \cite[Proposition 8.10.3]{thurston1979geometry} but his proof left something to be desired as it did not justify why the limit is geodesic, or why the convergence respects the normal vectors.

\begin{lemma}\label{measured limits are almost thurston}
Let $(\lambda_i, \mu_i)$ be measured geodesic laminations.
If $(\lambda_i, \mu_i) \to (\lambda, \mu)$ in the weak topology of measures, then for every $x \in \supp \lambda$ there exists $x_i \in \supp \lambda_i$ such that $x_i \to x$.
\end{lemma}
\begin{proof}
For every $\varepsilon > 0$, $\mu(B(x, \varepsilon)) > 0$, so by the portmanteau theorem,
$$\liminf_{i \to \infty} \mu_i(B(x, 2\varepsilon)) \geq \mu(B(x, \varepsilon)).$$
Therefore there exists $x_i \in \supp \lambda_i \cap B(x, 2\varepsilon)$ if $i$ is large enough.
\end{proof}

\begin{lemma}\label{weak convergence of FLG}
Let $(u_i)$ be a sequence of functions of least gradient which is bounded in $L^1$, and $T$ a current such that $\dif u_i \to T$ weakly.
Then there exists a function $u$ of least gradient such that along a subsequence, $u_i \to u$ weakly in $BV$, and $\dif u = T$.
\end{lemma}
\begin{proof}
By the Miranda stability theorem, $(u_i)$ has a subsequential limit $u$ in $L^1$, which has least gradient.
For any compactly supported $d-1$-form $\varphi$,
\begin{align*}
\int_M T \wedge \varphi
&= \lim_{i \to \infty} \int_M \dif u_i \wedge \varphi
= -\lim_{i \to \infty} \int_M u_i \dif \varphi = - \int_M u \dif \varphi.
\end{align*}
where the last equality is because the Radon-Nikod\'ym derivative of $\dif \varphi$ with respect to Lebesgue measure is in $L^\infty$ and hence $L^1(\dif \varphi)$ is a weaker space than $L^1$. The last expression here is $\int_M \dif u \wedge \varphi$.
So $T = \dif u$ and $\dif u_i \to \dif u$ in the weak topology of measures; that is, $u_i \to u$ weakly in $BV$.
\end{proof}

\begin{lemma}\label{limits of measured geodesic lams are geodesic}
Let $(\lambda_i, \mu_i)$ be measured geodesic laminations.
If $(\lambda_i, \mu_i) \to (\lambda, \mu)$ in the weak topology of measures, then $\lambda$ is geodesic.
\end{lemma}
\begin{proof}
Let $x \in \supp \lambda$ and $r > 0$ such that $B := B(x, r)$ is contractible.
In $B$, we can write $T_{\mu_i} = \dif u_i$ for some sequence of functions of least gradient $u_i \in BV(B)$.
Since $u_i$ is only defined up to a constant, we impose $\int_M \star u_i = 0$, so by Poincar\'e's inequality,
$$||u_i||_{L^1(B)} \lesssim r\int_B \star |T_{\mu_i}| \lesssim 1$$
where the last bound is because $(T_{\mu_i})$ is compact in the weak topology of measures.
So by Lemma \ref{weak convergence of FLG}, $u_i \to u$ where $u$ has least gradient and satisfies $T_\mu = \dif u$.
So by \cite{BackusFLG}, the leaf $N$ containing $x$ is a geodesic and $x$ is not an endpoint of $N$.
But $x$ was arbitrary, so $N$ is in fact complete.
\end{proof}

\begin{proposition}\label{measured implies Thurston}
Let $(\lambda_i, \mu_i)$ be measured geodesic laminations.
If $(\lambda_i, \mu_i) \to (\lambda, \mu)$ in the weak topology of measures, then $\lambda$ is a geodesic lamination and $\lambda_i \to \lambda$ in Thurston's geometric topology.
\end{proposition}
\begin{proof}
By Lemmata \ref{measured limits are almost thurston} and \ref{convergence of geodesic lams in thurston},
$\lambda$ is a geodesic lamination and for every $x \in \supp \lambda$, $\varepsilon > 0$, and large $i$, there exists $y \in \supp \lambda_i \cap B(x, \varepsilon)$.
To get convergence of the normals we choose a Lipschitz $d-1$-form $\varphi$ which extends $\star \normal$.
Then $\int_K T_\mu \wedge \varphi = \mu(K)$ for any $\mu$-continuity set $K$, so by the portmanteau theorem,
$$\lim_{i \to \infty} \frac{\int_K T_{\mu_i} \wedge \varphi}{\mu_i(K)} = \frac{\int_K T_\mu \wedge \varphi}{\mu(K)} = 1.$$
On the other hand, if we assume that there exists $\delta, \varepsilon > 0$ such that for every $y \in \supp \lambda_i \cap B(x, \varepsilon)$,
$$|\sin(\normal_i(y) - \normal(x))| \geq \delta,$$
then possibly after shrinking $\varepsilon$ we may assume that $B(x, \varepsilon)$ is a $\mu$-continuity set, so that
$$\int_K T_{\mu_i} \wedge \varphi = \int_K \normal_i\mu_i \wedge \star \normal \leq (1 - O(\delta)) \mu_i(K)$$
which contradicts $\delta > 0$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%

\section{Convergence in flow boxes}
Let $C^{1-}$ be the Fr\'echet space $\bigcap_{\alpha < 1} C^\alpha$, where $C^\alpha$ are H\"older spaces.
The following definition is one possible interpretation of the vague definition of \cite[Lemma II.1.2]{ColdingMinicozziV}.



We allow the limiting flow box to not be Lipschitz, but just $C^{1-}$, if necessary.
Just as for Thurston's geometric topology, convergence on the level of flow boxes does not satisfy uniqueness of limits, and hence we speak of \dfn{maximal limits} to allow us to speak of a unique limit of a sequence.

\begin{proposition}
Let $(\lambda, \mu)$ be a measured geodesic lamination.
Then there exist finite sublaminations $\lambda_i$ of $\lambda$ such that $\lambda$ is a maxima limit of $(\lambda_i)$ on the level of flow boxes.
\end{proposition}
\begin{proof}
Let $(x_i)$ be a dense sequence in $\supp \lambda$ and $N_i$ the leaf containing $x_i$.
Since $\lambda$ admits a transverse measure, $N_i$ is closed, so setting $\lambda_i := \bigcup_{j < i} N_j$, $\lambda_i$ is a finite lamination.
Now if $N$ is a leaf in $\lambda$, either $N = N_i$ for some $i$, in which case clearly $N$ is a limiting leaf of $\lambda_i$, or for every $x \in N$ there exists a subsequence $(x_{i_j})$ such that $x_{i_j} \to x$.
In partiular, $N_{i_j}$ approximates $N$ in Thurston's geometric topology (here the convergence of normal vectors is because it's a geodesic lamination).
Moreover, if $F$ is a flow box for $\lambda$ near $x$, then $F$ is also a flow box for $N_i$ for every $i$, hence the convergence in flow boxes.
\end{proof}


%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Compactness}
\begin{definition}
A sequence of laminations $(\lambda_i)$ is \dfn{tight} if there exists a compact set $K \subseteq M$ such that for every $i$ there exists a leaf $N_i$ of $\lambda_i$, such that $N_i \cap K$ is nonempty.
\end{definition}

\begin{lemma}\label{limit of minimals is minimal}
Let $(N_i)$ be a sequence of complete embedded minimal hypersurfaces of uniformly bounded curvature and $N_i \to N$ in $\Hypspace M$.
Then $N$ is a complete minimal hypersurface.
\end{lemma}
\begin{proof}
In a neighborhood of any point on $N$, we can write $N_i = \partial U_i$ where $U_i$ is a set of least perimeter.
Let $u_i := 1_{U_i}$; then $u_i$ is a function of least gradient.

If we set $N = \partial U$, $u := 1_U$, then $u_i \to u$ we claim that pointwise away from $N$, and hence almost everywhere.
This can be seen by writing $N_i$ as the graph of $f_i$, so $U_i = \{y < f_i(x)\}$, and this is possible due to the assumption of bounded curvature.
Similarly we write $U = \{y < f(x)\}$.
Now if $(x, y) \in U$ but $(x, y) \notin U_i$ for arbitrarily large $i$, then $f_i(x) \leq y < f(x)$ so $(x, f_i(x)) \in N_i$ converges to a point $(x, \tilde y)$ with $\tilde y \neq y$, thus $(x, \tilde y) \notin N$.
This violates the convergence $N_i \to N$ in $\Hypspace N$.

So $u_i \to u$ almost everywhere, and hence by dominated convergence also converges in $L^1$.
By the Miranda stability theorem it follows that $U$ has least perimeter, and so by \cite{BackusFLG}, $N$ is a complete minimal hypersurface.
\end{proof}

\begin{lemma}\label{convergence of geodesic lams in thurston}
Assume that $(\lambda_i)$ is a sequence of geodesic laminations, and $\lambda$ a geodesic lamination, such that for every leaf $N$ of $\lambda$, there is a leaf $N_i$ of $\lambda_i$ such that $N_i \to N$ pointwise.
Then $\lambda_i \to \lambda$ in Thurston's geometric topology.
\end{lemma}
\begin{proof}
We must show that for $x \in N$ and $x_i \in N_i$, $N_i$ a leaf of $\lambda_i$ such that $x_i \to x$, $\normal_i := \normal_{N_i}(x_i)$ converges to $\normal := \normal_N(x)$ in the cosphere bundle $S'M$.
If this fails, then we work in normal coordinates based at $x$, so we can think of points of $M$ as vectors in $\RR^d$, and points of $S'M$ as unit vectors in $\RR^d$.
In such coordinates we may view $N$ as the first axis.
Possibly after taking a subsequence of $(\lambda_i)$, we may assume that no matter what sequence $(x_i)$ we choose,
$$|\sin \angle(\normal_i, \normal)| \geq \varepsilon.$$
The geodesic curvature of $N_i$ with respect to the euclidean metric on the tangent space is bounded independently of $i$ in terms of the scalar curvature $R$ of $M$ near $x$, so there exists $r = r(R) > 0$ independent of $i$ such that $N_i$ \dfn{avoids cones} in the sense that for every $v \in \RR^d$ such that $0 < |v| < r$ and $|\sin v| < \varepsilon/2$, $v + x_i$ does not lie in $N_i$.
After shrinking $\varepsilon$, we may assume that $\varepsilon < \max(r/2, 1/100)$.

To obtain a contradiction, we choose $y$ to lie on $N$ and satisfy $\dist(x, y) = \varepsilon$.
For $i$ large, $|x_i| < \varepsilon^2/10$, so
$$0 < \varepsilon - \varepsilon^2 \leq |y - x_i| < 2\varepsilon = r$$
and hence (using a superscript $j$ to indicate the $j$th coordinate)
$$|y^1 - x_i^1| \geq |y - x_i| - |x_i^2| \geq \varepsilon - 2\varepsilon^2.$$
Consider the triangle $\Delta$ whose vertices are $x_i, y$, and $z_i := (y^1, x_i^2)$.
Then $\Delta$ is a right triangle, and its smallest angle $\theta_\Delta$ satisfies
$$|\sin \theta_\Delta| = \frac{|x_i^2|}{|y^1 - x_i^1|} < \frac{\varepsilon^2}{\varepsilon - 2\varepsilon^2} < \frac{\varepsilon}{4}.$$
Thus for every large $i$, $y$ is contained in the cone avoided by $N_i$, and in fact for any $y_i \in N_i$, $\dist(y, y_i) \geq \varepsilon/4$.
But we argued above that any point of $N$ could be approximated by points of $N_i$ for $i$ large, so this is a contradiction.
\end{proof}

\begin{theorem}\label{compactness in flow boxes and Thurston}
Let $(\lambda_i)$ be a tight sequence of finite minimal laminations with bounded curvature. Then:
\begin{enumerate}
\item After passing to a subsequence, we may assume that $(\lambda_i)$ converges to a maximal limit $\lambda$ on the level of flow boxes, and hence in Thurston's geometric topology.
\item The limit $\lambda$ is a Lipschitz minimal lamination satisfying the same curvature bounds as $(\lambda_i)$.
\item If $\mu_i$ is a transverse measure to $\lambda_i$ and for every compact $K \subseteq M$, $\mu_i(K) \lesssim_K 1$, then after passing to a further subsequence, we may assume that $(\lambda_i, \mu_i)$ in fact converges to some measured minimal lamination $(\tilde \lambda, \mu)$ in the weak topology of measures, where $\tilde \lambda$ is a sublamination of the maximal lamination $\lambda$.
\end{enumerate}
\end{theorem}
\begin{proof}
We first select $x_0 \in M$; we will construct suitable flow boxes $F_i$ for $\lambda_i$ with image $B(x_0, r)$.
Here $r$ is to be determined, but at least is smaller than $R(x_0)/2$, where $R(x_0)$ is the injectivity radius of $M$ at $x_0$.

Let $\varepsilon > 0$.
After selecting $r$ small depending on $g$, $\varepsilon$, and the curvature bounds on $\lambda_i$, and rescaling $g$, we may assume that the leaves $N_{i1}, \dots, N_{ik_i}$ have all second fundamental forms of size $\leq \varepsilon$ on $B(x_0, r)$.
Then if $r$ is small enough depending on $g$, the second fundamental forms $\Two_{ij}$ of $N_{ij}$ with respect to the euclidean metric in $x_0$-normal coordinates are of size $\leq 2\varepsilon$ on $B(x_0, r)$.

If $r$ is small enough, then because there is a Lipschitz conormal $1$-form $\normal_i$ to $\lambda_i$ in $B(x_0, r)$, we can represent $N_{i1}, \dots, N_{ik_i}$ as graphs over the equatorial hyperplane in $B(x_0, r)$, possibly after applying a rotation $R_i$.
The uniform bounds on $\Two_{ij}$ ensure that $r$ is independent of $i$.
Thus we have $x_0$-normal coordinates $(x_i, y_i)$ on $B(x_0, r)$ where
$$N_{ij} = \{y_i = f_{ij}(x_i)\}$$
and $\normal_i(0, 0) = \dif y_i$.
Since $|\Two_{ij}| \leq 2\varepsilon$, if $\varepsilon \leq 1$,
\begin{equation}\label{bound on derivatives}
||\dif f_{ij}||_{C^0} \leq |\normal_i(0, f_{ij}(0))| + O(\varepsilon r) \lesssim (1 + \varepsilon) r \lesssim r.
\end{equation}

By the maximum principle, if $f_{ij}(x_i) = f_{ij'}(x_i)$ for some $i, j, j'$, then $j = j'$.
So after applying a permutation to $\{1, \dots, k_i\}$ we may assume that $j < j'$ implies $f_{ij} < f_{ij'}$.
Then if we set $u_{ij} := f_{i,j+1} - f_{ij}$, then $u_{ij} > 0$ by the maximum principle, and we get a second-order divergence-form operator $P_{ij}(\varepsilon, r)$ such that
\begin{equation}\label{elliptic PDE}
-\Delta u_{ij} = P_{ij}(\varepsilon, r) u_{ij}
\end{equation}
and $P_{ij}(\varepsilon, r)$ is perturbative as $\varepsilon, r \to 0$ in the sense that its coefficients can be made arbitrarily small in $C^0(B_{r/2})$ by taking $\varepsilon, r$ small.
In fact, by \cite[Chapter 7]{colding2011course}, if we set
$$F(x, y, p, q) := h^{ij} (q_{ij} + \Gamma_{ij}^0 + p_i \Gamma_{0j}^0 + p_j \Gamma^0_{i0} + p_i p_j \Gamma^0_{00}) - p_m h^{ij} (\Gamma_{ij}^m + p_i \Gamma_{0j}^m + p_i \Gamma_{i0}^m + q_{ij} \Gamma_{00}^m),$$
where $\Gamma_{\mu \nu}^\lambda$ are Christoffel symbols for $M$ at $(x, y)$, $x_0 := (0, 0)$, the zero index corresponds to $y$, and $h^{ij}$ are the components of the inverse of
$$h_{ij} = g_{ij} + p_i g_{j0} + p_j g_{0i} + p_i p_j g_{00}$$
evaluted at $(x, y)$, then $h_{ij} = \delta_{ij} + O(|x|^2 + y^2 + |p|^2)$, $\Gamma_{\mu \nu}^\lambda = O(|x|^2 + y^2)$, and
$$F(x, f_{i, j + 1}(x), \dif f_{i, j + 1}(x), \partial^2 f_{i, j + 1}(x)) - F(x, f_{ij}(x), \dif f_{ij}(x), \partial^2 f_{ij}(x)) = 0.$$
By (\ref{bound on derivatives}), we conclude that the dominant term is $\Delta u_{ij}$ and the perturbative terms are $O(r + \varepsilon)$ in $C^0$, as required by (\ref{elliptic PDE}).
TODO: Reindex all this, don't want the ij's to be Einstein indices and leaf indices
It follows that if $\varepsilon, r$ are small, then $-\Delta - P$ is uniformly elliptic, with ellipticity constant close to that of $-\Delta$.
So by the Harnack inequality, for any $0 < \delta < 1/2$,
\begin{equation}\label{Harnack bound}
\sup_{B_{\delta r}} u_{ij} \leq e^{O(\delta)} \inf_{B_{\delta r}} u_{ij}.
\end{equation}

We now set $\eta_{ij} := f_{ij}(0)$ and define
$$\varphi_i(\xi_i, \eta_i) := \sum_{j=1}^{k_j - 1} 1_{[\eta_{ij}, \eta_{i,j+1})}(\eta_i) \left[f_{ij}(\xi_i) + \frac{\eta_i - \eta_{ij}}{\eta_{i,j+1} - \eta_{ij}} u_{ij}(x_i)\right].$$
That is, the change of coordinates
$$(x_i, y_i) = (\xi_i, \varphi_i(\xi_i, \eta_i))$$
flattens out the graphs of the $f_{ij}$ to hyperplanes.
Moreover
\begin{align*}
\dif \varphi_i &= \sum_{j=1}^{k_j - 1} 1_{[\eta_{ij}, \eta_{i, j + 1})}(\eta_i) \left[\dif f_{ij} + \frac{\eta_i - \eta_{ij}}{\eta_{i,j+1} - \eta_{ij}} \dif u_{ij} + \frac{u_{ij}}{\eta_{i,j+1} - \eta_{ij}} \dif \eta_i\right] \\
&\qquad + \sum_{j = 2}^{k_j - 1} \delta_{\eta_{ij}}(\eta_i)\left[f_{i,j-1}(\xi_i) + u_{i,j-1}(\xi_i) - f_{i,j}(\xi_i)\right] \\
&= \sum_{j=1}^{k_j - 1} 1_{[\eta_{ij}, \eta_{i, j + 1})}(\eta_i) \left[\dif f_{ij} + \frac{\eta_i - \eta_{ij}}{\eta_{i,j+1} - \eta_{ij}} \dif u_{ij} + \frac{u_{ij}}{\eta_{i,j+1} - \eta_{ij}} \dif \eta_i\right].
\end{align*}
Now $\eta_i - \eta_{ij} \in [0, \eta_{i, j + 1} - \eta_{ij}]$ for $\eta_i \in [\eta_{ij}, \eta_{i,j+1})$, so the first two terms are bounded by (\ref{bound on derivatives}) as
$$\left|\left|\dif f_{ij} + \frac{\eta_i - \eta_{ij}}{\eta_{i,j+1} - \eta_{ij}} \dif u_{ij}\right|\right|_{C^0(B_{\delta r})}
\leq 3\sup_{j' \in \{1, \dots, k_i\}} ||\dif f_{ij'}||_{C^0(B_{\delta r})}
\lesssim \delta r.$$
From (\ref{Harnack bound}),
$$\frac{u_{ij}}{\eta_{i,j+1} - \eta_{ij}} = \frac{u_{ij}(0) + \sup_{B_{\delta r}} u_{ij} - \inf_{B_{\delta r}} u_{ij}}{u_{ij}(0)} = 1 + O(\delta r).$$
In conclusion
$$\dif \varphi_i = \dif \eta_i + O(\delta r).$$
Therefore the flow boxes $F_i(\xi_i, \eta_i) := (x_i, y_i)$ have Lipschitz norm and conorm in $(1/2, 3/2)$ if $\delta$ is chosen small enough independently of $i$.
So by a compactness argument, a subsequence of flow boxes and their inverses converge in $C^{1-}$ to a local diffeomorphism $F$.
Since $x_0$ was arbitrary, a diagonal argument ensures that we can cover $M$ by sets $U_\alpha := B(x_0, \delta r)$ in which these flow boxes $F_{i\alpha}$ converge to some local diffeomorphism $F_\alpha$.
By taking a further subsequence and rotating, we may assume that $R_i$ converges to the identity as well.

Since $(\lambda_i)$ is tight, we can find a precompact ball $B$ in $M$ so that every leaf of $\lambda_i$ passes through $B/2$.
Then by compactness of $\Hypspace \overline B$, we obtain limits of every sequence $(N_i)$ of $\lambda_i$ in $\Hypspace \overline B$.
By Lemma \ref{limit of minimals is minimal}, such a limit $N$ is a geodesic.
Then $F_{i\alpha}^{-1}(N_i \cap U_\alpha) = \{\eta = \eta_i\}$ for some $\eta_i$, and by the $C^{1-}$-convergence of the $F_i$, the hyperplanes $\{\eta = \eta_i\}$ must converge to some hyperplane $\{\eta = \eta_\infty\}$, so that $F_\alpha(\{\eta = \eta_\infty\}) = N \cap U_\alpha$.
So the limiting geodesics $N$ form a geodesic lamination $\lambda$ and $F_\alpha$ is a flow box for $\lambda$.
Diagonalizing against larger and larger compact balls, we extend the geodesic lamination $\lambda$ in $B$ to a lamination in $M$.

The above convergence is in Thurston's geometric topology.
Indeed, if $N$ is a leaf of $\lambda$, then we can find leaves $N_i$ such that $N_i \to N$ pointwise.
So by Lemma \ref{convergence of geodesic lams in thurston} and the fact that $N$ is a geodesic, we conclude the desired convergence.

Finally, if $\mu_i$ is transverse to $\lambda_i$, then we observe that since $(\lambda_i)$ is tight, we can find a compact subset $K$ of $M$ which meets every leaf of $\lambda$.
Then $\lambda, \lambda_i$ are orientable on $K \cap U_\alpha$, so we can take $T_{\mu_i}|_{K \cap U_\alpha}$ and obtain a convergent subsequence to some $T_\mu|_{K \cap U_\alpha}$; here we use the uniform bounds on $\mu_i(K)$.
Diagonalizing again, we obtain $\mu := |T_\mu|$, which is a transverse measure to $\lambda|_K$, and is globally defined since it is not oriented.
Since $K$ meets every leaf of $\lambda$, $\mu$ extends uniquely to all of $M$ as a transverse measure to $\lambda$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Convergence in flow boxes versus the weak topology of measures}
\begin{theorem}\label{measures implies flow boxes}
Let $(\lambda_i, \mu_i)$ be measured geodesic laminations.
If $(\lambda_i, \mu_i) \to (\lambda, \mu)$ in the weak topology of measures, then $\lambda_i \to \lambda$ on the level of flow boxes.
\end{theorem}
\begin{proof}
The question is local, so we may remove all leaves of $\lambda$ except those that meet some precompact open set.
Then, for each $i$, let $(\lambda_{ij})_j$ be finite sublaminations of $\lambda_i$ converging to $\lambda_i$ on the level of flow boxes.
Then any subsequence $(i_k)$, has a further subsequence $(i_{k_\ell})$ such that $\lambda_{i_{k_\ell} i_{k_\ell}} \to \lambda$ on the level of flow boxes by Theorem \ref{compactness in flow boxes and Thurston}.
Indeed, as geodesic laminations, they have bounded curvature, and $(\lambda_i)$ is tight by construction.
So $\lambda_{i_{k_\ell}} \to \lambda$ on the level of flow boxes, but $(i_k)$ was arbitrary, so $\lambda_i \to \lambda$ on the level of flow boxes.
\end{proof}

% \begin{lemma}
% For every geodesic lamination $\lambda$ and every $x \in \supp \lambda$, there is a transverse curve to $\lambda$ through $x$.
% \end{lemma}
% \begin{proof}
% Since $\normal_\lambda^\sharp$ is Lipschitz with no zeroes, it extends to a Lipschitz vector field $X$ with no zeroes in a neighborhood of $\supp \lambda$.
% Then any integral curve of $X$ through $x$ is transverse to $\lambda$.
% \end{proof}

% \begin{lemma}
% Let $\lambda$ be a geodesic lamination and $\gamma$ a transverse curve to $\lambda$.
% Then $\lambda$ extends to a geodesic foliation $\overline \lambda$ in a neighborhood of $\gamma(I)$, such that the Lipschitz constant of $\normal_{\overline \lambda}$ is the Lipschitz constant of $\normal_\lambda$.
% \end{lemma}
% \begin{proof}
% Extend $(\normal_\lambda^\sharp)^\perp$, which is tangent to the leaves of $\lambda$, to a Lipschitz vector field $Y$ on a neighborhood of $\gamma(I)$ with a comparable Lipschitz constant to $(\normal_\lambda^\sharp)^\perp$. Then
% $$F(k, y) = \exp_{\gamma(k)}(yY(\gamma(k)))$$
% is a flow box for $\lambda$ in a neighborhood of $\gamma$ such that every fiber $\{k\} \times \RR$, not just those for which $k \in K_F$, is a geodesic.
% So $F$ is a flow box for a geodesic foliation $\overline \lambda$ which extends $\lambda$ near $\gamma(I)$.
% Moreover, $\normal_{\overline \lambda} = -(Y^\perp)^\flat$ has a comparable Lipschitz constant to $\normal_\lambda$ by construction.
% \end{proof}

% \begin{lemma}
% If $\lambda_i \to \lambda$ in Thurston's geometric topology and $x \in N$, $N$ a leaf of $\lambda$, then for any sequence of leaves $N_i$ of $\lambda_i$ which contain $x_i \in N_i$ such that $x_i \to x$, $\normal_{N_i} \to \normal_N$ uniformly in a neighborhood of $x$.
% \end{lemma}
% \begin{proof}
% By definition of Thurston's geometric topology, $\normal_{N_i}(x_i) \to \normal_N(x)$.
% However, $N_i, N$ are geodesics, so $\normal_{N_i}, \normal_N$ are covariantly constant.
% In particular they remain close on small sets.
% \end{proof}

% \begin{lemma}
% If $\lambda$ is a geodesic lamination, and $\gamma$ is a transverse curve, there exists a Lipschitz unit-length vector field $X$ along $\gamma$ such that if $\gamma(t) \in N$ for some leaf $N$ of $\lambda$, then $X(\gamma(t))$ is tangent to $N$.
% Moreover, the map that sends a geodesic lamination to a Lipschitz vector field is continuous for Thurston's geometric topology.
% \end{lemma}
% \begin{proof}
% First, $(\normal_\lambda^\sharp)^\perp$ is Lipschitz, unit-length, tangent to the leaves of $\lambda$, and defined on $\supp \lambda \cap \gamma(I)$.
% We then let
% \begin{equation}\label{tangent frame to a lamination}
% 	X := E(\normal_\lambda^\sharp)^\perp.
% \end{equation}

% To see the continuity, suppose that $\lambda_i \to \lambda$ in Thurston's geometric topology.
% Let $x \in \supp \lambda$; by taking a subsequence we can find $x_i \in \supp \lambda_i$ such that $x_i \to x$ and $\normal_{\lambda_i}(x_i) \to \normal_\lambda(x)$.
% So if $N_i, N$ are the leaves containing $x_i, x$, then $\normal_{N_i} \to \normal_N$ uniformly in a neighborhood of $x$.
% Thus (TODO) $\normal_{\lambda_i} \circ \gamma \to \normal_\lambda \circ \gamma$ in $W^{1, \infty}_p(I, M)$, hence if we define $X_i$ by (\ref{tangent frame to a lamination}), $X_i \to X$ in $W^{1, \infty}(I, M)$.
% \end{proof}


% \begin{proof}
% By Proposition \ref{measured implies Thurston}, $\lambda_i \to \lambda$ in Thurston's geometric topology and $\lambda$ is a geodesic lamination.
% Let $\gamma$ be a transverse curve to $\lambda$, and choose a Lipschitz unit vector field $X$ along $\gamma$ which is tangent to the leaves of $\lambda$.
% Then we set
% $$F(k, y) := \exp_{\gamma(k)}(yX(\gamma(k))).$$
% Since $\exp_{\gamma(k)}$ sends lines through the origin to geodesics through $\gamma(k)$, and $X$ is Lipschitz, $F$ is a flow box for $\lambda$.
% Since $\gamma$ is eventually transverse to $\lambda_i$, we obtain flow boxes $F_i$ for each of the $\lambda_i$, with associated vector fields $X_i$ as well, such that $X_i \to X$ in $W^{1, \infty}$. Therefore $F_i \to F$ in $W^{1, \infty}$.
% \end{proof}



%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Construction of minimal laminations}
We conclude this paper by applying the above theory to prove a condition for a collection of minimal hypersurfaces to form a minimal lamination.
This condition will be used in the companion papers \cite{BackusFLG, DaskalopoulosPrep2}.

\begin{definition}
A \dfn{minimal partition} is a closed subset $\lambda$ of $M$ which has been partitioned into (disjoint, embedded, without boundary) minimal hypersurfaces, called the \dfn{leaves} of $\lambda$.
\end{definition}

\begin{theorem}\label{partition implies lamination}
Let $\lambda$ be a minimal partition such that the second fundamental forms of the leaves of $\lambda$ are locally uniformly bounded.
Then $\lambda$ is a minimal lamination.
\end{theorem}
\begin{proof}
Let $(x_i)$ be a dense sequence in $\supp \lambda$, and let $N_i$ be the leaf containing $x_i$.
Then the union of $N_1, \dots, N_i$ is the support of a lamination $\lambda_i$, so that the curvatures of $\lambda_i$ are locally uniformly bounded.
So $(\lambda_i)$ has a maximal subsequential limit $\tilde \lambda$ on the level of flow boxes by Theorem \ref{compactness in flow boxes and Thurston}.
In particular, the leaves of $\tilde \lambda$ are exactly the limits in Hausdorff distance of the leaves of $\lambda$, but $\Leaves \lambda$ is closed so $\lambda = \tilde \lambda$.
\end{proof}

\begin{corollary}
Let $\lambda$ be a geodesic partition. Then $\lambda$ is a geodesic lamination.
\end{corollary}

The above result is also useful when combined with the below stable Bernstein theorem.\footnote{so-called because this theorem is essentially equivalent to the assertion that a stable two-sided minimal hypersurface in $\RR^d$ is a hyperplane.}

\begin{proposition}[stable Bernstein theorem]
Let $3 \leq d \leq 4$ and let $N$ be a two-sided stable minimal hypersurface in $B_r \subseteq M$, $r \lesssim 1$, where $M$ has bounded geometry and dimension $d$.
Then on $B_{r/2}$, $|\Two_N| \lesssim_M r^{-1}$.
\end{proposition}
\begin{proof}
For $d = 3$ see \cite[Corollary 2.11]{colding2011course}, and for $d = 4$ see \cite{Chodosh2021}.
\end{proof}

\begin{corollary}
Let $\lambda$ be a minimal partition in $M$ whose leaves are two-sided and stable, where $M$ has dimension $3 \leq d \leq 4$.
Then $\lambda$ is a minimal lamination.
\end{corollary}

\printbibliography

\end{document}
