\documentclass[reqno,11pt]{amsart}
\usepackage[letterpaper, margin=1in]{geometry}
\RequirePackage{amsmath,amssymb,amsthm,graphicx,mathrsfs,url,slashed,subcaption}
\RequirePackage[usenames,dvipsnames]{xcolor}
\RequirePackage[colorlinks=true,linkcolor=Red,citecolor=Green]{hyperref}
\RequirePackage{amsxtra}
\usepackage{cancel}
\usepackage{tikz-cd}

% \setlength{\textheight}{9.3in} \setlength{\oddsidemargin}{-0.25in}
% \setlength{\evensidemargin}{-0.25in} \setlength{\textwidth}{7in}
% \setlength{\topmargin}{-0.25in} \setlength{\headheight}{0.18in}
% \setlength{\marginparwidth}{1.0in}
% \setlength{\abovedisplayskip}{0.2in}
% \setlength{\belowdisplayskip}{0.2in}
% \setlength{\parskip}{0.05in}
%\renewcommand{\baselinestretch}{1.05}

\title{Regularity of sets of least perimeter on Riemannian manifolds}
\author{Aidan Backus}
\address{Department of Mathematics, Brown University}
\email{aidan\_backus@brown.edu}
\date{\today}

\newcommand{\NN}{\mathbf{N}}
\newcommand{\ZZ}{\mathbf{Z}}
\newcommand{\QQ}{\mathbf{Q}}
\newcommand{\RR}{\mathbf{R}}
\newcommand{\CC}{\mathbf{C}}
\newcommand{\DD}{\mathbf{D}}
\newcommand{\PP}{\mathbf P}
\newcommand{\MM}{\mathbf M}
\newcommand{\II}{\mathbf I}
\newcommand{\Hyp}{\mathbf H}
\newcommand{\Sph}{\mathbf S}
\newcommand{\Group}{\mathbf G}
\newcommand{\GL}{\mathbf{GL}}
\newcommand{\Orth}{\mathbf{O}}
\newcommand{\SpOrth}{\mathbf{SO}}
\newcommand{\Ball}{\mathbf{B}}

\DeclareMathOperator*{\Expect}{\mathbf E}

\DeclareMathOperator{\avg}{avg}
\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\cent}{center}
\DeclareMathOperator{\ch}{ch}
\DeclareMathOperator{\codim}{codim}
\DeclareMathOperator{\Cyl}{Cyl}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\Exc}{Exc}
\newcommand{\ext}{\mathrm{ext}}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Iso}{Iso}
\DeclareMathOperator{\Jac}{Jac}
\DeclareMathOperator{\Lip}{Lip}
\DeclareMathOperator{\Met}{Met}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\rad}{rad}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Rm}{Rm}
\DeclareMathOperator{\Hess}{Hess}
\DeclareMathOperator{\Hol}{Hol}
\DeclareMathOperator{\Prop}{Prop}
\DeclareMathOperator{\Radon}{Radon}
\DeclareMathOperator*{\Res}{Res}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\singsupp}{sing~supp}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\Tan}{Tan}
\newcommand{\tr}{\operatorname{tr}}

\newcommand{\Mink}{\mathbf m}
\newcommand{\Ric}{\mathrm{Ric}}
\newcommand{\Riem}{\mathrm{Riem}}
\newcommand*\dif{\mathop{}\!\mathrm{d}}
\newcommand*\Dif{\mathop{}\!\mathrm{D}}
\newcommand{\LapQL}{\Delta^{\mathrm{ql}}}

\newcommand{\dbar}{\overline \partial}

\DeclareMathOperator{\atanh}{atanh}
\DeclareMathOperator{\csch}{csch}
\DeclareMathOperator{\sech}{sech}

\DeclareMathOperator{\Div}{div}
\DeclareMathOperator{\Gram}{Gram}
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\spn}{span}
\DeclareMathOperator{\Ell}{Ell}
\DeclareMathOperator{\WF}{WF}

\newcommand{\Two}{\mathrm{I\!I}}

\newcommand{\Lagrange}{\mathscr L}
\newcommand{\DirQL}{\mathscr D^{\mathrm{ql}}}
\newcommand{\DirL}{\mathscr D}

\newcommand{\Hilb}{\mathcal H}
\newcommand{\Homology}{\mathrm H}
\newcommand{\normal}{\mathbf n}
\newcommand{\radial}{\mathbf r}
\newcommand{\evect}{\mathbf e}
\newcommand{\vol}{\mathrm{vol}}

\newcommand{\Bmu}{\boldsymbol \mu}
\newcommand{\Bnu}{\boldsymbol \nu}
\newcommand{\Blambda}{\boldsymbol \lambda}

\newcommand{\pic}{\vspace{30mm}}
\newcommand{\dfn}[1]{\emph{#1}\index{#1}}

\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}

\newcommand{\loc}{\mathrm{loc}}
\newcommand{\cpt}{\mathrm{cpt}}

\def\Japan#1{\left \langle #1 \right \rangle}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{badtheorem}[theorem]{``Theorem"}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{sublemma}[theorem]{Sublemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{assumption}[theorem]{Assumption}

\newtheorem{mainthm}{Theorem}
\renewcommand{\themainthm}{\Alph{mainthm}}

% \newtheorem{claim}{Claim}[theorem]
% \renewcommand{\theclaim}{\thetheorem\Alph{claim}}
\newtheorem*{claim}{Claim}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}

\newtheorem{exercise}[theorem]{Discussion topic}
\newtheorem{homework}[theorem]{Homework}
\newtheorem{problem}[theorem]{Problem}

\makeatletter
\newcommand{\proofpart}[2]{%
  \par
  \addvspace{\medskipamount}%
  \noindent\emph{Part #1: #2.}
}
\makeatother



\numberwithin{equation}{section}


% Mean
\def\Xint#1{\mathchoice
{\XXint\displaystyle\textstyle{#1}}%
{\XXint\textstyle\scriptstyle{#1}}%
{\XXint\scriptstyle\scriptscriptstyle{#1}}%
{\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
\!\int}
\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$ }
\vcenter{\hbox{$#2#3$ }}\kern-.6\wd0}}
\def\ddashint{\Xint=}
\def\dashint{\Xint-}

\newcommand\todo[1]{\textcolor{red}{TODO: #1}}

\usepackage[backend=bibtex,style=alphabetic,giveninits=true]{biblatex}
\renewcommand*{\bibfont}{\normalfont\footnotesize}
\addbibresource{topics.bib}
\renewbibmacro{in:}{}
\DeclareFieldFormat{pages}{#1}


\begin{document}
\begin{abstract}
    We show that the reduced boundary of a set of least perimeter in a Riemannian manifold of dimension $\leq 7$ is a smooth minimal hypersurface.
    On euclidean space, this is a classical result of de Giorgi and Miranda, but on Riemannian manifolds it has remained folklore.
\end{abstract}

\maketitle

\section{Introduction}
An open subset $U$ of a Riemannian manifold-with-boundary $M$ is said to have \dfn{least perimeter} if its indicator function $1_U$ has least gradient -- that is, for every $\varphi \in BV(M)$ with $\varphi|_{\partial M} = 0$, 
$$\|\dif 1_U\|_{TV} \leq \|\dif 1_U + \dif \varphi\|_{TV},$$
where $\|\cdot\|_{TV}$ denotes the total variation norm of a Radon measure.
It was proven by de Giorgi and Miranda in the 1960s that if $M$ is an open subset of $\RR^d$, $d \leq 7$, and $U$ has least perimeter, then $U$ is bounded by minimal hypersurfaces \cite{deGiorgi61, Miranda66}.

The arguments of de Giorgi and Miranda rely on two facts about euclidean space:
\begin{enumerate}
\item The tangent bundle to euclidean space is flat, so one can take averages of sections of the tangent bundle.
\item A harmonic function on euclidean space can be decomposed into harmonic polynomials.
\end{enumerate}
In spite of this, the generalization of the de Giorgi--Miranda regularity theorem to Riemannian manifolds $M$ has remained folklore, at least as far as we are aware.

In this paper we establish the de Giorgi--Miranda regularity theorem in general:

\begin{theorem}\label{main thm}
Let $M$ be a Riemannian manifold-with-boundary of dimension $\leq 7$, and let $U \subset M$ have least perimeter.
Then $U$ is bounded by smooth minimal hypersurfaces.
\end{theorem}

The main point of the proof is to establish certain estimates on the parallel transport operator.
Once this has been done, one can imitate the argument of de Giorgi and Miranda, as exposited in the book of Giusti \cite{Giusti77}, taking special care to ensure that one does not obtain any terms involving the curvature of the metric.

Theorem \ref{main thm} has many standard corollaries.
For example, by imitating the argument of \cite{górny2017planar} and simply replacing the de Giorgi--Miranda theorem with Theorem \ref{main thm}, one can easily show:

\begin{corollary}
Let $M$ be a simply connected Riemannian manifold-with-boundary of dimension $\leq 7$, and let $u$ be a function of least gradient on $M$.
Then there exist a continuous function $u_c$ of least gradient, and a jump function $u_j$ of least gradient, such that $u = u_c + u_j$.
\end{corollary}

In an upcoming work \cite{BackusCML}, we will use Theorem \ref{main thm} to prove the following theorem:

\begin{corollary}
Let $T$ be a closed $d-1$-current on a Riemannian manifold of dimension $\leq 4$.
Then $T$ is the Ruelle-Sullivan current of a minimal lamination if and only if $T$ is locally the derivative of a $1$-harmonic function.
\end{corollary}

\subsection{Outline of the paper}
We establish some preliminaries in \S\ref{Prelims}, including estimates on the parallel propagator.
In \S\ref{MollifierSection}, we state the monotonicity formula for functions of least gradient and recall some of its standard consequences.
In \S\ref{DeGiorgiSec}, we state and partially prove the de Giorgi lemma, which controls the oscillation of the normal vector to a minimal perimeter and is well-known to imply Theorem \ref{main thm}.
In \S\ref{Mollifiers}, we complete the proof of the de Giorgi lemma by establishing a technical result about how mollification preserves the oscillation of the normal vector.

\subsection{Acknowledgements}
I would like to thank Georgios Daskalopoulos for suggesting this project and for many helpful discussions, and Christine Breiner for helpful comments.

This research was supported by the National Science Foundation's Graduate Research Fellowship Program under Grant No. DGE-2040433.

%%%%%%%%%%%%%%%%%%%%%%%%
\section{Preliminaries}\label{Prelims}
\subsection{Notation and conventions}
We shall always assume that $M$ has bounded geometry.
This assumption is no loss, since all of our results are local anyways, and only serves to ease notation.

When using the Einstein convention, Greek indices range over $0, 1, \dots$ while Latin indices range over $1, \dots$.
We write $y := x^0$.

We write $\Japan \xi := \sqrt{1 + |\xi|^2}$ for the Japanese norm of a vector $\xi$.
We write $\Ball^d$ for the unit ball in euclidean space, and $\Sph^{d - 1}$ for the unit sphere.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sets of locally finite perimeter}
We recall some elementary facts from geometric measure theory.
Standard references on euclidean space include \cite{simon1983GMT,Giusti77}, and we will only give details when the Riemannian case is significantly more involved than its euclidean counterpart.

Given a function $u \in L^1_\loc(M)$, we can define its \dfn{total variation measure} in an open set $U$ by
$$\|\dif u\|_{TV(U)} = \sup_{\|\psi\|_{C^0} \leq 1} \int_M u \dif \psi$$
where $\psi$ ranges over all $d-1$-forms with compact support in $U$.
We sometimes abuse notation and write $\star |\dif u|$ for the total variation measure, even if $|\dif u|$ is a distribution but not a function.
Similarly we sometimes write $\star \partial_i u$, if we have chosen a coordinate system $(\partial_i)$.

If the total variation measure is a Radon measure, we say that $u \in BV_\loc(M)$, or that $u$ has \dfn{locally bounded variation}.
This is a diffeomorphism-invariant condition.
If $u$ has locally bounded variation and $U \subseteq M$ is open with Lipschitz boundary, then by \cite[Theorem 2.10]{Miranda67}, the trace $u|_{\partial U} \in L^1_\loc(\partial U)$ is well-defined.
Moreover, by \cite[Theorem 4.14]{simon1983GMT}, there exists a $\star|\dif u|$-measurable section $f$, the \dfn{polar part} of $\dif u$, of the cosphere bundle $S'M$ such that
\begin{equation}\label{RNy formula}
\partial_i u = f_i |\dif u|.
\end{equation}
As in \cite{Miranda66, Giusti77}, most of the technical work in this paper amounts to controlling the oscillation of a polar section $f$ at fine scales.
In order to make this precise, we shall need to take averages of $f$ and apply a version of the Lebesgue differentiation theorem for curved vector bundles.

\begin{proposition}[Lebesgue differentiation theorem for a vector bundle]\label{LebesgueDiff}
Let $E \to M$ be a vector bundle over an oriented smooth manifold $M$, $\omega$ a Radon measure, and $f \in L^1_\loc(M, E, \omega)$.
Then there exists an $\omega$-null set $Z \subset M$ such that for every Riemannian metric on $M$, every trivialization $(F_1, \dots, F_\ell)$ of $E$ with dual trivialization $(F'_1, \dots, F'_\ell)$ of $E'$, and every $P \in M \setminus Z$,
$$f(P) = \lim_{r \to 0} \sum_{i=1}^\ell \left[\frac{\int_{B(P, r)} (F'_i, f) \dif \omega}{\omega(B(P, r))}\right] F_i(P).$$
\end{proposition}

We shall apply this proposition with $E := T'M$, $F_i = \dif x^i$.
Note carefully that the terms inside the limit \emph{are} dependent on the metric and the choice of trivialization; the content of this result is that \emph{the set of Lebesgue points is diffeomorphism-invariant}.

\begin{proof}
Choose a flat Riemannian metric, let $\mathcal F = ((F_i), (F_i'))$ be a pair of paralellizations of $E, E'$ such that $(F_i', F_j) = \delta_{ij}$, and $\ell$ the rank of $E$.
Then for every $\delta > 0$ there exists $\tilde f \in C_c(M, E)$ such that $\|f - \tilde f\|_{L^1(\omega)} < \delta$, thus
\begin{align*}
&\left|\sum_{i=1}^\ell \left[(F_i'(x), f(x)) - \dashint_{B(x, r)} (F_i', f) \dif \omega\right] F_i(x)\right| \\
&\qquad \leq \left|\sum_{i=1}^\ell (F_i'(x), f(x) - \tilde f(x)) F_i(x)\right| + \dashint_{B(x, r)} \left|\sum_{i=1}^\ell (F_i', f - \tilde f)F_i(x) \dif \omega \right| \\
&\qquad \qquad + \left|\sum_{i=1}^\ell \left[(F_i'(x), \tilde f(x)) - \dashint_{B(x, r)} (F_i, \tilde f) \dif \omega\right] F_i(x)\right| \\
&\qquad =: I_1(x) + I_{2, r}(x) + I_{3, r}(x).
\end{align*}
Here the integral defining $I_{2, r}(x)$ is valued in the fiber $E_x$.

By the proof of the Lebesgue differentiation theorem, $\{I_1 > \varepsilon\} \subseteq \{|f - \tilde f| > \varepsilon\}$ and $\{I_{2, r} > \varepsilon\} \subseteq \{\dashint_{B(x, r)} |f - \tilde f|\dif \omega\}$, which are $\mathcal F$-independent sets of $\omega$-measure $\lesssim \delta/\varepsilon$.
Meanwhile $I_{3, r} \to 0$ pointwise as $r \to 0$, so for
\begin{equation}\label{definition of null set}
Z_{\varepsilon, \mathcal F} := \left\{x \in M: \limsup_{r \to 0} \left|\sum_{i=1}^\ell \left[(F_i'(x), f(x)) - \dashint_{B(x, r)} (F_i', f) \dif \omega\right] F_i(x)\right| > 2\varepsilon\right\},
\end{equation}
one has
$$Z_{\varepsilon, \mathcal F} \subseteq \bigcap_{r > 0}\bigcup_{s <r} \{I_{1, s} > \varepsilon\} \cup \{I_{2, s} > \varepsilon\}.$$
The right-hand side is independent of $\mathcal F$ and $\delta$, but has $\omega$-measure $\lesssim \delta/\varepsilon$, so it is $\omega$-null.
Thus the union taken over all possible $\mathcal F$ and $\varepsilon$ is also $\omega$-null.

Now let $g$ be a Riemannian metric and $h$ our flat reference metric.
Then the balls $B_g(x, r)$ have bounded eccentricity with respect to $h$, so we can replace $B(x, r)$ with $B_g(x, r)$ in (\ref{definition of null set}) without affecting $Z_{\varepsilon, \mathcal F}$ \cite[Chapter 3, Corollary 1.7]{stein2009real}.
\end{proof}

\begin{corollary}
The polar section $f: M \to S'M$ in (\ref{RNy formula}) satisfies
\begin{equation}\label{Lebesgue point definition}
    f_\mu(P) = \left[\lim_{r \to 0} \frac{\int_{B(P, r)} \partial_\mu u \dif V}{\|\dif u\|_{TV(B(P, r))}}\right]
\end{equation}
for any coordinate system $(x^\mu)$ and any Riemannian metric $g$, and $\star|\dif u|$-almost every $P$.
The exceptional set does not depend on $(x^\mu)$ or $g$.
\end{corollary}

It follows from the above corollary that the following definitions, largely taken from \cite[Definition 3.3]{Giusti77}, which a priori refer to the metric or to a choice of coordinate system, are actually completely determined by the smooth structure on $M$.

\begin{definition}
Let $U \subseteq M$. We say that $U$ has \dfn{locally finite perimeter} if $1_U \in BV_\loc(M)$.
In that case we make the following definitions:
\begin{enumerate}
\item The \dfn{measure-theoretic boundary} $\partial U$ is the set of points whose Lebesgue density with respect to $M$ is $\in (0, 1)$.
\item The polar section of $1_U$ is called the \dfn{conormal $1$-form} $\normal_U$ to $\partial U$.
\item The set of points $P$ for which $\normal_U(P)$ exists is the \dfn{reduced boundary} $\partial^* U$.
\item The \dfn{perimeter} $|\partial^* U \cap E|$ in a Borel set $E$ is $\|d1_U\|_{TV(E)}$.
\end{enumerate}
\end{definition}

Choosing a coordinate system on $M$ in which the volume form is $\dif x^0 \wedge \cdots \wedge \dif x^{d - 1}$, we see from \cite[Chapter 4]{Giusti77} that the following properties of the reduced boundary hold:

\begin{proposition}\label{locality of Caccioppoli}
    Let $U$ be a set of locally finite perimeter.
    Then:
    \begin{enumerate}
    % \item $\partial^* U$ is either empty or $d-1$-dimensional in the Hausdorff sense, and is $d-1$-rectifiable.
    \item $\partial^* U$ is a dense subset of $\partial U$.
    \item If $\normal_U$ extends to a continuous $1$-form on $\partial U$, then $\partial^* U = \partial U$ is a $C^1$ embedded hypersurface.
    \item If $\partial^* U = \partial U$ is a $C^1$ hypersurface, then $\normal_U$ is the conormal $1$-form on $\partial U$ as defined in differential topology, and the total variation measure of $1_U$ is the surface measure on $\partial U$.
\end{enumerate}
\end{proposition}

% As a first application of Proposition \ref{locality of Caccioppoli} we recover the following formulation of the coarea formula.

\begin{proposition}[coarea formula]\label{Coarea2}
Let $u \in BV_\loc(M)$ and $E$ an open set. Then
\begin{equation}\label{coarea formula}
\|\dif u\|_{TV(E)} = \int_{-\infty}^\infty |E \cap \partial^* \{u > y\}| \dif y.
\end{equation}
\end{proposition}
\begin{proof}
Reasoning identically to \cite[Theorem 1.23]{Giusti77}, we may assume that $u \in C^\infty(M)$.
If this is true and also $u$ has no critical points, then (\ref{coarea formula}) follows from Fubini's theorem, the fact that $|E \cap \partial \{u > y\}|$ is the surface area of $E \cap \{u = y\}$ (by Proposition \ref{locality of Caccioppoli}), and the change-of-variables formula.
However the left-hand side of (\ref{coarea formula}) is unaffected by critical points of $u$, and the right-hand side of (\ref{coarea formula}) is unaffected by critical values of $u$ by Sard's theorem, so (\ref{coarea formula}) holds even if $u \in C^\infty(M)$ has critical points.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Functions of least gradient}
We write
$$\eta(u, U) := \inf_{v|_{\partial U} = 0} \|\dif(u + v)\|_{TV(U)}$$
for $u \in BV_\loc(M)$ and $U \subseteq M$ open with Lipschitz boundary, thus $u$ has least gradient iff $\eta(u, M) = \|\dif u\|_{TV(M)}$.
If $u, v \in BV(U)$, then using the coarea formula and reasoning analogously to \cite[Lemma 5.6]{Giusti77}, we obtain the a priori estimates
\begin{align}
|\eta(u, U) - \eta(v, U)| &\leq \|u - v\|_{L^1(\partial U)} \label{a priori estimate 1} \\
\eta(u, U) &\leq \|u\|_{L^1(\partial U)} \leq |\partial U| \cdot \|u\|_{L^\infty(M)}. \label{a priori estimate 2}
\end{align}

\begin{proposition}[Miranda stability theorem]\label{Miranda convergence}
If a sequence of functions $(u_n)$ (not necessarily of the same trace) satisfies for every open $U \Subset M$ with Lipschitz boundary
$$\limsup_{n \to \infty} \|\dif u_n\|_{TV(U)} \leq \liminf_{n \to \infty} \eta(u_n, U) < \infty,$$
and $u_n \to u$ in $L^1_\loc(M)$, then $u$ has least gradient, and $\dif u_n \to \dif u$ in the weak topology of measures.
\end{proposition}
\begin{proof}
The proof is similar to \cite[Teorema 3 and Osservazione 3]{Miranda67}, if we observe that we are allowed to add a term of size $o(1)$ to the right-hand side of the inequalities \cite[(2.8), (2.9), (2.13), and (2.14)]{Miranda67}.
The fact that the convergence in the weak topology of measures is equivalent to the convergence in the sense of \cite[Osservazione 3]{Miranda67} follows from the portmanteau theorem \cite[Theorem 13.16]{klenke2013probability}.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The parallel propagator}
The key new ingredient for the proof of the regularity theorem on Riemannian manifolds is the parallel propagator
$$K(P, Q): T_Q'M \to T_P'M$$
which is defined whenever there exists a unique geodesic $\gamma$ from $Q$ to $P$.
It sends a cotangent vector to $M$ at $Q$ to its parallel transport along $\gamma$, which is a cotangent vector to $M$ at $P$.

To express the parallel propagator in coordinates, we recall from \cite[Chapter II, \S2]{baez1994gauge} that if $\Gamma$ is a square matrix of $1$-forms, its \dfn{path-ordered exponential} along a curve $\gamma: [0, 1] \to M$ is given by 
$$\mathcal Pe^{-\int_\gamma \Gamma} := \sum_{n=0}^\infty (-1)^n \int_{\Delta_n} \prod_{m=1}^n (\Gamma(\gamma(t_i)), \gamma'(t_i)) \dif t$$
where 
$$\Delta_n := \{t \in [0, 1]^n: t_1 \leq t_2 \leq \cdots \leq t_n\}$$
is the standard $n$-simplex.
The path-ordered exponential is then a square matrix; it is defined by a convergent series if $\Gamma$ is continuous \cite[Chapter II, \S2]{baez1994gauge}, and to first order, the path-ordered exponential is
\begin{equation}\label{path ordered exponential taylor series}
\mathcal Pe^{-\int_\gamma \Gamma} = I + O(|\gamma| \cdot \|\Gamma\|_{C^0}).
\end{equation}

\begin{proposition}
Suppose that $U(P, Q)$ exists, and let $\Gamma$ be the Christoffel symbols of the Levi-Civita connection of $M$ acting on the cotangent bundle in some coordinate system, viewed as a $d \times d$-matrix of $1$-forms.
Let $\gamma$ be the unique geodesic from $Q$ to $P$. Then in coordinates,
\begin{equation}\label{path ordered exponential is propagator}
K(P, Q) = \mathcal Pe^{-\int_\gamma \Gamma}.
\end{equation}
\end{proposition}

To be more precise, if we use our coordinate system to identify $T_P'M$ and $T_Q'M$ with $\RR^d$, then the $d\times d$-matrix obtained from the path-ordered exponential can be viewed as a linear map
$$\mathcal Pe^{-\int_\gamma \Gamma}: T_Q'M \to T_P'M.$$
The assertion is that this map is exactly equal to $K(P, Q)$.
For a proof, see \cite[Chapter II, \S2]{baez1994gauge}.

We now use the parallel propagator to define an intrinsic averaging operator on the cotangent bundle.
Such an operator will necessarily depend on the choice of basepoint, since there is no canonical way to identify all of the cotangent spaces to $M$.

\begin{definition}
Let $\mu$ be a Radon measure, $U \subseteq M$ open with finite $\mu$-measure, $P \in U$, and $\xi$ a $1$-form on $U$.
Suppose that for every $Q \in U$ there exists a unique geodesic from $Q$ to $P$.
Then the \dfn{average} of $\xi$ in $U$ with respect to $\mu, P$ is 
$$\avg_{U, P, \mu} \xi := \frac{1}{\mu(U)} \int_U K(P, Q)\xi(Q) \dif \mu(Q).$$
\end{definition}

To be more explicit, $K(P, Q)\xi(Q)$ is a cotangent vector to $P$, so we have a vector-valued map 
\begin{align*}
U &\to T_PM \\
Q &\mapsto K(P, Q)\xi(Q).
\end{align*}
This vector-valued map has target a single vector space, rather than a vector bundle, so it makes sense to take its vector-valued integral.

\begin{proposition}\label{translation invariance}
Let $\mu$ be a Radon measure, $U \subseteq M$ open with finite $\mu$-measure, $P, Q \in U$, and $\xi$ a $1$-form on $U$.
Suppose also that $U$ has diameter $\leq \rho$ and for any $R \in U$ there exist unique geodesics from $R$ to $P$ and $Q$.
Then 
$$||\avg_{U, P, \mu} \xi| - |\avg_{U, Q, \mu} \xi|| \lesssim \rho^2 \|\xi\|_{L^\infty(\mu)}.$$
\end{proposition}
\begin{proof}
Choose normal coordinates at $P$, so that $\|\Gamma\|_{C^0(U)} \lesssim \rho$.
Moreover, the geodesics from $R$ to $P, Q$ have length $\leq \rho$.
Therefore by the reverse triangle inequality,
\begin{align*}
||\avg_{U, P, \mu} \xi| - |\avg_{U, Q, \mu} \xi||
&= \frac{1}{\mu(U)} \left|\left|\int_U \mathcal P e^{-\int_R^P \Gamma} \xi(R) \dif \mu(R)\right| - \left|\int_U \mathcal P e^{-\int_R^Q \Gamma} \xi(R) \dif \mu(R)\right|\right| \\
&\leq \frac{\|\xi\|_{L^\infty(\mu)}}{\mu(U)} \int_U \left|\mathcal P e^{-\int_R^P \Gamma} - \mathcal Pe^{-\int_R^Q \Gamma}\right| \dif \mu(R) \\
&\leq \|\xi\|_{L^\infty(\mu)} \sup_{R \in U} \left|\mathcal Pe^{-\int_R^P \Gamma} - \mathcal Pe^{-\int_R^Q \Gamma}\right|.
\end{align*}
By (\ref{path ordered exponential taylor series}), this quantity is $\lesssim \rho^2 \|\xi\|_{L^\infty(\mu)}$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Monotonicity formula}\label{MollifierSection}
Let $u$ be a function of least gradient.
If $u$ is defined on euclidean space, then $u$ satisfies a monotonicity formula \cite[Theorem 5.12]{Giusti77}, and the main idea of the proof is to bound the growth of $\|\dif u\|_{TV}$ using the vector-valued integral of $\dif u$ -- that is,
\begin{equation}\label{integral of du}
I(u, P, r) := \int_{B(P, r)} K(P, Q) \frac{\dif u}{|\dif u|}(Q) \dif \mu(Q)
\end{equation}
where $\dif u/|\dif u|$ is the polar section and $\dif \mu := \star|\dif u|$ is the total variation measure.
We now apply the above averaging techniques to extend the monotonicity formula to the manifold case.

\begin{proposition}[monotonicity formula]\label{Monotone}
Let $u$ be a function of least gradient on a Riemannian manifold $M$, and let $\mu$ be its total variation measure.
Then there exists $A \geq 0$ and $r_* > 0$, which only depend on $M$, such that for $0 < r \leq r_*$,
\begin{equation}\label{weak monotonicity}
\frac{\dif}{\dif r}\left[e^{Ar^2}r^{1 - d} \mu(B(P, r))\right] \geq 0.
\end{equation}
and for $0 < r_1 < r_2 \leq r_*$,
\begin{align*}
&|r_2^{1 - d} I(u, P, r_2) - r_1^{1 - d} I(u, P, r_1)|^2 \\
&\qquad \lesssim \left(1 + (d - 1) \log \frac{r_2}{r_1}\right) \left(r_2^{1 - d}\mu(B(P, r_2)) \right)
\left(\int_{r_1}^{r_2} \partial_r \left[e^{Ar^2} r^{1 - d} \mu(B(P, r))\right] \dif r\right)\\
&\qquad \qquad + |K|^2 r_2^{6-2d} \left(\mu(B(P, r_2))\right)^2.
\end{align*}
\end{proposition}

Throughout the proof of Proposition \ref{Monotone} we write $\dif \sigma$ for the usual volume form on $\Sph^{d - 1}$.
We begin with an estimate for a smoothed out version of functions of least gradient.
Its euclidean case can be isolated from the proof of \cite[Lemma 5.8]{Giusti77}.

\begin{lemma}\label{monotonicity lemma}
There exists $A$ such that for every $u \in C^1(B_R)$, $0 < r_1 < r_2 < R$, if we let
$$E(r) = \mu(B_r) - \eta(u, r),$$
so that $E(R) = 0$ iff $u$ has least gradient, then there exists $A \geq 0$ such that for $R > 0$ small,
\begin{equation}\label{monotonicity lemma eqn}
0 \leq \int_{B_{r_2} \setminus B_{r_1}} \star r^{1 - d}\frac{(\partial_ru)^2}{|\dif u|} \leq 2\int_{r_1}^{r_2} \partial_r \left[e^{Ar^2} r^{1-d}\int_{B_r} \star |\dif u|\right] + \frac{O(E(r))}{r^d} \dif r.
\end{equation}
\end{lemma}
\begin{proof}
This result is coordinate-invariant, so we may use whichever coordinates are convenient: we in fact use normal polar coordinates $(r, \theta)$.
We fix $s \in [r_1, r_2]$ and introduce a competitor $v(r, \theta) = u(s, \theta)$.
From the definition of $\eta$,
\begin{equation}\label{consequence of least gradient monotone}
    \eta(u, s) \leq \int_U \star |\dif v| = \int_0^s \int_{\partial B_r} \star_r |\dif v| \dif r.
\end{equation}
We now recall that
$$\dif S_{\partial B_\rho}(\theta) = \left[\rho^{d - 1} - \frac{\rho^{d + 1}}{6} \Ric_P(\theta, \theta) + O(\rho^{d + 2})\right] \dif \sigma(\theta)$$
where the implied constant depends on the curvature of $M$.
Thus we can find $A > 0$ such that for all $\rho$ small enough that $e^{A\rho^2} \sqrt{\det g|_{\partial B_\rho}}$ is monotone in $\rho$ for some $A > 0$, as long as $\rho$ is small enough.
Applying $\partial_r v = 0$ it follows that
\begin{equation}\label{introduce the ricci tensor}
\int_{\partial B_r} \star_r |\dif v| \leq e^{As^2} \frac{\tilde r^{d - 1}}{s^{d - 1}} \int_{\partial B_s} \star_s |\dif v|.
\end{equation}
Applying (\ref{consequence of least gradient monotone}) and Fubini's theorem,
\begin{align*}
\eta(u, s) &\leq e^{As^2} \int_0^s \frac{r^{d - 1}}{s^{d - 1}} \dif r \cdot \int_{\partial B_s} \star_s |\dif v| = \frac{s e^{As^2}}{d} \int_{\partial B_s} \star_s |\dif v|\\
&\leq \frac{s e^{As^2}}{d - 1} \int_{\partial B_s} \star_s |\dif v|.
\end{align*}
By Gauss' lemma, $\dif v$ is the orthogonal projection of $\dif u$ onto $T' \partial B_s$, and its orthocomplement is $\partial_r u$. Therefore by Taylor's theorem,
$$\int_{\partial B_s} \star_s |\dif v| \leq \int_{\partial B_s} \star_s |\dif u| \sqrt{1 - \frac{(\partial_r u)^2}{|\dif u|^2}} \leq \int_{\partial B_s} \star_s \left[|\dif u| - \frac{(\partial_r u)^2}{2 |\dif u|}\right]$$
or in other words
\begin{align*}
\int_{\partial B_s} \star_s \frac{(\partial_r u)^2}{2|\dif u|} &\leq \int_{\partial B_s} \star_s |\dif u| - \frac{d - 1}{s} e^{-As^2} \eta(u, s)\\
&\leq \int_{\partial B_s} \star_s |\dif u| - \frac{d - 1}{s} e^{-As^2} \int_{B_s} \star |\dif u| - O(s^{-1}E(s)).
\end{align*}
We moreover have for $\tilde A \geq 0$ that
$$e^{-\tilde As^2} \partial_s \left[e^{\tilde As^2} s^{1 - d} \int_{B_s} \star |\dif u|\right] = \left[2\tilde As^{2 - d} - \frac{d - 1}{s^d}\right]\int_{B_s} \star |\dif u| + s^{1 - d} \int_{\partial B_s} \star_s |\dif u|$$
so if we choose $\tilde A$ so that
$$-\frac{d - 1}{s} e^{-As^2} = 2\tilde As - \frac{d - 1}{s}$$
then
$$s^{1 - d} \int_{\partial B_s} \star_s |\dif u| - (d - 1)\frac{e^{-\tilde As^2}}{s^d} \int_{B_s} \star|\dif u| \leq e^{-\tilde As^2} \partial_s\left(e^{\tilde As^2} s^{1 - d} \int_{B_s} \star|\dif u|\right).$$
We moreover have $e^{-\tilde As^2} \leq 1$, so we can now integrate with respect to $\dif s$ and rename $\tilde A$ to $A$ to conclude.
\end{proof}

\begin{proof}[Proof of Proposition \ref{Monotone}]
Working in normal coordinates $(x^\mu)$ based at $P$, we first computing using (\ref{path ordered exponential taylor series})
$$ := r^{1 - d} I(u, P, r) = r^{1 - d} \left[\int_{B(P, r)} \star \partial_\alpha u\right] \dif x^\alpha(P) + O(r^{3 - d}) \|\dif u\|_{TV(B(P, r))}.$$
Let $(\iota_0, \dots, \iota_{d - 1}): \Sph^{d - 1} \to \RR^d$ be the embedding of the unit sphere.
Applying the fundamental theorem of calculus,
\begin{align*}
(I_r)_\alpha &= \left[\int_{\Sph^{d - 1}} u(r, \theta) \iota_\alpha(\theta) \dif \sigma(\theta)\right] + O(r^{3 - d}) \|\dif u\|_{TV(B(P, r))}
\end{align*}
and hence
\begin{equation}\label{monotone dump the metric}
|I_{r_2} - I_{r_1}| \leq \int_{\Sph^{d - 1}} |u(r_2, \theta) - u(r_1, \theta)| \dif \sigma(\theta) + O(r^{3 - d}) \|\dif u\|_{TV(B(P, r))}.
\end{equation}
Henceforth we write $B_r := B(P, r)$.
The metric $g$ plays no role in the dominant term of (\ref{monotone dump the metric}), so we may use \cite[Lemma 5.3]{Giusti77} to bound
$$0 \leq \int_{\Sph^{d - 1}} |u(r_2, \theta) - u(r_1, \theta)| \dif \sigma(\theta) \leq \int_{\Sph^{d - 1}} \int_{r_1}^{r_2} r^{1 - d}|\partial_r u(r, \theta)| \dif r \dif\sigma(\theta).$$
To reintroduce the metric we posit that $r_2$ is small enough that $\dif r \dif \sigma(\theta) \leq \star 2$.
We therefore have
\begin{equation}\label{monotone before cs}
\int_{\Sph^{d - 1}} \int_{r_1}^{r_2} r^{1 - d}|\partial_r u(r, \theta)| \dif r \dif\sigma(\theta) \leq 2 \int_{B_{r_2} \setminus B_{r_1}} \star r^{1 - d}|\partial_r u|
\end{equation}
and if we apply the Cauchy-Schwarz inequality and approximate $u$ by $C^1$ functions as on \cite[pg68]{Giusti77}, it follows from Lemma \ref{monotonicity lemma} that the right-hand side of (\ref{monotone before cs}) is
$$\lesssim \sqrt{\int_{B_{r_2} \setminus B_{r_1}} \star r^{1 - d} |\dif u|} \sqrt{\int_{r_1}^{r_2} \partial_r \left[e^{Ar^2} r^{1-d}\int_{B_r} \star |\dif u|\right] \dif r}.$$
The monotonicity (\ref{weak monotonicity}) follows at once.

Integrating by parts,
\begin{align*}
\int_{B_{r_2} \setminus B_{r_1}} \star r^{1 - d} |\dif u| &= \int_{r_1}^{r_2} r^{1 - d} \partial_r \int_{B_r} \star |\dif u| \dif r \\
&\leq r^{1 - d} \int_{B_r} \star |\dif u| + (d - 1) \int_{r_1}^{r_2} r^{-d} \int_{B_r} \star |\dif u| \dif r.
\end{align*}
Using (\ref{weak monotonicity}) we bound this second integral as
\begin{align*}
\int_{r_1}^{r_2} r^{-d} \int_{B_r} \star |\dif u| \dif r &\leq r^{1 - d} \log \frac{r_2}{r_1} \int_{B_{r_2}} \star |\dif u|.
\end{align*}
If we set
$$J_r := r^{1 - d} \int_{B_r} \star |\dif u|$$
then we can sum up our progress so far as
$$|I_{r_2} - I_{r_1}| \lesssim \sqrt{\left[1 + \log \frac{r_2}{r_1}\right] J_{r_2}} \sqrt{e^{Ar_2^2} J_{r_2} - e^{Ar_1^2} J_{r_1}} + r_2^2 J_{r_2}.$$
The claim now follows by squaring both sides and applying Cauchy-Schwarz.
\end{proof}

Since we are considering the regularity of minimal surfaces, we need to use the $d \leq 7$ hypothesis to show the regularity of tangent cones to minimal surfaces.

\begin{definition}
    For a function $u$ on $M$, $P \in M$ we define the \dfn{blowup} of $u$ at $P$ to be the net of functions $u_t: T_PM \to \RR$, given by
    $$u_t(v) = u\left(\exp_P(tv)\right).$$
\end{definition}

\begin{corollary}\label{blowup theorem}
Suppose that $U$ is an open set with least perimeter in $B(P, r)$, $P \in \partial^* U$, and $u = 1_U$.
Furthermore, suppose that $d \leq 7$.
Then the blowup $(u_t)$ of $u$ converges as $t \to 0$ along a subsequence (that we also denote $t \to 0$) in $L^1_\loc$ and almost everywhere, to the indicator function $v$ of a half-space $C \subset T_PM$ such that $0 \in \partial C$.
Moreover, $\dif u_t \to \dif v$ in the weak topology of measures.
\end{corollary}
\begin{proof}
If we only require that $\partial C$ is a minimal cone rather than a hyperplane, then this is a standard consequence of the Miranda stability theorem (Proposition \ref{Miranda convergence}) and the monotonicity formula (\ref{weak monotonicity}), see for example \cite[Theorem 9.3]{Giusti77} for the euclidean case which is essentially identical to the proof here.
However, a minimal cone of dimension $\leq 6$ is necessarily a hyperplane \cite[Theorem 9.10 and Theorem 10.10]{Giusti77}.
\end{proof}

We now generalize the surface area estimates of \cite[Remark 5.13]{Giusti77}.

\begin{corollary}\label{doubling dimension}
If $d \leq 7$ then there exists $A \geq 0$ such that for every set $U$ of least perimeter in a ball $B_r = B(P, r)$, with $P \in \partial^* U$, and $r > 0$ small,
$$|\Ball^{d - 1}|e^{-Ar^2}r^{d - 1} \leq |\partial^*U \cap B_r| \leq |\Sph^{d - 1}|e^{Ar^2} r^{d - 1}.$$
\end{corollary}
\begin{proof}
The upper bound on $|\partial^* U \cap B_r|$ is obtained by using (\ref{a priori estimate 2}) and the fact that the surface area of $\partial B_r$ is $|\Sph^{d - 1}|(1 + O(r^2))r^{d - 1}$.
This can be seen by integrating $\star_{\partial B_r} 1$ along $\partial B_r$ in normal coordinates and applying the Taylor expansion of the Riemannian measure.
The lower bound is obtained from the monotonicity formula, which implies that
$$\limsup_{\rho \to 0} e^{-A\rho^2} \rho^{1 - d} |\partial^* U \cap B_\rho| \leq |\partial^* U \cap B_r|.$$
To control the left-hand side we take a blowup $(u_\rho)$ of $1_U$.
By Corollary \ref{blowup theorem} we can pass to a subsequence so that $u_\rho \to 1_C$ for $C$ a half-space, which in particular is transverse to $B'_1$, where the prime denotes the euclidean metric on the tangent space.
Then
\begin{align*}
\limsup_{\rho \to 0} e^{-A\rho^2} \rho^{1 - d} |\partial^* U \cap B_\rho| &= \lim_{\rho \to 0} e^{O(\rho^2)} \int_{B'_1} \star'|\dif u_\rho|' = \int_{B'_1} \star'|\dif 1_C|.
\end{align*}
This last term is $|\partial C \cap B'_1|$, the measure of the intersection of the euclidean unit ball with a hyperplane through its origin.
In other words it is the measure $|\Ball^{d - 1}|$ of the unit ball in $\RR^{d - 1}$.
\end{proof}

We will frequently use Corollary \ref{doubling dimension} to bound error terms.
For example, if $u = 1_U$ where $U$ has least perimeter, then the error term in the monotonicity formula is of size $O(r_2^{d + 1})$.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The de Giorgi lemma}\label{DeGiorgiSec}
We now introduce the quantity which governs the rate of convergence of the Lebesgue differentiation theorem for $\normal_U$, whenever $U$ is a set of locally finite perimeter.
More precisely, let $\mu$ be the surface measure of $\partial^* U$, which is the total variation measure of $1_U$.
We study the convergence of the approximation
$$\normal_U(P, r) := \avg_{B(P, r), P, \mu} \normal_U.$$

\begin{definition}
The \dfn{excess} of a set $U \subset M$ of locally finite perimeter at $P \in \partial U$, such that $\partial^* U$ has surface measure $\mu$, in an open set $A \ni P$ with Lipschitz boundary is
$$\Exc_A(U, P) := \mu(U)\left(1 - \left|\avg_{A, P, \mu} \normal_U\right|\right).$$
For $\rho > 0$ we write $\Exc_\rho(U, P) := \Exc_{B(P, \rho)}(U, P)$.
\end{definition}

Since parallel transport preserves length, we have
$$|K(P, Q) \normal_U(Q)| = 1$$
and, averaging $K(P, Q) \normal_U(Q)$ in $Q$, we conclude
\begin{equation}\label{normal isnt too big}
|\normal_U(P, r)| \leq 1.
\end{equation}
From this we obtain the following monotonicity property: if $P \in A'$ and $A' \subseteq A$, then
\begin{equation}\label{approximate monotone}
0 \leq \Exc_{A'}(U, P) \leq \Exc_A(U, P).
\end{equation}
Moreover, by Proposition \ref{translation invariance}, if $P, Q \in A$ then 
\begin{equation}\label{translation invariant excess}
|\Exc_A(U, P) - \Exc_A(U, Q)| \lesssim (\diam A)^2 |\partial^* U \cap A|.
\end{equation}

Following \cite{Miranda66,Giusti77,deGiorgi61}, we proceed by controlling the excess using the following de Giorgi lemma, which is the Riemannian generalization of \cite[Theorem 8.1]{Giusti77}:

\begin{proposition}[de Giorgi lemma]\label{de Giorgi}
There exist constants $C, c, \rho_* > 0$ which only depend on $M$, such that for every $P \in M$, $\rho$ such that $0 < \rho < \rho_*$, and set $U \subset M$ of least perimeter such that
\begin{equation}\label{base case}
\Exc_\rho(U, P) \leq c\rho^{d - 1},
\end{equation}
we have
\begin{equation}\label{dGL concl}
\Exc_{\rho/2}(U, P) \leq 2^{-d} \Exc_\rho(U, P) + C\rho^{d + 1}.
\end{equation}
\end{proposition}

Before we prove the de Giorgi lemma we explain why it implies Theorem \ref{main thm}, using an induction on scale.
The base case is proven by \cite[pg109]{Giusti77} and the inductive case is proven by \cite[Corollary 8.3]{Giusti77}.
There are not new ideas in these corollaries, but we do have a new error term in (\ref{dGL concl}) that was not in \cite[Theorem 8.1]{Giusti77}, so for completeness we reprove them.

\begin{corollary}[base case]
Assume the de Giorgi lemma, and let $U \subset M$ have least perimeter.
Then there exists $\rho = \rho(P) < \rho_*$, which is locally uniform in $P \in \partial U$, such that (\ref{base case}) holds.
\end{corollary}
\begin{proof}
Let $Q \in \partial^* U$; we shall choose $\rho$ uniformly in a small neighborhood of $Q$, which is enough since $\partial^* U$ is dense in $\partial U$.
Since $\partial U$ has a tangent space at $Q$ by Corollary \ref{blowup theorem}, $\Exc_r(U, Q) \ll r^{d - 1}$, thus we can choose $r \in (0, \rho_*)$ such that $\Exc_r(U, Q) \leq c(r/2)^{d - 1}$.
For $P \in B(Q, r/4)$ we have $B(P, r/4) \subseteq B(Q, r/2)$ and hence by the de Giorgi lemma, for $\rho := r/4$, (\ref{base case}) holds.
\end{proof}

\begin{corollary}[inductive case]
Assume the de Giorgi lemma, and let $U \subset M$ have least perimeter.
Then $\normal_U$ extends to a continuous $1$-form on $\partial U$, where $\partial U$ is endowed with the subspace topology induced by $M$.
\end{corollary}
\begin{proof}
We use the de Giorgi lemma inductively as in \cite[Theorem 8.2]{Giusti77}.
Suppose that $r/2 < s < r = \rho/2^n$ for some $n$ and some $\rho$ satisfying (\ref{base case}).
To fix notation, let
$\xi := \normal_U(P, r)$, $\eta = \normal_U(P, s)$, $m := |\partial^* U \cap B(P, s)|$, $M := |\partial^* U \cap B(P, r)|$, and $\gamma_n := \Exc_{\rho/2^n}(U, P)$.
We first estimate
$$|\xi - \eta|^2 = |\xi|^2 + |\eta|^2 - 2 g^{-1}(\xi, \eta) \leq 2(1 - g^{-1}(\xi, \eta)).$$
To estimate the right-hand side, let $\mu$ be the surface measure on $\partial^* U$. Then
$$m(1 - g^{-1}(\xi, \eta)) = \int_{B(P, s)} (1 - g^{-1}(\xi, K(P, Q) \normal_U(Q))) \dif \mu(Q) =: I.$$
By the Cauchy-Schwarz inequality, (\ref{normal isnt too big}), and the fact that $K(P, Q)$ preserves length,
$$g^{-1}(\xi, K(P, Q) \normal_U(Q)) \leq 1,$$
which implies that, since $s \leq r$,
\begin{align*}
I
&\leq \int_{B(P, r)} (1 - g^{-1}(\xi, K(P, Q) \normal_U(Q))) \dif \mu(Q) 
\leq M(1 - |\xi|^2) \leq 2M(1 - |\xi|).
\end{align*}
But $M(1 - |\xi|)$ is exactly the definition of $\Exc_r(U, P)$, so putting everything together and using Corollary \ref{doubling dimension} to bound $m \gtrsim s^{d - 1} \gtrsim r^{d - 1}$, we have the bound
\begin{equation}\label{just need the excess}
|\xi - \eta|^2 \lesssim r^{1 - d} \Exc_r(U, P).
\end{equation}
Then, since $r = \rho(V)/2^n$, $\Exc_r(U, P) = \gamma_n$, and for $C' := C\rho^{d + 1}$, we have by the de Giorgi lemma that
\begin{equation}\label{induction on gamma}
\gamma_n \leq \frac{\gamma_0}{2^{nd}} + \sum_{k=0}^n \frac{C'}{2^{k(d + 1) + (n - k)d}} \leq \frac{\gamma_0 + C'}{2^{nd}}.
\end{equation}
Indeed, by induction,
\begin{align*}
\gamma_{n + 1}
&\leq \frac{\gamma_0}{2^{(n + 1)d}} + 2^{-d} \sum_{k=0}^n \frac{C'}{2^{k(d + 1) + (n - k)d}} + \frac{C'}{2^{(n + 1)(d + 1)}} \\
&= \frac{\gamma_0}{2^{(n + 1)d}} + \sum_{k=0}^{n + 1} \frac{C'}{2^{k(d + 1) + (n + 1 - k)d}}
\end{align*}
and (\ref{induction on gamma}) follows by summing the geometric series.
Reasoning as on \cite[pg100]{Giusti77} we conclude that for \emph{any} $s < r < \rho$, not just those of the form $r/2 < s < r = \rho/2^n$,
$$|\normal_U(P, r) - \normal_U(P, s)| \lesssim \sqrt{\frac{r}{\rho}}$$
and so $\normal_U(\cdot, r) \to \normal_U$ locally uniformly.
Since $\normal_U(\cdot, r)$ is continuous, the claim follows.
\end{proof}

\begin{proof}[Proof of Theorem \ref{main thm}, assuming the de Giorgi lemma]
If $U$ is a set of least perimeter, then $\normal_U$ is continuous for the subspace topology on $\partial U$, so by Proposition \ref{locality of Caccioppoli} it is a $C^1$ embedded hypersurface.
But $U$ is a global minimizer of area, so it is a $C^1$ stable minimal hypersurface, hence a $C^\infty$ stable minimal hypersurface.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The smooth case}
We begin the proof of the de Giorgi lemma with the following special case which is analogous to \cite[Lemma 6.4]{Giusti77}.
To state it, let $c_0$ be a given small constant, and $\alpha = \frac{1}{2} + O(c_0)$ also given.
These constants will be chosen later only depending on $M$, not on $U$.
Implied constants and constants called $C$ shall also only depend on $M$.

\begin{proposition}[de Giorgi lemma, $C^1$ case]\label{Miranda44}
Suppose that $c_0$ is small enough depending on $M$.
Then there exists a constant $c_1$, which only depends on $c_0$ and $M$, such that the following holds.
For every $\rho, \beta > 0$ small enough depending on $c_0$, and every set $U$ with $C^1$ boundary in $B(P, \rho)$, suppose that 
$$\Exc_\rho(U, P) \leq \beta.$$
Furthermore, suppose that there is a vector field $X$, which is a coordinate vector field for some choice of normal coordinates based at $P$, such that on $B(P, \rho)$,
\begin{equation}
    (\normal_U, X) \geq e^{-o(c_1^2)}. \label{Miranda44 normal hyp}
\end{equation}
Finally, suppose that
\begin{equation}
|\partial^* U \cap B(P, \rho)| \leq \eta(U, B(P, \rho)) + c_1 \beta. \label{Miranda44 minimality hyp}
\end{equation}
Then
\begin{equation}\label{Miranda44 concl}
\Exc_{\alpha \rho} (U, P) \leq (\alpha^{d + 1} + c_0) \beta + C\rho^{d + 1}.
\end{equation}
\end{proposition}

To set up the proof, suppose that we have normal coordinates $(x^0, \dots, x^{d - 1})$ based at $P$ such that $X = \partial_0$.
We shall say that a $C^1$ hypersurface $N \subset M$ is \dfn{graphical} over a relatively open set $\Omega \subseteq \{x^0 = 0\}$ if for every integral curve $\gamma$ of $X$ passing through $\Omega$, $N$ intersects exactly once, and transversely.\footnote{Elsewhere in the literature this condition is called \dfn{strongly graphical}, and \dfn{graphical} means that the intersection may fail to be transverse.}
In that case we may define $u(x)$, $x \in \Omega$, to be the unique time at which a particle traveling at unit speed along an integral curve, starting at $x$, passes through $N$; since $N$ is $C^1$ and intersects the integral curve transversely, $\dif u$ is continuous.

\begin{lemma}
There is a map $\Lagrange$ taking $1$-jets on $\Omega$ to area forms on $\Omega$, with the following property: for any graphical hypersurface $N$ over $\Omega$, with graphical function $u$, the area of $N$ is
$$|N| = \int_\Omega \Lagrange(u, \dif u).$$
Moreover, if $\|\dif u\|_{C^0} \leq 1$, then (with $\dif x = \dif x^1 \wedge \cdots \wedge \dif x^{d - 1}$)
$$\Lagrange(u, \dif u) = \sqrt{1 + |\dif u|^2 + O(|x|^2 + \|u\|_{C^0}^2)} \dif x.$$
\end{lemma}
\begin{proof}
Let $\Psi: \Omega \to N$ be the diffeomorphism $x \mapsto (x, u(x))$.
For any $v, w \in T_x \Omega$, let $h(v, w) := g(x, u(x))(v, w)$, $h(X, v) := g(x, u(x))(X(x, u(x)), v)$, and $h(X, X) := g(x, u(x))(X(x, u(x)), X(x, u(x)))$.
These quantities are defined in coordinates, because we can use the coordinates to identify $T_{(x, 0)}M$ with $T_{(x, u(x))} M$, so that the quadratic form $g(x, u(x))$ on $T_{(x, u(x))} M$ acts on $T_{(x, 0)}M$.
In these coordinates we have $\Psi_* v = v + (\partial_v u) X(x, u(x))$, so that
\begin{align*}
\Psi^* g(v, w)
&= g(\Psi_* v, \Psi_* w) \\
&= h(v, w) + h(X, v) \partial_w u + h(X, w) \partial_v u + h(X, X) \partial_v u \partial_w u.
\end{align*}
If we set $v = \partial_i$ and $w = \partial_j$, $i, j = 1, \dots, d - 1$, then from the fact that the coordinates are normal we obtain $h(v, w) = \delta_{ij} + O(|x|^2 + u(x)^2)$, $h(X, v) = O(|x|^2 + u(x)^2)$, and $h(X, X) = 1 + O(|x|^2 + u(x)^2)$.
Moreover, $\partial_i u \partial_j u$ are the components of $\dif u \otimes \dif u$.
In conclusion,
$$\Psi^* g(\partial_i, \partial_j) = I + \dif u \otimes \dif u + O((|x|^2 + u(x)^2)(1 + |\dif u(x)|)).$$
However, $1 + |\dif u(x)| \leq 2$, which can be absorbed.
Furthermore, by \cite[(24)]{Petersen2008}, the determinant of $I + \dif u \otimes \dif u$ is $1 + (|\dif u|')^2$, which up to an error of size $|x|^2 + u(x)^2$ is equal to $|\dif u|^2$, since we are in normal coordinates.
The pullback of the area form of $N$ by $\Psi$ is $\sqrt{\det((\Psi^* g(\partial_i, \partial_j))_{ij})}$, so we get 
\begin{align*}
|N| &= \int_\Omega \sqrt{1 + |\dif u(x)|^2 + O(|x|^2 + \|u\|_{C^0}^2)} \dif x. \qedhere
\end{align*}
\end{proof}

Let us write $\mathscr B_\rho$ for a ball in $\Omega$, of radius $\rho$ centered on the coordinate origin $P$.
If the graphical hypersurface $N$ intersects $\Omega$ at $P$, then on $\mathscr B_\rho$,
\begin{equation}\label{approximate by euclidean lagrangian}
\Lagrange(u, \dif u) - \Lagrange(u, \avg_\rho \dif u) = \left(\Japan{\dif u} - \Japan{\avg_\rho \dif u}\right) \dif x + O(\rho^2)
\end{equation}
where $\avg_\rho := \avg_{\mathscr B_\rho, P, \dif x}$, taken using the flat metric on $\Omega$ induced by the coordinates $x$.
Here $\Japan{\dif u} := \sqrt{1 + |\dif u|^2}$ is the Japanese norm of $\dif u$, so that $\Japan{\dif u} \dif x$ is exactly the Lagrangian density for the euclidean minimal surface equation.
Thus, we reduce the problem of estimating $\Lagrange(u, \dif u) - \Lagrange(u, \avg_\rho \dif u)$ to the euclidean case, and hence can show the following analogue of \cite[Lemma 6.3]{Giusti77}.

\begin{lemma}[de Giorgi lemma, minimal graphs]\label{Miranda43}
There exists $c_1 = c_1(c_0, M) > 0$ such that for every $\beta, \rho > 0$, the following holds.
Let $N$ be a $C^1$ graphical hypersurface over $\mathscr B_\rho$, with graphical function $u$, which intersects $\Omega$ at $P$.
Let $I$ be the union of all integral curves of $X$ through $\mathscr B_\rho$.
Suppose that $\|\dif u\|_{C^1} \leq c_1$, and that
\begin{align}
\int_{\mathscr B_\rho} \Lagrange(u, \dif u) - \Lagrange(u, \avg_\rho \dif u) &\leq \beta \label{Miranda43 oscillation}, \\
\int_{\mathscr B_\rho} \Lagrange(u, \dif u) &\leq \eta(N, I) + c_1 \beta \label{Miranda43 minimality}.
\end{align}
Then for any $\tilde \alpha \in (0, 1)$,
\begin{equation}\label{Miranda43 concl}
\int_{\mathscr B_{\tilde \alpha \rho}} \Lagrange(u, \dif u) - \Lagrange(u, \avg_{\tilde \alpha \rho} \dif u) \leq \left(\tilde \alpha^{d + 1} + \frac{c_0}{2}\right) \beta + C\rho^{d + 1}.
\end{equation}
\end{lemma}
\begin{proof}
Let $v$ be the harmonic function on $\mathscr B_\rho$ (with its euclidean metric) which equals $u$ on $\partial \mathscr B_\rho$.
By elliptic regularity, the maximum principle for harmonic functions, and (\ref{Miranda43 minimality}),
$$\|v\|_{C^1} \lesssim \|v\|_{C^0} \leq \|u\|_{C^0} \leq \rho \|\dif u\|_{C^1} \leq c_1.$$
In particular, $\Japan{\dif v} \lesssim 1$ and $v(x)^2 \lesssim \rho^2$, so by (\ref{approximate by euclidean lagrangian}) and the fact that $|\mathscr B_\rho| \sim \rho^{d - 1}$,
\begin{align*}
&\left|\int_{\mathscr B_\rho} \Lagrange(u, \dif u) - \Lagrange(v, \dif v) - \Japan{\dif u} \dif x + \Japan{\dif v} \dif x\right| \\
&\qquad \lesssim \int_{\mathscr B_\rho} (|x|^2 + u(x)^2) \Japan{\dif u} \dif x + (|x|^2 + v(x)^2) \Japan{\dif v} \dif x
\lesssim \rho^{d + 1}.
\end{align*}
Let $K$ be the graph of $v$, viewed as a submanifold of $M$.
Since $u$ and $v$ have the same trace, $|K| \geq \eta(K, I) = \eta(N, I)$, so
\begin{align*}
\int_{\mathscr B_\rho} \Lagrange(u, \dif u) - \Lagrange(v, \dif v) &\leq \int_{\mathscr B_\rho} \Lagrange(u, \dif u) - \eta(\Psi_w(\Omega), I) \leq c_1 \beta.
\end{align*}
We can then replace $\beta$ with $\beta + O(\rho^{d + 1})$ so that $u, v$ meet the hypotheses of \cite[Lemma 6.2]{Giusti77} which gives the result if $c_1$ is small enough.
\end{proof}

We have thus established the $C^1$ de Giorgi lemma, under the additional assumptions that $P \in N$, and that $N$ is graphical with a graphical function $u$ such that $\|\dif u\|_{C^0} \leq c_1$.
We shall deal with the assumption that $P \in N$ later, but first we remove the assumption that $N$ is graphical with small derivative.
The key point is that this assumption has \emph{no geometric significance}: at a sufficiently fine scale $\rho$, it will hold as long as our coordinates were chosen so that $X$ is the normal vector to $N$ at $P$.

\begin{lemma}\label{hypersurfaces are graphical}
Suppose that $\rho$ is small enough depending on $c_1$.
Let $U$ be a set with $C^1$ boundary in $B(P, \rho)$.
Suppose that $X$ is a vector field which is the coordinate vector field for some normal coordinates $(x^\mu)$ based at $Q \in B(P, \rho)$, such that on $B(P, \rho)$,
\begin{align}
(\normal_U, X) &\geq e^{-o(c_1^2)}. \label{rep as a good graph hyp}
\end{align}
Then there exists a relatively open set $\Omega \subseteq \{x^0 = 0\}$ such that $\partial U$ is $X$-graphical over $\Omega$, and its graphical function $u$ satisfies $\|\dif u\|_{C^0} \leq c_1$.

Moreover, there exists a ball $\mathscr B \subseteq \Omega$ (with respect to the flat metric on $\Omega$) with the following property.
Let $\Omega^\alpha$ be the set of points in $\Omega$ which are initial data for $X$-integral curves which intersect $\partial U \cap B(P, \alpha \rho)$.
Then 
\begin{equation}\label{rep as a good graph set nests}
    \Omega^\alpha \subseteq \left(\alpha + \frac{c_0}{2}\right) \mathscr B \subset \mathscr B \subseteq \Omega.
\end{equation}
\end{lemma}
\begin{proof}
We first observe that if $\rho$ is small enough depending on $M$, then $B(P, \rho)$ appears convex in the coordinates $(x^\mu)$.
Indeed, $(x^\mu)$ is $O(\rho)$-close in the $C^2$ topology to a normal coordinate system $(y^\mu)$ based at $P$, so it suffices to show that $B(P, \rho)$ appears uniformly strictly convex as $\rho \to 0$ in the coordinates $(y^\mu)$.
However, in those coordinates $(y^\mu)$, $\partial B(P, \rho)$ appears to be a sphere of radius $\sim \rho$, so its principal curvatures appear to be $\gtrsim \rho^{-1}$, implying the uniformly strict convexity.

By \cite[Theorem 4.8]{Giusti77}, the convexity in coordinates of $B(P, \rho)$ and (\ref{rep as a good graph hyp}), there exists $\Omega \subseteq \{x^0 = 0\}$ and $u \in C^1(\Omega)$ such that $\partial U$ is the graph of $u$ and 
$$\|\dif u\|_{C^0} \leq \sup_{x_1, x_2 \in \Omega} \frac{|u(x_1) - u(x_2)|}{|x_1 - x_2|} \leq e^{o(c_1^2)}\sqrt{1 - e^{-o(c_1^2)}} \leq c_1.$$

It remains to construct $\mathscr B$ satisfying (\ref{rep as a good graph set nests}); we adapt the proof of \cite[(6.25)]{Giusti77} to this setting.
We may assume that $\Omega^\alpha$ is nonempty; otherwise just choose any ball $\mathscr B$ in the open set $\Omega$.
Let $y := x^0$ and $x := (x_1, \dots, x_{d - 1})$ denote the coordinate functions.
Since $\Omega^\alpha$ is nonempty, there exists $(x_*, y_*) \in \partial U \cap B(P, \alpha \rho)$, and then for $(x, y) \in \partial U$, we obtain from the bound on $\dif u$ that $|y - y_*| \lesssim c_1 \rho$.
Writing $(x_\natural, y_\natural)$ for the coordinates of $P$, we see from the Pythagorean theorem and the approximate euclideanness of the metric that for $(x, y) \in \partial B(P, \rho)$,
$$|x - x_\natural|^2 + (y - y_\natural)^2 = \rho^2 + O(\rho^3).$$
If we additionally impose $(x, y) \in \partial U$, then it follows that 
$$|x - x_\natural|^2 + (y_* - y_\natural)^2 = \rho^2 + O(c_1 \rho^2)$$
at least if $\rho$ is small depending on $c_1$.
But $(x, y) \in \partial U \cap \partial B(P, \rho)$ iff $x \in \partial \Omega$ and $y = u(x)$.
So $\partial \Omega$ lies in an annulus:
$$\partial \Omega \subset \{x: (1 - Cc_1)\rho^2 - (y_* - y_\natural)^2 \leq |x - x_\natural|^2 \leq (1 + Cc_1)\rho^2 - (y_* - y_\natural)^2\}$$
for some constant $C$ depending on $M$ only.
A similar argument with $(x, y) \in \partial U \cap \partial B(P, \alpha \rho)$ (instead of $\partial B(P, \rho)$) shows that
$$\partial \Omega^\alpha \subset \{x: (1 - Cc_1) \alpha^2 \rho^2 - (y_* - y_\natural)^2 \leq |x - x_\natural|^2 \leq (1 + Cc_1)\alpha^2 \rho^2 - (y_* - y_\natural)^2\}.$$
If we take $c_1$ small enough depending on $C, c_0$, then we may choose
$$\mathscr B := \mathscr B\left(x_\natural, \sqrt{(1 - Cc_1)\rho^2 - (y_* - y_\natural)^2}\right).$$
Then $\mathscr B \subseteq \Omega$ and since 
\begin{align*}
\left(\alpha + \frac{c_0}{2}\right) \sqrt{(1 - Cc_1)\rho^2 - (y_* - y_\natural)^2} &\geq \sqrt{(1 + Cc_1)\alpha^2 \rho^2 - (y_* - y_\natural)^2},
\end{align*}
we also have $(\alpha + c_0/2) \mathscr B \supseteq \Omega^\alpha$.
\end{proof}

\begin{proof}[Proof of Proposition \ref{Miranda44}]
Throughout this proof we assume that there exists some $Q \in \partial U \cap B(P, \alpha \rho)$.
If not, then (\ref{Miranda44 concl}) is vacuous since then $\Exc_{\alpha \rho} (U, P) = 0$.

Consider normal coordinates $(x^\mu)$ based at $Q$, such that at $Q$, $Y := \partial_0$ points in the same direction as $X$.
If $\rho$ is small enough depending on $c_1$, then by (\ref{Miranda44 normal hyp}), we have 
$$(\normal_U, Y) \geq e^{-o(c_1^2)}.$$
By Lemma \ref{hypersurfaces are graphical}, it follows that $\partial U$ is $Y$-graphical over some set $\Omega \subseteq \{x^0 = 0\}$, with graphical function $u$ satisfying $\|\dif u\|_{C^0} \leq c_1$.
Moreover, there exists a ball $\mathscr B \subseteq \Omega$ satisfying (\ref{rep as a good graph set nests}).
Let $J_t := \{(x, y) \in B(P, \rho): x \in t\mathscr B\}$ be the cylinder over $t\mathscr B$.

For $t \leq 1$, we can use (\ref{path ordered exponential taylor series}) and Corollary \ref{doubling dimension} to compute in the coordinates $(x^\mu)$ that
\begin{align*}
    \Exc_{J_t}(U, Q) &= |\partial U \cap J_t| - \int_{J_t} \mathcal Pe^{-\int_R^Q \Gamma} \normal_U(R) \dif \mu(R) \\
    &= |\partial U \cap J_t| - \left[\int_{J_t} (\normal_U)_\alpha \dif \mu\right] \dif x^\alpha + O(\rho^{d + 1}).
\end{align*}
Here $\Gamma$ are Christoffel symbols and $\mu$ is the surface measure.

Let $\nu$ denote the surface measure with respect to the euclidean metric on $M$ obtained from coordinates, and let $\normal_U'$ denote the conormal $1$-form with respect to the euclidean metric.
On \cite[pg83]{Giusti77} it is shown that
\begin{equation}\label{Excess versus Lagrangian}
\left|\left[\int_{J_t} (\normal_U')_\alpha \dif \nu\right] \dif x^\alpha\right| = \Japan{\avg_{t\rho} \dif u} |t\mathscr B|.
\end{equation}
By Corollary \ref{doubling dimension} and the approximate euclideanness of the metric, we can replace $\dif \nu$ with $\dif \mu$, and $\normal_U'$ with $\normal_U$, in (\ref{Excess versus Lagrangian}), at the price of an error of size $O(\rho^{d + 1})$.
So from (\ref{approximate by euclidean lagrangian}), we deduce 
\begin{align*}
\Exc_{J_t}(U, P)
&= \Exc_{J_t}(U, Q) + O(\rho^{d + 1}) \\
&= |\partial U \cap J_t| - \int_{t\mathscr B} \Japan{\avg_{t\rho} \dif u} \dif x + O(\rho^{d + 1}) \\
&= \int_{t\mathscr B} \Lagrange(u, \dif u) - \Lagrange(u, \avg_{t\rho} \dif u) + O(\rho^{d + 1}).
\end{align*}
Applying (\ref{rep as a good graph set nests}) and the properties (\ref{approximate monotone}) and (\ref{translation invariant excess}) of the excess, we obtain from Lemma \ref{Miranda43} with $\tilde \alpha := \alpha + c_0/2$ that
\begin{align*}
\Exc_{\alpha \rho}(U, P) 
&\leq \Exc_{J_{\alpha + c_0/2}}(U, P) + O(\rho^{d + 1}) \\
&\leq \int_{(\alpha + c_0/2) \mathscr B} \Lagrange(u, \dif u) - \Lagrange(u, \avg_{(\alpha + c_0)\rho} \dif u) + O(\rho^{d + 1}) \\
&\leq \left(\left(\alpha + \frac{c_0}{2}\right)^{d + 1} + \frac{c_0}{2}\right) \int_{\mathscr B} \Lagrange(u, \dif u) - \Lagrange(u, \avg_\rho \dif u) + O(\rho^{d + 1}).
\end{align*}
Since $\alpha < 1$, if $c_0$ is smaller than an absolute constant, $((\alpha + c_0/2)^{d + 1} + c_0/2) \leq \alpha^{d + 1} + c_0$.
We can then reverse the above computation to obtain 
\begin{align*}(\alpha^{d + 1} + c_0) \int_{\mathscr B} \Lagrange(u, \dif u) - \Lagrange(u, \avg_\rho \dif u) + O(\rho^{d + 1})& \leq (\alpha^{d + 1} + c_0) \Exc_\rho(U, P) + O(\rho^{d + 1}). \qedhere 
\end{align*}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Reduction to the smooth case}
In \S\ref{Mollifiers} we shall establish the following lemma, an analogue of \cite[Lemma 7.5]{Giusti77}:

\begin{lemma}\label{single mollify}
There exists $c_2(c_1, M) > 0$ such that for any ball $B(P, \rho)$, with $\rho$ sufficiently small depending on $c_1$, and any set $U$ of least perimeter in $B(P, \rho)$ such that
\begin{equation}\label{single mollify hyp}
\Exc_\rho (U, P) \leq c_2 \rho^{d - 1},
\end{equation}
there exists an open set $V \subseteq B(P, (1 - c_1)\rho)$ and a vector field $X$ such that $\partial V$ is $C^1$,
\begin{equation}
    |\partial V \cap B(P, (1 - c_1)\rho)| \leq \eta(V, B(P, (1 - c_1)\rho)) + c_1 \Exc_\rho (U, P), \label{single mollify minimality}
\end{equation}
and $X$ is a coordinate vector field for some normal coordinate chart based at $P$ satisfying
\begin{equation}
(\normal_V, X) \geq e^{-o(c_1^2)}. \label{single mollify normal}
\end{equation}
Moreover, for $\rho \lesssim \varpi \leq (1 - c_1) \rho$,
\begin{equation}
|\Exc_\varpi (U, P) - \Exc_\varpi (V, P)| \leq c_1 \Exc_\rho (U, P) + C\rho^{d + 1}. \label{single mollify excess}
\end{equation}
\end{lemma}

\begin{proof}[Proof of the de Giorgi lemma, assuming Lemma \ref{single mollify}]
Choose $c_0 \leq 2^{-(d + 2)}$ small enough to meet the hypotheses of Proposition \ref{Miranda44}.
Let $\alpha := 1/(2(1 - c_0))$, $c_1 = c_1(c_0) \leq c_0/4$ small enough to meet the hypotheses of Proposition \ref{Miranda44} $c := c_2 = c_2(c_0, c_1)$ as in Lemma \ref{single mollify}.

Let $U$ be a set of least perimeter satisfying (\ref{single mollify hyp}).
Let $V, X$ be as in Lemma \ref{single mollify}.
Then by (\ref{approximate monotone}) and (\ref{single mollify excess}),
\begin{align*}
\Exc_{(1 - c_1) \rho} (V, P) &\leq \Exc_{(1 - c_1) \rho} (U, P) + c_1 \Exc_\rho (U, P) + O(\rho^{d + 1}) \\
&\leq (1 + c_1) \Exc_\rho (U, P) + O(\rho^{d + 1}).
\end{align*}
Since $1/2 = \alpha (1 - c_0)$, by Proposition \ref{Miranda44}, it follows that
\begin{align*}
    \Exc_{\rho/2} (V, P) &\leq (2^{-(d + 1)} + c_0) (1 + c_1) \Exc_\rho (U, P) + O(\rho^{d + 1}) \\
    &\leq (2^{-(d + 1)} + 2^{-(d + 2)}) \Exc_\rho (U, P) + O(\rho^{d + 1})
\end{align*}
Finally, by (\ref{approximate monotone}) and (\ref{single mollify excess}),
\begin{align*}
    \Exc_{\rho/2} (U, P)
    &\leq \Exc_{\rho/2} (V, P) + c_1 \Exc_\rho (U, P) + O(\rho^{d + 1}) \\
    &\leq (2^{-(d + 1)} + 2^{-(d + 2)} + 2^{-(d + 3)}) \Exc_\rho (U, P) + O(\rho^{d + 1})\\
    &\leq 2^{-d} \Exc_\rho (U, P) + O(\rho^{d + 1}). \qedhere
\end{align*}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mollification} \label{Mollifiers}
The goal of this lemma is to prove Lemma \ref{single mollify}, which we carry out by convolution with a smoother function.

In order to do this, we need to choose a normal coordinate system $(x^\mu)$ at some ``coordinate origin'' $O \in M$; then convolution $f * g$ is defined in terms of the affine structure and volume form induced by the cordinate system $(x^\mu)$, rather than intrinsically.
In particular, convolution is not even approximately diffeomorphism-invariant, and unlike in \cite{Giusti77} we must take care to ensure that this will not create unacceptable error terms.

Following \cite[Chapter 7]{Giusti77} we define the convolution kernel
$$\chi_\varepsilon(x) := \frac{d + 1}{|\Ball^d|} \varepsilon^{-d}1_{|x| < \varepsilon} \left(1 - \frac{|x|}{\varepsilon}\right)$$
We write $B_\varepsilon := B(O, \varepsilon)$ and $u_\varepsilon := u * \chi_\varepsilon$ whenever $u \in BV(B_{2\varepsilon})$.
When applied to sets of least perimeter with small excess, this convolution operator satisfies the following analogue of \cite[Theorem 7.3]{Giusti77}, which asserts that the convolution has a normal vector which is $C^0$ close to a constant.

\begin{proposition}\label{main mollifier lemma}
Let $q := \min(1/4, 1/(2(d - 1))$.
There exists $c > 0$ such that for every $0 < \rho \lesssim 1$, every $0 < \gamma \lesssim 1$, and every set $U$ of least perimeter such that
\begin{equation}\label{hypothesis on main mollifier lemma}
\Exc_\rho(U, O) \leq \gamma \rho^{d - 1},
\end{equation}
if we let $\varepsilon := \gamma^4\rho$, $\sigma := \gamma^{1/(2(d - 1))}\rho$, and $\varphi := (1_U)_\varepsilon$, then:
\begin{enumerate}
\item For every $y \in (c\gamma^2, 1 - c\gamma^2)$, the level set $\partial \{\varphi > y\} \cap B_{\rho - 2\sigma}$ is a $C^1$ hypersurface.
\item Possibly after applying a rotation around $O$, for every $x \in B_{\rho - 2\sigma}$ such that $c\gamma^2 < \varphi(x) < 1 - c\gamma^2$,
\begin{equation}\label{claim on main mollifier lemma}
\frac{\dif \varphi}{|\dif \varphi|} \geq 1 - O(\gamma^q) - O(\rho^2).
\end{equation}
\end{enumerate}
\end{proposition}

The estimate (\ref{claim on main mollifier lemma}) says that the normal vector to level sets $\{\varphi = y\}$, where $y$ is far from $\{0, 1\}$, is approximately constant in the coordinates $(x^\mu)$.
In particular, if $\gamma^q + \rho^2 \ll c_1$, then the level sets of $\varphi$ satisfy (\ref{Miranda44 normal hyp}).
In the application to the de Giorgi lemma, we can think of $c_1$ as ``order $1$'' (at least compared to the tiny parameters in Proposition \ref{main mollifier lemma}) and $\gamma \sim \rho^2$, in which case
$$1 \gg \gamma^q \gg \sigma \gg \rho \gg \gamma \gg \varepsilon \gg \varepsilon \delta > 0.$$
Since we can take losses of size $O(\rho^2)$, we shall mean by $|\cdot|$ the euclidean norm in the coordinates, and we shall work with the euclidean measure rather than $\star 1$.

The idea behind the proof of Proposition \ref{main mollifier lemma}, as in \cite{Giusti77}, is to cover $\partial^* U$ by small balls and apply the monotonicity formula in each ball.
Our next lemma thus estimates the behavior of $u$ in a single ball.

\begin{lemma}\label{mollifier sublemma}
Let $\delta := \gamma^d$, and let $u := 1_U$ satisfy (in the sense of distributions)
\begin{equation}\label{hypothesis on mollifier sublemma}
\int_{B_\rho} (|\dif u| - \partial_0 u)(z) \dif z \leq \rho^{d - 1} \gamma.
\end{equation}
Then for every $P \in \partial^* U \cap B_\varepsilon$ and $x \in B_{\rho - 2\sigma}$,
$$(1_{B(P, 2\delta\varepsilon)}(|\dif u| - \partial_0 u))_\varepsilon(x) \lesssim \gamma^q (1_{B(P, \delta\varepsilon)} |\dif u|)_\varepsilon(x).$$
\end{lemma}
\begin{proof}
We first claim that for $r > 0$ so small that $B(P, 2r) \subseteq B_\varepsilon$,
\begin{equation}\label{bound the kernel}
\sup_{y \in B(P, 2r)} \chi_\varepsilon(x - y) \lesssim \inf_{y \in B(P, r)} \chi_\varepsilon(x - y).
\end{equation}
In the euclidean case, this result can be isolated from the proof of \cite[Theorem 7.3]{Giusti77}.
Otherwise, we can use the smallness of $\varepsilon$ and the Taylor expansion of the metric to approximate $g$-balls by euclidean balls.
This suffices to prove (\ref{bound the kernel}), since $\chi_\varepsilon$ is uniformly continuous.

Now let $V := B(P, 2\delta\varepsilon)$.
Integrating (\ref{bound the kernel}) against $1_V(|\dif u| - \partial_0)$,
\begin{equation}\label{kernel bounded}
    (1_V(|\dif u| - \partial_0))_\varepsilon(x) \lesssim \inf_{y \in B(P, \delta\varepsilon)} \chi_\varepsilon(x - y) \int_V (|\dif u| - \partial_0 u)(z) \dif z.
\end{equation}
To estimate the right-hand side of (\ref{kernel bounded}), we assume that $\gamma$ is chosen so small that $\sigma > 2\delta\varepsilon$.
We then set $W := B(P, \sigma)$ and apply the weak monotonicity formula (\ref{weak monotonicity}) to get
$$(2\delta\varepsilon)^{1 - d} \int_V |\dif u|(z) \dif z \leq e^{O(|K| \sigma^2)} \sigma^{1 - d} \int_W |\dif u|(z) \dif z$$
and hence
\begin{align*}
(2\delta\varepsilon)^{1 - d} \int_V (|\dif u| - \partial_0 u)(z) \dif z
&\leq \sigma^{1 - d} \int_W (|\dif u| - \partial_0 u)(z) \dif z + O(|K|) \sigma^{3 - d} \int_W |\dif u|(z) \dif z \\
&\qquad + \sigma^{1 - d} \int_W \partial_0 u(z) \dif z - (2\delta\varepsilon)^{1 - d} \int_V \partial_0 u(z) \dif z\\
&=: I_1 + I_2 + I_3 - I_4.
\end{align*}

Before estimating the $I_i$, we next claim that we may assume that
\begin{equation}\label{submollifier gauge assumption}
|\partial_0 - \partial_0^P| \lesssim \gamma^{1/(d - 1)}
\end{equation}
on $W$. Indeed, the statement of Lemma \ref{mollifier sublemma} is independent of the choice of gauge $\Phi^P$, so we can make a gauge transformation, apply Lemma \ref{DoVF lemma}, and recall $O \in W$ (since $\sigma > \varepsilon$) to get $|\partial_0 - \partial_0^P| \lesssim \sigma^2$
which gives (\ref{submollifier gauge assumption}).

We then use (\ref{hypothesis on mollifier sublemma}) to bound
$$I_1 \leq \sigma^{1 - d} \int_{B_\rho} (|\dif u| - \partial_0 u)(z) \dif z \leq \gamma^{\frac{1 - d}{2(d - 1)}} \gamma = \gamma^{1/2}.$$
Also by Corollary \ref{doubling dimension} we have $I_2 \lesssim |K| \gamma^{1/(d - 1)}$.
We can ignore the $K$ since we are allowing constants to blow up as $K \to \infty$.

To estimate $I_3 - I_4$, we recall the notation (\ref{integral of du}) for vector-valued integrals, and apply the monotonicity formula, Proposition \ref{Monotone}, to compute
\begin{align*}
    I_3 - I_4 &= \sigma^{1 - d} I(u, Q, \sigma)_0 - (2 \delta \varepsilon)^{1 - d} I(u, Q, 2\delta\varepsilon)_0 + O(|K| \sigma^2) \\
    &\leq |\sigma^{1 - d} I(u, Q, \sigma) - (2 \delta \varepsilon)^{1 - d} I(u, Q, 2 \delta \varepsilon)| + O(|K| \sigma^2)) \\
    &\lesssim \sqrt{1 + (d - 1) \log \frac{\sigma}{2\delta\varepsilon}} \sqrt{\sigma^{1 - d} \int_W \star |\dif u|} \sqrt{\int_{2\delta\varepsilon}^\sigma \partial_r \left[e^{Ar^2} r^{1 - d} \int_{B(Q, r)} \star |\dif u|\right] \dif r}\\
&\qquad + |K| \sigma^{3 - d} \int_W \star |\dif u| + |K| \sigma^2 \\
&=: J_1 J_2 J_3 + J_4 + J_5.
\end{align*}
Here $0 \leq A \lesssim |K|$ is as in Proposition \ref{Monotone}.
By definition we have $J_1 \lesssim -\log \gamma$, $J_4 = I_1 \lesssim \gamma^{1/(d - 1)}$, and $J_5 \lesssim \gamma^{1/(d - 1)}$.
From Corollary \ref{doubling dimension} we have $J_2 \lesssim 1$.
We then need to bound $J_3$:

\begin{sublemma}
    We have $J_3 \lesssim \gamma^q$.
\end{sublemma}
\begin{proof}
We bound
\begin{align*}
J_3^2 &\leq \sigma^{1 - d} \int_W |\dif u|(z) \dif z - (2 \delta \varepsilon)^{1 - d} \int_V |\dif u|(z) \dif z + O(|K|) \sigma^{3 - d} \int_W \star |\dif u| \\
&= \sigma^{1 - d} \int_W (|\dif u| - \partial_0 u)(z) \dif z + \sigma^{1 - d} \int_W (\partial_0 - \partial_0^P)u(z) \dif z + \sigma^{1 - d} \int_W \partial_0^Pu(z) \dif z \\
  &\qquad - (2 \delta\varepsilon)^{1 - d} \int_V \partial_0^P u(z) \dif z + O(|K|) \sigma^{3 - d} \int_W \star |\dif u| \\
&=: K_1 + K_2 + K_3 - K_4 + K_5.
\end{align*}
Then $K_1 = I_1 \leq \gamma^{1/2}$, $K_2 = I_2 \lesssim \gamma^{1/(d - 1)}$, and $K_5 = J_4 \lesssim \gamma^{1/(d - 1)}$.

To estimate $K_3 - K_4$, we introduce the $d - 1$-form
$$\psi := \dif x^1_P \wedge \cdots \wedge \dif x^{d - 1}_P.$$
Then
\begin{equation}\label{K3 calculus}
K_3 = \sigma^{1 - d} \int_W \dif u \wedge \psi = \sigma^{1 - d} \int_{U \cap \partial W} \psi.
\end{equation}
We decompose
$$\partial W = \Gamma_+ \cup \Gamma_0 \cup \Gamma_-$$
where $\pm x^0_P > 0$ on the hemispheres $\Gamma_\pm$ and $\Gamma_0$ is the equator.
Then all positive contributions to the integral in the right-hand side of (\ref{K3 calculus}) come from $\Gamma_+$.
Moreover, as $d-1$-chains in $M$, $\partial \Gamma_+ = \Gamma_0$.
However, if we set $N := \{x^0_P = 0\}$ and $W_0 := W \cap N$, then $\Gamma_0 = \partial W_0$, and $\Gamma_+$ and $W_0$ have the same homology class relative to $\Gamma_0$.
But $\dif \psi = 0$, so by Stokes' theorem,
$$K_3 \leq \sigma^{1 - d} \int_{\Gamma_+ \cap U} \psi \leq \sigma^{1 - d} \int_{\Gamma_+} \psi = \sigma^{1 - d} \int_{W_0} \psi.$$
Since $\psi$ is the euclidean volume form on $W_0$, and $W_0$ is a $d-1$-ball whose euclidean radius is $\leq \sigma + O(|K| \sigma^3)$, it follows that
\begin{equation}\label{K3 calculus 2}
K_3 \leq |\Ball^{d - 1}| + O(|K| \sigma^2).
\end{equation}
By Corollary \ref{doubling dimension}, the right-hand side of (\ref{K3 calculus 2}) is $\leq K_4 + O(\gamma^{1/(d - 1)})$.
Adding up all the $K_i$ we complete the proof.
\end{proof}

By (\ref{kernel bounded}),
$$(1_V(|\dif u| - \partial_0 u))_\varepsilon(x) \lesssim (\delta\varepsilon)^{d - 1} \gamma^q \inf_{y \in B(P, \delta\varepsilon)} \chi_\varepsilon(x - y).$$
We finally apply Corollary \ref{doubling dimension} to prove
\begin{align*}
(\delta\varepsilon)^{d - 1} \inf_{y \in B(P, \delta\varepsilon)} \chi_\varepsilon(x - y)
&\lesssim \int_{B(P, \delta \varepsilon)} \chi_\varepsilon(x - y) |\dif u|(y) \dif y = |\dif u|_\varepsilon(x). \qedhere
\end{align*}
\end{proof}

\begin{proof}[Proof of Proposition \ref{main mollifier lemma}]
Let $\delta := \gamma^d > 0$ and $u := 1_U$.
We greedily construct a cover $\mathcal V = \{V_n: 1 \leq n \leq N\}$ of $\partial^* U \cap B_{\varepsilon(1 - 2\delta)}$ by balls of radius $2\delta\varepsilon$, centered on points $Q_n \in \partial^* U \cap B_{\varepsilon(1 - \delta)}$, which is \dfn{efficient} in the sense that each $Q \in \partial^*U \cap B_{\varepsilon(1 - \delta)}$ only lies in $O_d(1)$ balls $V_n$.
We set $V_0 := B_\varepsilon \setminus B_{\varepsilon(1 - 2\delta)}$.
Then $\supp \dif u \subseteq \bigcup_n V_n$, so
\begin{equation}\label{sum over cover}
(|\dif u| - \partial_0 u)_\varepsilon \leq \sum_{n = 0}^N (1_{V_n} (|\dif u| - \partial_0 u))_\varepsilon.
\end{equation}
After rotating the coordinate frame we may assume that $\dif u = \partial_0 u \dif x^0$. Then
\begin{equation}\label{excess is d0u}
    \int_{B(P, \rho)} (|\dif u| - \partial_0 u) \dif x = \Exc_\rho(U, P) + O(\rho^{d + 1}) \lesssim \rho^{d - 1}(\gamma + \rho^2).
\end{equation}
Applying (\ref{excess is d0u}) and reasoning similarly to \cite[pg92]{Giusti77}, one can show that there exists $c > 0$ such that
\begin{equation}\label{V0 case}
    (1_{V_0} (|\dif u| - \partial_0 u))_\varepsilon \lesssim (\gamma + \rho^2) |\dif u|_\varepsilon
\end{equation}
on $B_\sigma \cap \{c\gamma^2 < \varphi < 1 - c\gamma^2\}$.
% The proof of (\ref{V0 case}) is essentially given by \cite[pg92]{Giusti77}, so we just sketch it.
% For $y \in V_0$, $\chi_\varepsilon(x - y) \lesssim \delta/\varepsilon^d$, so using Corollary \ref{doubling dimension}, one can show
% $$\int_{V_0} \chi_\varepsilon(x - y)(|\psi| \cdot |\dif u| - \star(\dif u \wedge \psi))(y) \dif y \lesssim \frac{\gamma^d}{\varepsilon}.$$
% Here we used $\|\psi\|_{L^\infty} \lesssim 1$.
% One can use \cite[Lemma 7.1]{Giusti77}, the assumption $c\gamma^2 < \varphi < 1 - c\gamma^2$, and the fact that $g$ is a perturbation of the euclidean metric to obtain
% $$\int_{B_\varepsilon} \chi_\varepsilon(x - y) |\dif u|(y) \dif y \gtrsim \frac{\gamma^{d - 1}}{\varepsilon}$$
% which then implies (\ref{V0 case}).
Since $\mathcal V$ is efficient, we can sum (\ref{V0 case}) and Lemma \ref{mollifier sublemma} (with $\gamma$ replaced by $\gamma + \rho^2$) over $n$ in (\ref{sum over cover}) to show that
$$(|\dif u| - \partial_0 u)_\varepsilon \lesssim (\gamma^q + \rho^2) |\dif u|_\varepsilon$$
which implies (\ref{claim on main mollifier lemma}).
In particular near $\varphi^{-1}(y) \cap B_{\rho - 2\sigma}$, where $y \in (c\gamma^2, 1 - c\gamma^2)$, one has $|\dif u| > 0$.
Therefore $u$ is a $C^1$ submersion by \cite[Lemma 7.1]{Giusti77}, which completes the proof.
\end{proof}

\begin{proof}[Proof of Lemma \ref{single mollify}]
Suppose not.
Then there exist balls $B_n := B(P_n, \rho_n)$ and sets $U_n$ of least perimeter in $B_n$ such that
\begin{equation}\label{single mollify Excess assumption}
\gamma_n := \rho_n^{1 - d} \Exc_{\rho_n} (U_n, P_n) \leq n^{-2},
\end{equation}
and $\rho_n \leq 1/n$, but such that for every open set $V_n \subseteq B(P_n, (1 - \varepsilon) \rho_n)$ with $C^1$ boundary, and every vector field $X_n$ arising from a normal coordinate system at $P_n$, at least one of (\ref{single mollify normal}), (\ref{single mollify minimality}), or (\ref{single mollify excess}) fail.

Now let $w_n := (u_n)_{\gamma_n^4 \rho_n}$ be the mollification of $u_n$ with coordinate origin $O := P_n$, and let $X_n$ be the $\partial_0$ of the normal coordinates at $P_n$.
Draw $t \in [0, 1]$ uniformly at random, and let $c, q > 0$ be as in Proposition \ref{main mollifier lemma} (which we apply with $\gamma := n^{-2}$ and $\rho := \rho_n$), $a_n := c\gamma_n^2$, $b_n := 1 - c\gamma_n^2$.
By the coarea formula, Proposition \ref{Coarea2},
$$\int_{tB_n} \star |\dif w_n| = \int_0^1 |\partial^* \{w_n > y\} \cap tB_n| \dif y \geq \int_{a_n}^{b_n} |\partial^* \{w_n > y\} \cap tB_n| \dif y,$$
so by the mean value theorem, there exists $y_n \in (a_n, b_n)$ such that
$$|\partial^* \{w_n > y_n\} \cap tB_n| \leq \frac{1}{b_n - a_n} \int_{tB_n} \star |\dif w_n|.$$
We then set $V_n := \{w_n > y_n\}$, $v_n := 1_{V_n}$, so that $V_n$ has $C^1$ boundary in $tB_n$ and
\begin{equation}\label{MVT mollifier}
|V_n \cap tB_n| \leq \frac{1}{b_n - a_n} \int_{tB_n} \star |\dif w_n|.
\end{equation}
Since $\grad w_n$ is normal to the level sets of $w_n$, $\normal_{V_n} = \dif w_n/|\dif w_n|$.
So by (\ref{claim on main mollifier lemma}),
$$(\normal_{V_n}, X_n) \geq 1 - O(n^{-2q} + n^{-2}) = 1 - O(n^{-2q}).$$
Thus for $n$ large, $V_n$ satisfies (\ref{single mollify normal}).

Let $\Gamma_n := \partial(tB_n)$.
Arguing completely analogously to the proofs of \cite[(7.23), (7.22)]{Giusti77}, respectively, we see that almost surely,
\begin{align}
\|u_n - v_n\|_{L^1(\Gamma_n)} &\ll \gamma_n \label{trace of vn} \\
|\partial V_n \cap sB_n| &\leq |\partial^* U_n \cap sB_n| + \gamma_n. \label{difference of surface area}
\end{align}
Here we are using the fact that $\sum_n n^{-2} = \pi^2/6$ is finite.
The conjunction of (\ref{trace of vn}), (\ref{difference of surface area}), and (\ref{a priori estimate 1}) implies
\begin{equation}
||\partial^* U_n \cap tB_n| - |\partial V_n \cap tB_n|| \ll \gamma_n, \label{mollifier quant2}
\end{equation}
and the conjunction of (\ref{mollifier quant2}), (\ref{a priori estimate 1}), the fact that $U_n$ has least perimeter, (\ref{single mollify Excess assumption}) and (\ref{trace of vn}) implies (\ref{single mollify minimality}).

To derive a contradiction, we must show that $V_n$ satisfies (\ref{single mollify excess}) for $n$ large.
To this end we fix $\varpi \in [\rho/10, t\rho]$ where $t > 1 - c_1$ can be chosen at random almost surely.
Then
\begin{align*}
    |\Exc_\varpi(U_n, P_n) - \Exc_\varpi(V_n, P_n)|
    &\leq ||\partial^* U_n \cap t B_n| - |\partial V_n \cap t B_n||\\
    &\qquad + \left|\left[\int_{B_\varpi} \star \partial_\mu(u_n - v_n) \right] \dif x^\mu(P_n)\right| + O(\rho_n^{d + 1}) \\
    &=: I_1 + I_2 + I_3.
\end{align*}
By (\ref{mollifier quant2}), $I_1 < c_1 \Exc_\varpi(U_n, P_n)/2$ if $n$ is large, and $I_3$ is irrelevant.
Since $\dif \psi_\mu = 0$, by Stokes' theorem and (\ref{trace of vn}), if $n$ is large then
\begin{align*}
    I_2 &\leq \left|\left[\int_{\partial B_\varpi} \star_\varpi (\normal_{B_\varpi})_\mu (u_n - v_n)\right] \dif x^\mu(P_n)\right| \leq \frac{c_1}{2} \Exc_\varpi(U_n, P_n) + O(\rho_n^{d + 1}).
\end{align*}
This implies (\ref{single mollify excess}) for $n$ large and so contradicts our assumptions.
\end{proof}


\printbibliography

\end{document}