\documentclass[reqno,12pt,letterpaper]{amsart}
\RequirePackage{amsmath,amssymb,amsthm,graphicx,mathrsfs,url}
\RequirePackage[usenames,dvipsnames]{color}
\RequirePackage[colorlinks=true,linkcolor=Red,citecolor=Green]{hyperref}
\RequirePackage{amsxtra}
\usepackage{cancel}
\usepackage{tikz-cd}

\setlength{\textheight}{8.50in} \setlength{\oddsidemargin}{0.00in}
\setlength{\evensidemargin}{0.00in} \setlength{\textwidth}{6.08in}
\setlength{\topmargin}{0.00in} \setlength{\headheight}{0.18in}
\setlength{\marginparwidth}{1.0in}
\setlength{\abovedisplayskip}{0.2in}
\setlength{\belowdisplayskip}{0.2in}
\setlength{\parskip}{0.05in}
\renewcommand{\baselinestretch}{1.10}

\title[Geodesic laminations by minimal currents]{Geodesic laminations by minimal currents}
\author{Aidan Backus}
\date{July 2021}

\newcommand{\NN}{\mathbf{N}}
\newcommand{\ZZ}{\mathbf{Z}}
\newcommand{\QQ}{\mathbf{Q}}
\newcommand{\RR}{\mathbf{R}}
\newcommand{\CC}{\mathbf{C}}
\newcommand{\DD}{\mathbf{D}}
\newcommand{\PP}{\mathbf P}
\newcommand{\MM}{\mathbf M}
\newcommand{\II}{\mathbf I}
\newcommand{\Hyp}{\mathbf H}

\DeclareMathOperator{\card}{card}
\DeclareMathOperator{\cent}{center}
\DeclareMathOperator{\ch}{ch}
\DeclareMathOperator{\codim}{codim}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\Gal}{Gal}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\Jac}{Jac}
\DeclareMathOperator{\Lip}{Lip}
\DeclareMathOperator{\Met}{Met}
\DeclareMathOperator{\id}{id}
\DeclareMathOperator{\rad}{rad}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Radon}{Radon}
\DeclareMathOperator*{\Res}{Res}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\singsupp}{sing~supp}
\DeclareMathOperator{\Spec}{Spec}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\Tan}{Tan}
\newcommand{\tr}{\operatorname{tr}}

\newcommand{\Ric}{\mathrm{Ric}}
\newcommand{\Riem}{\mathrm{Riem}}

\newcommand{\dbar}{\overline \partial}

\DeclareMathOperator{\atanh}{atanh}
\DeclareMathOperator{\csch}{csch}
\DeclareMathOperator{\sech}{sech}

\DeclareMathOperator{\Div}{div}
\DeclareMathOperator{\grad}{grad}
\DeclareMathOperator{\Ell}{Ell}
\DeclareMathOperator{\WF}{WF}

\newcommand{\Hilb}{\mathcal H}
\newcommand{\normal}{\mathbf n}
\newcommand{\vol}{\mathrm{vol}}

\newcommand{\pic}{\vspace{30mm}}
\newcommand{\dfn}[1]{\emph{#1}\index{#1}}

\renewcommand{\Re}{\operatorname{Re}}
\renewcommand{\Im}{\operatorname{Im}}


\newtheorem{theorem}{Theorem}[section]
\newtheorem{badtheorem}[theorem]{``Theorem"}
\newtheorem{prop}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{axiom}[theorem]{Axiom}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumption}[theorem]{Assumption}

\newtheorem{exercise}[theorem]{Discussion topic}
\newtheorem{homework}[theorem]{Homework}
\newtheorem{problem}[theorem]{Problem}

\newtheorem{ack}{Acknowledgements}

\numberwithin{equation}{section}


% Mean
\def\Xint#1{\mathchoice
{\XXint\displaystyle\textstyle{#1}}%
{\XXint\textstyle\scriptstyle{#1}}%
{\XXint\scriptstyle\scriptscriptstyle{#1}}%
{\XXint\scriptscriptstyle\scriptscriptstyle{#1}}%
\!\int}
\def\XXint#1#2#3{{\setbox0=\hbox{$#1{#2#3}{\int}$ }
\vcenter{\hbox{$#2#3$ }}\kern-.6\wd0}}
\def\ddashint{\Xint=}
\def\dashint{\Xint-}

%\usepackage{color}
%\hypersetup{%
%    colorlinks=true, % make the links colored%
%    linkcolor=blue, % color TOC links in blue
%    urlcolor=red, % color URLs in red
%    linktoc=all % 'all' will create links for everything in the TOC
%Ning added hyperlinks to the table of contents 6/17/19
%}

% style=alphabetic
\usepackage[backend=bibtex,maxcitenames=50,maxnames=50]{biblatex}
\addbibresource{topics.bib}
\renewbibmacro{in:}{}
\DeclareFieldFormat{pages}{#1}

\begin{document}
\begin{abstract}
Topics exam, Fall 2021.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents

\section{Introduction}
Let $M$ be an oriented Riemannian manifold of metric $g$ and dimension $d \geq 2$.
Our first result is:

\begin{theorem}\label{main thm}
Let $u: M \to \RR$ be a function of least gradient, $d \leq 7$, and $A_y = \partial \{u > y\}$.
Then $(A_y)_{y \in \RR}$ is a lamination of $u$ by smooth minimal hypersurfaces, which are analytic if $g$ is.
\end{theorem}

We refer to Definition \ref{main definitions} for the definitions.
If $M = \RR^d$, then Theorem \ref{main thm} is essentially a standard result; see \cite[Proposition 3.4]{górny2017planar} for an exposition in that case.
A proof of an analogous result for currents is given by \cite[\S5.3]{federer2014geometric}; our proof uses a similar strategy but is rather ``hands-on" in that it avoids the use of homological integration theory.

In the case of a surface, Theorem \ref{main thm} can be stated in a somewhat stronger form.
One can compare and constrast this result to \cite[Theorem 5.2]{daskalopoulos2020transverse}, which deals with $\infty$-harmonic functions.

\begin{theorem}\label{main crly}
Let $\overline \Sigma$ be a convex surface with boundary and suppose that $u: \Sigma \to M$ is a function of least gradient defined on the interior $\Sigma$ of $\overline \Sigma$.
Then, if $A_y = \partial \{u > y\}$, $(A_y)_{y \in \RR}$ extends to a geodesic lamination of $\overline \Sigma$.
\end{theorem}

As a by-product of Theorem \ref{main thm}, we recover the Riemannian case of \cite[Theorem 5.3.20]{federer2014geometric}:

\begin{corollary}\label{main crly 2}
Every minimal $1$-current on a complete Riemannian surface can be written as the finite sum of disjoint geodesics.
\end{corollary}
%
% The statement of Theorem \ref{main thm} implies a great deal of rigidity for functions of least gradient.
% Let us just consider some of the consequences for the $2$-dimensional model geometries, where geodesic laminations are particularly well understood.
%
% \begin{example}[spherical geometry]
% Let $M = S^2$. Then, up to a rotation, the only functions of least gradient are of the form $u = y_N 1_N + y_S 1_S$ where $N,S$ are the northern and southern hemispheres of $M$ respectively.
% To see this we first argue that $u$ is two-valued; if not, so there are $y_1 < y_2 < y_3$ for which $\{u > y_i\}$ are all sets of positive measure, we can find great circles $C_1,\gamma_2$ so that $C_1$ bounds $\{u > y_3\}$ and $C_2$ bounds $\{u > y_2\}$.
% So $C_1,C_2$ are distinct and therefore intersect, a contradiction.
% So for any $u$ we can find $y_N, y_S$ such that the essential image of $u$ is $\{y_N, y_S\}$ and a great circle bounds $\{u = y_S\}$.
% Thus we may assume that $\{u = y_S\} = S$.
% \end{example}
%
% \begin{example}[euclidean geometry]
% Let $M = \RR^2$. Then, up to a rigid motion, the only functions of least gradient are of the form $u(x) = f(x_1)$ where $f' \geq 0$.
% Indeed, if $\{u > y\}$ is neither null nor conull then $\partial \{u > y\}$ is a line $\ell$, and we may assume that $\ell$ is $\{x_2 = 0\}$.
% If $\{u > z\}$ is also neither null nor conull but $\partial \{u > z\}$ is not parallel to $\{x_2 = 0\}$ then $\partial \{u > z\}$ intersects $\{x_2 = 0\}$ which is a contradiction.
% Thus $\partial_2 u = 0$, and the Euler-Lagrange equation (\ref{EulerLagrange}) simplifies to $f'' = 0$, so $f$ is monotone.
% Thus we may assume that $f$ is increasing.
% \end{example}
%
% \begin{example}[hyperbolic geometry]
% Let $M = \Hyp^2/\Gamma$ be a closed (orientable) hyperbolic surface.
% Then for every function $u$ of least gradient the image of $u$ is countable, say $\{y_j: j \in \NN\}$, and closed geodesics $\gamma_j$ such that $\gamma_j,\gamma_{j+1}$ bound $\{u = y_j\}$.
% To see this we recall that the closed geodesics on $M$ are in bijection with conjugacy classes of $\Gamma$, which is a discrete subgroup of $\mathrm{PSL}(\RR^2)$ and therefore is countable.
% Moreover the superlevel sets $\{u > y\}$ must be bounded by closed geodesics, for if not, they are bounded by geodesics that cluster on themselves, which is absurd.
% \end{example}

The hypothesis $d \leq 7$ in Theorem \ref{main thm} is sharp, owing to the following celebrated theorem \cite[Theorem 6.2.2]{Simons68} \cite[Theorem A]{BOMBIERI1969}.
Recall that a \dfn{minimal cone} $C$ is a cone of least perimeter with vertex at the origin.

\begin{theorem}\label{minimal cones in R8}
The following are equivalent:
\begin{enumerate}
\item $d \leq 7$.
\item The boundary of every minimal cone in $\RR^d$ is $C^1$.
\item The boundary of every minimal cone in $\RR^d$ is a hyperplane.
\end{enumerate}
\end{theorem}

In fact, the minimal cone
\begin{equation}\label{bad minimal cone}
\{(x, y) \in \RR^{4 + 4}: |x| < |y|\}
\end{equation}
is singular at $0$ \cite[Theorem A]{BOMBIERI1969}.
So if $M = \RR^d$ and $d \geq 8$, we can take $u$ to be the indicator function of a singular minimal cone in $\RR^d$.
It would be interesting to modify (\ref{bad minimal cone}) to more general Riemannian manifolds of dimension $8$.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Preliminaries}
\begin{notation}[presheaves]
If $F$ is a presheaf of function spaces, we write $u \in F_l(U)$ to mean that for every $V \Subset U$, $u \in F(V)$.
We write $u \in F_c(U)$ to mean that $u \in F(U)$ and $\supp u \Subset U$.
\end{notation}

\begin{notation}[vector bundles]
Let $E$ be a vector bundle, which we will always assume is normed, with dual $E'$.
If $u,v$ are sections of $E',E$ respectively, we write $(u, v)$ for their fiberwise pairing, which is a function $M \to \RR$.
We write $\langle u, v\rangle$ or $\int_M (u, v)$, for their $L^2$-duality pairing, which is a real number.
If $u$ is a section of $E$, we write $u \prec U$ to mean that $||u||_{L^\infty} \lesssim 1$ and $\supp u \Subset U$.
\end{notation}

\begin{definition}
We write $\mathcal D(U, E) = C^\infty_c(U, E)$ for the space of \dfn{test sections} of $E$.
The topological dual space $\mathcal D'(U, E) = \mathcal D(U, E')'$ is the space of \dfn{generalized sections} of $E$.
Test sections of the tangent bundle are known as \dfn{test fields}.
\end{definition}

\begin{definition}
We say that a section is \dfn{as smooth as possible} if it is smooth, and analytic if $g$ is.
\end{definition}

\begin{notation}[volume forms]
We reserve the letter $d$ for dimension or exterior differentiation, and so to avoid awkwardness such as $\int |du| dV$ we write $\vol$ for the Riemannian volume form.
We write $\vol'$ for the euclidean volume form on $\RR^d$ whenever it is relevant.
If $N$ is a closed submanifold we write $\vol_N$ to indicate the pullback of $\vol$ along the inclusion map $M \to N$.
\end{notation}

\begin{definition}
By an \dfn{$E$-valued Radon measure} we mean a generalized section of $E$ which is bounded on $L^\infty_l(U, E')$.
The \dfn{total variation} of an $E$-valued Radon measure $u$ on $U$ is $\sup_{f \prec U} \langle u, f\rangle$.
\end{definition}

\begin{definition}
A function $u \in L^1_l(M)$ has \dfn{bounded variation} if its distributional exterior derivative $du$ is a $T'M$-valued Radon measure of finite total variation.
We write $u \in BV(U)$ to mean that $u$ has bounded variation on $U$.
If $U$ is an open set and $1_U \in BV_l(M)$, we call $U$ a \dfn{Caccioppoli set}.
If $u \in BV_l(M)$ we will write $\int_M |du| ~\vol$ for the total variation of its derivative, to emphasize that the total variation depends on the volume form\footnote{Note carefully
that $\int_N |du| ~\vol$ can be nonzero even if $N$ is null (say, a hypersurface); in general we will be interested in functions which are \emph{not} absolutely continuous.}.
\end{definition}

It follows from the definitions that $|du| ~\vol$ is a generalized volume form and $\int_M |du| ~\vol$ is well-defined.
However, we will also have occasion to consider $\int_M du ~\vol$, which is \emph{not} well-defined.
In fact, if $x$ is the standard coordinate on $M = \RR^d$ (and $g$ is a metric on $\RR^d$), then $dx^j$ can just be viewed as a covector in $(\RR^d)'$ rather than a $1$-form on $\RR^d$.
In that case $du = \sum_j \partial_ju ~dx^j$ and we \emph{define}
$$\int_{\RR^d} du ~\vol = \sum_j \left(\int_{\RR^d} \partial_ju ~\vol \right) dx^j \in (\RR^d)'.$$
Thus we have the triangle inequality
\begin{equation}\label{connection triangle inequality}
\left|\int_{\RR^d} du ~\vol\right| \leq \int_{\RR^d} |du| ~\vol
\end{equation}
since if $du$ has compact support then
$$\sup_{v \in \RR^d} \left(\int_{\RR^d} du ~\vol, v\right) \leq \sup_{v \in \RR^d} \int_{\RR^d} (du, v) ~\vol \leq \sup_{X \prec \RR^d} \int_{\RR^d} (du, X) ~\vol.$$
The important point is that $\int_M du ~\vol$ will only ever be mentioned along with some fixed coordinate system.

Sequences $(u_n)$ in $BV_l(M)$ with $u_n \to u$ in $L^1_l(M)$ satisfy the lower semicontinuity property
\begin{equation}
\label{RieszMarkovDistr}
\int_M |du| ~\vol \leq \liminf_{n \to \infty} \int_M |du_n| ~\vol.
\end{equation}
which follows by testing against smooth functions.
Moreover, the forgetful map
\begin{equation}\label{Forget}
BV_l(M) \to L^1_l(M)
\end{equation}
is compact. We refer to \cite[Chapter 1]{Giusti77} for more on functions of bounded variation.

The main objects of study in this paper are functions of least gradient:

\begin{definition}\label{main definitions}
A function $u \in BV_l(M)$ has \dfn{least gradient} if for every $v \in BV_c(M)$ and $\supp v \subseteq U \Subset M$,
$$\int_U |du| ~\vol \leq \int_U |du + dv| ~\vol.$$
A set $U$ has \dfn{least perimeter} if $1_U$ has least gradient.
\end{definition}

Functions of least gradient can be viewed as the correct notion of weak solution to the Dirichlet problem for the Euler-Lagrange equation
\begin{equation}\label{EulerLagrange}
\Div \frac{\grad u}{|\grad u|} = 0.
\end{equation}
In particular, if $U$ has least perimeter, $u = 1_U$, and $U$ has enough regularity one can view $\grad u/|\grad u|$ as the unit normal field to $U$, so one may view the Euler-Lagrange equation as meaning that $\partial U$ has no mean curvature.
However, by Theorem \ref{minimal cones in R8}, it is impossible to define the mean curvature of $\partial U$ in general.

We recall that if $U$ has least perimeter and can be expressed as the graph of a $C^1$ function $f$, then $f$ solves the minimal surface equation (MSE), which is by definition the Euler-Lagrange equation with Lagrangian density
\begin{equation}\label{Lagrangian density}
\mathscr L = \sqrt{1 + |df|^2} ~\vol.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Strategy of the proof}

Let us recall the strategy of the proof of Theorem \ref{main thm} in case $M = \RR^d$\footnote{For a much more detailed exposition, see \cite{Giusti77}.}.
One first reduces to the case that $u$ is the indicator function of a set whose boundary $N$ has least perimeter.
Blowing up $N$ near a point creates a sequence $(N_j)$ of minimal hypersurfaces converging to the boundary of a minimal cone $C$, which necessarily must be a hyperplane by Theorem \ref{minimal cones in R8}.
In particular, $C$ has a normal vector.
One then uses the de Giorgi lemma \cite{deGiorgi61} to show that if $j$ is large enough then $N_j$ also has a normal vector, but $N_j$ was just a rescaling of $N$, so $N$ has a normal vector and $C$ is the tangent space to $N$.
This implies that $N$ is $C^1$, and one may now use standard elliptic regularity theory to conclude the claim.

This strategy must be modified somewhat in the presence of curvature.
The bulk of the work is proving a suitable de Giorgi lemma, and in fact we show:

\begin{proposition}[Proposition \ref{induction 2}, imprecise]\label{imprecise induction 2}
Let $U \subseteq \RR^d$ be a set of least perimeter with respect to a metric $g$ on $\RR^d$ such that lines through the origin are geodesics, and let
$$\Lambda(U, \rho) = \int_{B(p, \rho)} |d1_U| ~\vol - \left|\int_{B(p, \rho)} d1_U ~\vol\right|.$$
Then if $\Lambda(U, \rho)\rho^{1 - d}$ and $\alpha$ are small enough,
$$\Lambda(U, \alpha \rho) < \alpha^d \Lambda(U, \rho).$$
\end{proposition}

There are two obstructions to a proof of Proposition \ref{imprecise induction 2} which were not present in the euclidean case:
\begin{enumerate}
\item The lack of a sufficiently strong monotonicity formula.
\item The lack of scale-invariance of the volume form (and hence the notion of least perimeter).
\end{enumerate}
Both of these essentially amount to a lack of scale-invariance.
To deal with the first two issues we identify a scale-invariant quantity $\Gamma$; roughly $1/\Gamma$ measures ``how difficult it is to approximate a minimal hypersurface by a $C^1$ hypersurface".


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Outline of the paper}
We begin with the preliminaries in Section \ref{RiemMeasureThy}, which records basic results on functions of approximately least gradient, including a generalization of Miranda's theorem \cite[Teorema 3]{Miranda67} on the stability of functions of least gradient.
We also show that the reduced boundary of a Caccioppoli set is metric-independent.

We are then ready to prove Theorems \ref{main thm} and \ref{main crly}.
In Section \ref{inequalities} we prove a monotonicity formula and an isoperimetric inequality, which show that the blowup of a set of least perimeter is as smooth as possible.
Then, in Section \ref{de Giorgi section}, we prove Proposition \ref{imprecise induction 2}.
In Section \ref{proof of main thm} we combine Proposition \ref{imprecise induction 2} with the results of Section \ref{inequalities} to prove Theorems \ref{main thm} and \ref{main crly}.

In Appendix \ref{coarea section} we deduce a coarea formula that is used throughout the paper, and in Appendix \ref{chains of geodesics} we recall facts about $1$-currents which show that Theorem \ref{main crly} implies Corollary \ref{main crly 2}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Acknowledgements}
I would like to thank Georgios Daskalopoulos for suggesting this project and for many helpful discussions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Riemannian measure theory}\label{RiemMeasureThy}
Let us now review several standard results in geometric measure theory, generalized to the case of Riemannian manifolds.

\subsection{Trace theorem}
We begin with a trace theorem and its consequences.

\begin{proposition}[trace theorem]\label{traces}
Let $U$ be an open set such that $N = \partial U$ is a Lipschitz hypersurface.
For every $u \in BV_l(M)$ there exists a trace $v \in L^1_l(N)$ such that for every test field $X$ on $M$,
\begin{equation}\label{Miranda IBP}
\int_U (du, X) ~\vol + \int_U u ~\mathcal L_XdV = \int_N vg(X, \nu) ~dA.
\end{equation}
Moreover, if $W$ is a small neighborhood of $\partial U$ in $U$, then $v$ is already determined by $u|W$, and $||v||_{L^\infty(N)} \leq ||u||_{L^\infty(U)}$.
\end{proposition}
\begin{proof}
By applying a partition of unity, we may work in coordinates $(x_1, x')$ such that $N$ is a graph in the sense that for every $c \in \RR$, $N \cap \{x_1 = c\}$ is a single point.
One can then imitate the proof of \cite[Teorema 1]{Miranda67}, replacing any integrations by parts with its analogue
\begin{equation}
\label{Riemannian IBP}
\int_U (du, X) ~dV + \int_U u ~\mathcal L_XdV = \int_{\partial U} ug(X, \nu) ~dA
\end{equation}
The construction in \cite[Teorema 1]{Miranda67} shows that only the values of $u$ near $N$ affect $v$ and gives the bound in $L^\infty$.
%
%
%Let $\varphi$ be a smooth flow on $M$ such that for every $x \in N$ and $t \in [0, 1]$, $\varphi_t(x) \in U$.
%Let $N_t$ be the set obtained by flowing $N$ along $\varphi$ for time $t$.
%Then, as long as $\varphi$ is chosen to flow slowly enough (so as to avoid self-intersections), $N_t$ is a Lipschitz hypersurface for every $t \in [0, 1]$.
%Moreover, $N_t$ partitions $U$ into a closed outer region $\{\varphi_s(x): x \in \partial U, s \in [0, t]\}$ and an open inner region $U_t$.
%
%There exist smooth functions $u_n$ on $U$ such that $u_n \to u$ locally uniformly.
%Applying (\ref{Riemannian IBP}) to $u_n$ on $U_t$ and using the locally uniform convergence, we have
%\begin{equation}
%\label{interior Miranda IBP}
%\langle 1_{U_t}du, X\rangle + \int_{U_t} u ~\mathcal L_XdV = \int_{N_t} g(uX, \nu) ~dA.
%\end{equation}
%By continuity from below, the left-hand side of (\ref{interior Miranda IBP}) converges to the left-hand side of (\ref{Miranda IBP}) as $t \to 0$.
%
%If we set $v_t(x) = u(\varphi_t(x))$, then $v_t(x)$ is defined for $x \in \partial U$ and $t \in [0, 1]$.
%Moreover, for every $0 < t_1 < t_2$,
%\begin{equation}
%\label{mean-value inequality}
%\int_N |v_{t_1} - v_{t_2}| ~dA \leq |du|_{TV}(U_{t_2} \setminus \overline{U_{t_1}}).
%\end{equation}
%Indeed, (\ref{mean-value inequality}) holds if $u$ is smooth by the mean-value theorem, and for the general case we can find $u_n$ smooth such that $u_n \to u$ in $L^2_l$.
%Expanding out both sides of (\ref{mean-value inequality}),
%\begin{equation}
%\label{approx mean-value inequality}
%\int_N |(\varphi_{t_1})^* u_n - (\varphi_{t_2})^* u_n| ~dA \leq -\inf_{X \prec U_{t_2} \setminus \overline{U_{t_1}}} \int_M u_n\mathcal L_XdV.
%\end{equation}
%Both sides of (\ref{approx mean-value inequality}) are preserved as $n \to \infty$, due to the convergence in $L^2_l$.
%Therefore (\ref{mean-value inequality}) holds.
%
%Since $|du|_{TV}(U) < \infty$, measure continuity and (\ref{mean-value inequality}) implies that
%$$\lim_{t_1, t_2 \to 0} \int_N |v_{t_1} - v_{t_2}| ~dA = 0.$$
%Thus $(v_t)$ is Cauchy in $L^1(N)$ as $t \to 0$, say $v_t \to v$.
%As $g(X, \nu) \in L^\infty$ and $(L^1)^* = L^\infty$, $g(v_tX, \nu) \to g(vX, \nu)$ in $L^1(N)$.
%Therefore the right-hand side of (\ref{interior Miranda IBP}) satisfies
%$$\lim_{t \to 0} \int_{N_t} g(uX, \nu) ~dA = \lim_{t \to 0} \int_{\partial U} g(v_tX, \nu) ~dA \int_U g(vX, \nu) ~dA$$
%which completes the proof.
\end{proof}

In the sequel we will usually just write $u$ for the trace of $u$, or $u'$ if distinguishing the two functions is crucial.

\begin{proposition}[gluing $BV$ functions along a hypersurface]\label{gluing}
Let $N$ be a Lipschitz hypersurface which separates $M$ into $U_1,U_2$.
If $u_j \in BV(U_j)$ and $u \in L^1_l(M)$ is the function such that $u|U_j = u_j$, then $u \in BV(M)$.
Moreover,
\begin{equation}
\label{glued BV norm}
\int_N |du| ~\vol = \int_N |u_1 - u_2| ~\vol_N.
\end{equation}
\end{proposition}
\begin{proof}
In the proof of \cite[Teorema 2]{Miranda67}, replace \cite[Teorema 1]{Miranda67} with Proposition \ref{traces}.
\end{proof}

\begin{notation}
If $u \in BV(M)$ and $U \Subset M$, we write
$$\eta(u, U) = \inf_{v \prec U} \int_U |d(u+v)| ~\vol$$
so that $u$ has least gradient iff $\eta(u, U) = \int_U |du| ~\vol$ for every $U$.
\end{notation}

\begin{lemma}[a priori estimates]\label{estimates on good set}
Let $u, v \in BV(M)$, let $U \Subset M$ have a Lipschitz boundary $N$. Then
\begin{equation}\label{a priori estimate 1}
|\eta(u, U) - \eta(v, U)| \leq \int_N |u - v| ~\vol_N.
\end{equation}
In particular
\begin{equation}\label{a priori estimate 2}
\eta(u, U) \leq \int_N |u| ~\vol_N \leq ||u||_{L^\infty(N)} |N|.
\end{equation}
\end{lemma}
\begin{proof}
In the proof of \cite[Lemma 5.6]{Giusti77}, replace \cite[Proposition 2.8]{Giusti77} with Proposition \ref{gluing} to prove (\ref{a priori estimate 1}).
Then (\ref{a priori estimate 2}) follows by taking $v = 0$.
\end{proof}

\subsection{The Miranda stability theorem}\label{MirandaStability}
In this section we show a semicontinuity theorem for the Lagrangian (\ref{Lagrangian density}).

The exponential pullback $\exp_p^* u$ of a function $u$ of least gradient defined near $p \in M$ need not have least gradient.
However, in a small ball $B$ around $p$, we will be able to show that $\eta(u, B) \approx |du|_{TV}(B)$ in a sense to be made precise later.
This observation motivates the following definition.

\begin{definition}
A sequence $(u_n)$ of functions in $BV(M)$ has \dfn{approximately least gradient} if
$$\limsup_{n \to \infty} \int_U |du_n| ~\vol \leq \liminf_{n \to \infty} \eta(u_n, U)$$
uniformly as $U$ ranges over open sets $\Subset M$.
\end{definition}

\begin{definition}
Let $(u_n)$ be a sequence in $BV_l(M)$ which converges in $L^1_l$ to $u$.
We say that a Lipschitz hypersurface $N$ \dfn{has no singularities} of $(u_n)$ if:
\begin{enumerate}
\item \label{cond1Mir} $\sup_n \int_N |du_n| ~\vol = 0$.
\item \label{cond2Mir} $(u_n)$ is bounded in $L^1(N, \vol_N)$.
\item \label{cond3Mir} $\int_N |du| ~\vol = 0$.
\item \label{cond4Mir} $u_n \to u$ in $L^1(N, \vol_N)$.
\end{enumerate}
We say that $N$ \dfn{has no singularities} of $u \in BV_l(M)$ if $N$ has no singularities of the sequence $u_n = u$.
By Condition $k$ we mean the $k$th bullet in the above list.
\end{definition}

\begin{lemma}\label{probabilistic method}
Let $(u_n)$ be a sequence in $BV_l(M)$ which converges in $L^1_l(U)$. Then:
\begin{enumerate}
\item \label{probabilistic balls} For every $x \in M$ and $R > 0$ such that $B(x, R) \Subset M$ and almost every $r \in (0, R]$, $\partial B(x, r)$ has no singularities of $(u_n)$.
\item \label{probabilistic hypersurfaces} For every $U \Subset M$ there exists $U \subseteq V \Subset M$ such that $\partial V$ has no singularities of $(u_n)$.
\end{enumerate}
\end{lemma}
\begin{proof}
We first prove (\ref{probabilistic balls}).
Let $r$ be drawn from $[R/2, R]$ uniformly at random; we claim that almost surely, $\partial B(x, r)$ has no singularities of $(u_n)$.
Let
$$A = \{s > 0: \int_{\partial B(x, s)} |du| ~\vol > 0\}.$$
Then
$$\sum_{s \in A} \int_{\partial B(x, s)} |du| ~\vol \leq \int_{\partial B(x, R)} |du| ~\vol < \infty$$
since $|du|$ is a Radon measure and $B(x, R) \Subset M$.
Therefore $A$ is countable,
%Let $A_n = \{|du_n|_{TV}(N) > 0\}$ and let $A_\infty = \{|du|_{TV}(N) > 0\}$.
%Then for every $n \in \NN \cup \{\infty\}$, writing $u_\infty = u$,
%$$\sum_{s \in A_n} |du_n|_{TV}(\partial B(x, s)) \leq |du_n|_{TV}(B(x, R)) < \infty$$
%since $|du_n|_{TV}$ is a Radon measure and $B(x, R) \Subset M$.
%Since each of the summands is nonzero by definition of $A_n$, it follows that $A_n$ is countable, and in particular null.
%Therefore Conditions \ref{cond1Mir} and \ref{cond3Mir} hold almost surely.
so Condition \ref{cond3Mir} holds almost surely.
We omit the proof that the other conditions hold almost surely as it is similar.

To prove (\ref{probabilistic hypersurfaces}), let $U \Subset W \Subset M$, and for every $x \in \partial U$, let $R_x \in (0, d(x, \partial W))$.
Then, by (\ref{probabilistic balls}), for every $x \in \partial U$, there exists $r_x \in (0, R_x)$ such that $\partial B(x, r_x)$ has no singularities of $(u_n)$.
Let $\mathcal U$ be the open cover of $\overline U$ given by the balls $B(x, r_x)$, as well as $U$ itself.
Since $\overline U$ is compact, there exists a finite subcover $\mathcal U_0$ of $\mathcal U$.
Let $V$ be the union of the sets in $\mathcal U_0$.
Then $\partial V$ is a union of finitely many spheres $\partial B(x, r_x)$ which have no singularities, and therefore has no singularities.
\end{proof}

We recall that $BV_l(M)$ is not separable, so it will be useful to have a somewhat weaker topology on $BV_l(M)$, as follows:

\begin{definition}
A sequence of functions $(u_n)$ in $BV_l(M)$ converges \dfn{in total variation on sets with no singularities} to $u \in BV_l(M)$ if for every set $A \Subset M$ such that $\partial A$ has no singularities,
\begin{equation}\label{convergence in TV}
\lim_{n \to \infty} \int_A |du_n| ~\vol = \int_A |du| ~\vol.
\end{equation}
\end{definition}

\begin{proposition}[Miranda stability theorem]\label{Miranda convergence}
If a sequence of functions $(u_n)$ has approximately least gradient and converges in $L^1_l$, then its limit $u$ has least gradient, and $u_n \to u$ in total variation on sets with no singularities.
\end{proposition}
\begin{proof}
By Lemma \ref{probabilistic method} for every $U$ open $\Subset M$ we can find $U \subseteq V \Subset M$ such that $V$ is open and $\partial V$ has no singularities.

We first prove that $u \in BV_l(M)$.
Let $v_n = (1 - 1_U)u_n$, so $v_n \in BV_l(M)$ by Proposition \ref{gluing}.
Since $(u_n)$ has approximately least gradient, if $n$ is large enough then
$$\int_{\overline V} |du_n| ~\vol \leq \eta(u_n, \overline V) + 1 \leq \int_{\overline V} |du_n| ~\vol + 1.$$
So by Proposition \ref{gluing} and Condition \ref{cond1Mir},
$$\int_U |du_n| ~\vol \leq \int_{\overline V} |du_n| ~\vol + 1 = \int_{\partial V} |u_n| ~\vol_{\partial V} + 1.$$
Thus, by (\ref{RieszMarkovDistr}) and Condition \ref{cond2Mir},
$$\int_U |du| ~\vol \leq \limsup_{n \to \infty} \int_U |du_n| ~\vol \leq \limsup_{n \to \infty} \int_{\partial V} |u_n| ~\vol_{\partial V} + 1 < \infty.$$
Therefore $u \in BV_l(M)$.

Let $v$ be a perturbation of $u$, thus $v \in BV_l(M)$ and $u = v$ on $M \setminus U$.
Such a perturbation exists, since $u \in BV_l(M)$.
Let
$$v_n(x) = \begin{cases}
v(x), &x \in V\\
u_n(x), &x \notin V
\end{cases}.$$
By Proposition \ref{gluing} and Condition \ref{cond3Mir}, $v_n \in BV_l$ and
\begin{equation}\label{gluing vn}
\int_{\overline V} |dv_n| ~\vol = \int_V |dv| ~\vol + \int_{\partial V} |u - u_n| ~\vol_{\partial V}.
\end{equation}
Let $\varepsilon > 0$. Then for $n$ large enough, since $(u_n)$ has approximately least gradient and $u_n - v_n$ is trace-free,
$$\int_V |du_n| ~\vol \leq \int_{\overline V} |du_n| ~\vol \leq \eta(u_n, \overline V) + \varepsilon \leq \int_{\overline V} |dv_n| ~\vol + \varepsilon.$$
By (\ref{gluing vn}), it follows that
$$\int_V |du_n| ~\vol \leq \int_V |dv| ~\vol + \int_{\partial V} |u - u_n| ~\vol_{\partial V} + \varepsilon.$$
By Condition \ref{cond4Mir} and (\ref{RieszMarkovDistr}),
$$\int_V |du| ~\vol \leq \int_V |dv| ~\vol + \varepsilon.$$
Since $\varepsilon > 0$ and $u = v$ on $M \setminus U$, it follows that $\int_U |du| ~\vol \leq \int_U |dv| ~\vol + \varepsilon$.
Taking $\varepsilon \to 0$, it follows that $u$ has least gradient.

Finally we prove (\ref{convergence in TV}).
Let $(u_{n_\ell})_\ell$ be a subsequence of $(u_n)$ such that $\int_A |du_{n_\ell}| ~\vol$ is Cauchy in $\ell$.
By Lemma \ref{estimates on good set} and Condition \ref{cond4Mir}, we have
$$\lim_{n \to \infty} |\eta(u, A) - \eta(u_n, A)| \leq \lim_{n \to \infty} \int_{\partial A} |u - u_n| ~\vol_{\partial A} = 0.$$
Since $(u_n)$ has approximately least gradient, for every $\varepsilon > 0$ and $n$ large enough, $\int_A |du_n| ~\vol \leq \eta(u_n, A) + \varepsilon$.
Thus by (\ref{RieszMarkovDistr}) and the fact that $(\int_A |du_{n_\ell}| ~\vol)_\ell$ is Cauchy,
\begin{align*}
\int_A |du| ~\vol &\leq \lim_{\ell \to \infty} \int_A |du_{n_\ell}| ~\vol \leq \lim_{\ell \to \infty} \eta(u_{n_\ell}, A) + \varepsilon\\
&= \eta(u, A) + \varepsilon = \int_A |du| ~\vol + \varepsilon
\end{align*}
where the last equality follows because $u$ has least gradient.
Since $\varepsilon$ and the subsequence $(u_{n_\ell})$ were arbitrary, the inequalities collapse to give (\ref{convergence in TV}).
\end{proof}

\begin{corollary}\label{level sets are minimal}
For every $u$ of least gradient, the superlevel sets $\{u > t\}$ have least perimeter.
\end{corollary}
\begin{proof}
In the proof of \cite[Theorem 1]{BOMBIERI1969}, replace \cite[Theorem 1.6]{Miranda66} with Proposition \ref{Coarea2} and replace \cite[Theorem 3]{Miranda67} with Proposition \ref{Miranda convergence}.
\end{proof}

\begin{corollary}\label{compactness}
Let $(u_n)$ be a sequence of approximately least gradient which is uniformly bounded in $L^\infty$.
Then there is a subsequence of $(u_n)$ which converges in $L^1_l$, almost everywhere, and in total variation on sets with no singularities to a function $u$ of least gradient.

In particular, if $(u_n)$ is a sequence of indicator functions of Caccioppoli sets, then $u$ is an indicator function of a set of least perimeter.
\end{corollary}
\begin{proof}
If $n$ is large enough, then by Proposition \ref{traces}, for every $U \Subset M$,
$$\int_U |du_n| ~\vol \leq \eta(u_n, U) + 1 \leq |\partial U|\cdot ||u_n||_{L^\infty(M)} + 1$$
which gives a uniform bound in $BV_l$.
Since the forgetful map (\ref{Forget}) is compact, a subsequence of $(u_n)$ converges to a function $u$ in $L^1_l$.
By Proposition \ref{Miranda convergence}, $u$ has least gradient and (\ref{convergence in TV}) holds.
By taking a further subsequence we can guarantee the convergence pointwise almost everywhere.
The convergence almost everywhere implies that if $(u_n)$ is a sequence of indicator functions, then there is a Caccioppoli set $U$ such that $u = 1_U$, which necessarily has least perimeter.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Change of metric}\label{change of metric}
Let us investigate how the volume form depends on the choice of metric.
If $\vol$ is the volume form of $g$, $\vol'$ the flat volume form induced by exponential normal coordinates $x$ centered at $p$, we have
the Taylor expansion \cite[p59]{chow2006hamilton}
\begin{equation}\label{Taylor expansion of determinant}
\frac{\vol}{\vol'} = 1 - \frac{1}{3} \Ric_p(x, x) - \frac{1}{6} \nabla \Ric(x, x, x) + a_4x^4 + \cdots
\end{equation}
where $a_j$, $j \geq 4$, are functions of the Riemann tensor and its covariant derivatives.

Let $g_1,g_2$ be metrics defined near a point $p$ and let $\beta = (e_1, \dots, e_d)$ be an orthonormal basis for $T_pM$ with respect to $(g_1)_p$ and $(g_2)_p$.
If $\vol_i$, resp. $\Ric^{(i)}$, is the volume form, resp. Ricci tensor, associated to $g_i$, in the geodesic normal coordinate $x$ associated to $\beta$ in a small ball $B(p, \rho)$, then by (\ref{Taylor expansion of determinant}),
\begin{equation}\label{change of volume form}
\frac{\vol_2}{\vol_1} = \frac{1 - \Ric_p^{(2)}(x, x)/6 + O(x^3)}{1 - \Ric_p^{(1)}(x, x)/6 + O(x^3)} = 1 + O(\rho^2).
\end{equation}

% \begin{proof}
% Recall that if $\omega = de_1 \wedge \cdots \wedge de_d$ then by (\ref{Taylor expansion of determinant}),
% $$\frac{\vol_1}{\omega} = 1 - \frac{1}{6}\Ric_p^{(1)}(x, x) + O(x^3)$$
% while if $y = Ax$ is the geodesic normal coordinate associated to $A_* \beta$ and $\Omega = d(Ae_1) \wedge \cdots \wedge d(Ae_d)$,
% \begin{align*}\frac{\vol_2}{\omega} &= \frac{\vol_2}{\Omega} \frac{\Omega}{\omega} = \left(1 - \frac{1}{6}\Ric_p^{2}(y, y) + O(y^3)\right)\det A \qedhere.\end{align*}
% \end{proof}
% In particular, if $\omega$ is a generalized $1$-form or generalized density on $B(p, \rho)$, writing
% $$\omega_j = (\omega, \grad^{(2)} y_j) = g_2(\omega_j, dy_j)$$
% we have from (\ref{change of volume form})
% \begin{align*}
% \int_{B(p, \rho)} \omega ~dV_2 &= \sum_j \left(\int_{B(p, \rho)} \omega_j ~dV_2\right) ~dy_j\\
% &= \sum_j \left(\int_{B(p, \rho)} \omega_j \frac{dV_2}{dV_1} ~dV_1\right) ~dy_j\\
% &= (1 + O(\rho^2)) \det A \sum_j \left(\int_{B(p, \rho)} \omega_j ~dV_1\right) ~dy_j\\
% &= (1 + O(\rho^2)) \det A \int_{B(p, \rho)} \omega ~dV_1.
% \end{align*}

\begin{notation}
We write $T^kM$ for the $k$th exterior power of $T_pM$ if $k \geq 0$ and $T^{-k}M$ for the $k$th exterior power of $T'_pM$ if $k \leq 0$.
\end{notation}

Let $\zeta_0$ be the injectivity radius of
$$\exp_p: T_pM \to M.$$
Let $a_j \in T^{-j}_pM$ be the Taylor coefficients in (\ref{Taylor expansion of determinant}) and let $k \geq 2$ be the least index $j \geq 1$ such that $a_j \neq 0$.
Then let $c = 2|a_k|$.
Since on $B(p, 1)$, $c|x|^k \leq c|x|^2$ and so
$$1 - \frac{c}{2}|x|^2 - O(x^{k+1}) \leq \frac{\vol}{\vol'} \leq 1 + \frac{c}{2}|x|^2 + O(x^{k+1}).$$
We select $\zeta < \zeta_0$ so small that on $B(p, \zeta) \cap B(p, 1)$, the term $O(x^{k+1})$ is at most $c/2$, so that
\begin{equation}\label{definition of c-zeta}
0.5 < 1 - c|x|^2 \leq \frac{\vol}{\vol'} \leq 1 + c|x|^2 < 2.
\end{equation}

By (\ref{Taylor expansion of determinant}), in the generic case $k = 2$, $c = |\Ric_p|/3$, which motivates the following definition:

\begin{definition}
Let $\zeta,c$ be as in (\ref{definition of c-zeta}).
We call $\zeta$ the \dfn{strong injectivity radius} of $p$ and $c$ the \dfn{Ricci-Taylor error}.
\end{definition}

Let us see how (\ref{definition of c-zeta}) behaves under a rescaling.
We rescale $g$ as $\widetilde g = \lambda g$, $\lambda > 0$, which defines a volume form $\widetilde \vol$.
We also rescale the flat metric suitably to get a volume form $\widetilde \vol'$.
Writing $F = \vol/\vol'$ and writing $F \approx 1 \pm c|x|^2$ on $B(p, \zeta)$ to indicate the relation (\ref{definition of c-zeta}), $\widetilde F = \widetilde \vol/\widetilde \vol'$, we deduce
$$\widetilde F = F \approx 1 \pm c|x|^2$$
on $B(p, \zeta) = \widetilde B(p, \widetilde \zeta)$ where $\widetilde \zeta = \lambda^{0.5} \zeta$ is given by the rescaling of the Riemannian distance function.
Therefore
\begin{equation}\label{scale invariance}
\text{upon applying the scaling above, }\widetilde \zeta = \lambda^{0.5} \zeta, ~\tilde c = c.
\end{equation}

Now we use (\ref{definition of c-zeta}) to control $\eta(\cdot, U)$ if $U$ is a small open set.
We write
$$\eta'(u, U) = \inf_{v \in BV_c(U)} \int_U |d(u + v)| ~\vol'$$
for the analogue of $\eta$ with respect to the flat metric induced by normal coordinates.

\begin{lemma}\label{flattening of eta}
If $U \subseteq B(p, \zeta) \cap B(p, r)$ with $r < 1$, $U \Subset M$, then for every $u \in BV_l(M)$,
\begin{equation}\label{flattening of eta equation}
(1 - cr^2) \eta'(u, U) \leq \eta(u, U) \leq (1 + cr^2) \eta'(u, U).
\end{equation}
\end{lemma}
\begin{proof}
Let $x$ be a normal coordinate centered on $p$.
Choose $(u_n)$ in $BV_l(M)$ which is a minimizing sequence for $\int_U |du_n| ~\vol'$ subject to the trace condition $u_n|\partial U = u|\partial U$.
Thus $\int_U |du_n| ~\vol' \to \eta'(u, U)$ and by (\ref{definition of c-zeta}),
$$(1 - cr^2) \int_U |du_n| ~\vol' \leq \int_U |du_n| (1 - c|x|^2) ~\vol' \leq \int_U |du_n| ~\vol = \eta(u, U).$$
Taking $n \to \infty$ we get $(1 - cr^2) \eta'(u, U) \leq \eta(u, U)$.
The other estimate in (\ref{flattening of eta equation}) is similar.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The reduced boundary}\label{reduced boundary}
The notion of reduced boundary was introduced in \cite{deGiorgi55}, and a general discussion of reduced boundary in Riemannian manifolds is given by \cite{Battista_2021}.
Here we formulate the definition of reduced boundary in a way that will be very convenient to us, and also metric-independent.

\begin{definition}
Let $U$ be an open Caccioppoli set and $p \in M$ satisfy $\int_{B(p, r)} |d1_U| ~\vol > 0$ for every $r > 0$.
By the \dfn{conormal $1$-form} to $U$ we mean the $1$-form
$$\normal(p) = \lim_{r \to 0} \frac{\int_{B(p, r)} d1_U ~\vol}{\int_{B(p, r)} |d1_U| ~\vol}$$
if this limit indeed exists.
\end{definition}

\begin{lemma}
Let $U$ be a Caccioppoli set.
Then the conormal $1$-form $\normal(p)$ exists for $|d1_U|~\vol$-almost every $p \in M$, and is independent of the choice of coordinates.
\end{lemma}
\begin{proof}
Suppose that we have coordinate functions $x,\hat x$ defined near $p$, and write $\normal(p)$, $\widehat \normal(p)$ to mean the conormal $1$-form with respect to $x,\hat x$ respectively.
By the Lebesgue differentiation theorem \cite[\S2.9]{federer2014geometric}, $\normal(p)$ and $\widehat \normal(p)$ as a Borel $1$-form $|d1_U| ~\vol$-almost everywhere.

Suppose that $p$ lies in the domain of both $\normal$ and $\hat \normal$, and let $v \in T_pM$.
The coordinates $x,\hat x$ uniquely define isomorphisms $F_q, \hat F_q: T_qM \to T_pM$ whenever $q$ is close to $p$.
Then $q \mapsto F_q$ is continuous as a section of the bundle of linear maps $TM \to T_pM$, and $F_p = \hat F_p = \id_p$, so $F_q - \hat F_q \to 0$ as $q \to p$.
It follows from the definitions and the triangle inequality (\ref{connection triangle inequality}) that
\begin{align*}
|\normal(p) - \hat \normal(p)| &\leq \limsup_{q \to p} ||F_q - \hat F_q|| = 0. \qedhere
\end{align*}
\end{proof}

\begin{definition}
Let $U$ be an open Caccioppoli set.
The \dfn{reduced boundary} $\partial^* U$ is the set of all $x \in M$ such that the conormal $1$-form $\normal(x)$ to $U$ exists and satisfies $|\normal(x)| = 1$.
The \dfn{tangent space} $T_x\partial^* U$ is the kernel of $\normal(x)$.
\end{definition}

\begin{lemma}\label{reduced boundary is metricfree}
Suppose that $g_1,g_2$ are Riemannian metrics on $M$ with volume forms $\vol_1,\vol_2$.
Let $U$ be an open Caccioppoli set with conormal $1$-form $\normal^{(j)}$ and reduced boundary $N^{(j)}$ with respect to $g_j$.
Then $\normal^{(1)} = \normal^{(2)}$ and $N^{(1)} = N^{(2)}$.
Moreover, if $\omega = d1_U$ and $d\nu_j = |\omega| ~\vol_j$, then $\nu_1,\nu_2$ are mutually absolutely continuous.
In particular, a set is $\nu_1$-rectifiable iff it is $\nu_2$-rectifiable.
\end{lemma}
\begin{proof}
Applying (\ref{change of volume form}) as $\rho \o 0$ and writing $\normal^{(j)}$ for the conormal $1$-form with respect to $g_j$, we obtain $\normal^{(1)} = \normal^{(2)}$
whenever one and hence both sides is finite.
The definition of the reduced boundary then gives $N^{(1)} = N^{(2)}$.
Thus $d\nu_2/d\nu_1 = \vol_2/\vol_1$, which is a strictly positive smooth function.
Rectifiability is preserved under mutual absolute continuity, so we are done.
\end{proof}

The above lemma motivates why it is more natural to study $\normal$ than the normal vector field $\normal^\sharp$, even though the latter is usually considered more fundamental in the euclidean setting.

\begin{notation}
If $U$ is a Caccioppoli set and $V$ is an open set, we write
$$|\partial^* U \cap V| = \int_V |d1_U| ~\vol.$$
We also write $|\partial^* U \cap V|' = \int_V |d1_U| ~\vol'$.
Since $\partial^* U$ has Hausdorff dimension $d - 1$, this notation does not collide with the notation $|V|$ for the volume of an open set, which has Hausdorff dimension $d$.
\end{notation}

\begin{proposition}\label{regularity of reduced boundary}
Let $U$ be a Caccioppoli set.
Then:
\begin{enumerate}
\item $\partial^* U$ is rectifiable with respect to the measure $E \mapsto |\partial^* U \cap E|$.
\item $\partial^* U$ is dense in $\partial U$.
\item If $\normal$ extends to a continuous $1$-form on $\partial U$, then $\partial^* U = \partial U$ is a $C^1$ hypersurface.
\item If $\normal$ extends to a continuous $1$-form on $\partial U$ and $U$ has least perimeter, then $\partial U$ is as smooth as possible.
\end{enumerate}
\end{proposition}
\begin{proof}
The claim is local, so we can work in a small ball $B(p, \rho)$ in which we can find a flat metric $g_1$, and set $g_2 = g$.
So by Lemma \ref{reduced boundary is metricfree}, the first three claims immediately follow from their euclidean counterparts, which are given in \cite[Chapter 4]{Giusti77} or \cite{deGiorgi55}.

Now we deduce the fourth claim from the third with a standard elliptic regularity argument.
Suppose that $\normal$ extends to a continuous $1$-form on $\partial U$ and $U$ has least perimeter.
Let $f: \RR^{d - 1} \to M$ be a $C^1$ parametrization of $\partial U$, so that $f$ is a classical minimizer of the Lagrangian $\mathscr L$ defined by (\ref{Lagrangian density}).
We select coordinates $(x, \xi) \in T^* M$ with $dx_i = \xi_i$ and recall that if $a^{ij} = \partial_{\xi_i} \partial_{\xi_j}\mathscr L$, then
\begin{equation}\label{elliptic PDE}
\partial_{x_i} (a^{ij} \partial_{x_j} f) = 0.
\end{equation}
Since $a^{ij}$ is H\"older continuous, we may iterate the interior Schauder estimate \cite[Theorem 6.22]{gilbarg2015elliptic} to show that $a^{ij}$ and $f$ are as smooth as possible.
\end{proof}

\section{Monotonicity and density formulae}\label{inequalities}
Let us recall the classical monotonicity formula for minimal hypersurfaces in Riemannian manifolds \cite[\S7]{MarquesXX}:

\begin{proposition}[monotonicity formula]\label{monotonicity formula}
Let $p \in M$ and let $r$ denote the distance to $p$.
Then there exists $A \in \RR$ which depends continuously on $p$ such that for every set $U$ of least perimeter such that $\partial^* U$ is smooth,
$$\partial_r e^{Ar^2}r^{1 - d} |\partial^* U \cap B(p, r)| \geq 0.$$
If $M$ is flat then we can take $A = 0$.
\end{proposition}

Unfortunately Proposition \ref{monotonicity formula} is not quite sufficient for our purposes, for two reasons: it presupposes that $\partial^* U$ is smooth, and we will frequently require $A \leq 0$ to carry out our arguments.
At least we will be able to use Proposition \ref{monotonicity formula} to prove the following standard corollary:

\begin{proposition}[area estimates]\label{area estimates}
Suppose that $N$ is a smooth minimal hypersurface in a small ball $B$ with $\kappa$ connected components in $B$.
If $N$ is the boundary of a Caccioppoli set then $\kappa$ is finite.
\end{proposition}
\begin{proof}
We apply Proposition \ref{monotonicity formula} to each of the connected components in each turn to get a lower bound on the area of $N$, namely $|N| \gtrsim_B \kappa$.
Since $N$ is the boundary of a Caccioppoli set, $B \mapsto |N \cap B|$ is a Radon measure, so $\kappa$ must be finite.
\end{proof}

We now prove a monotonicity formula for approximately minimal hypersurfaces on euclidean space; in doing so we will incur a logarithmic loss that we will later show is harmless.

\begin{lemma}[monotonicity formula for approximately least perimeter]\label{approximate monotonicity}
Let $E$ be a Caccioppoli set in $\RR^d$ equipped with its euclidean metric such that for some $h \geq 0$ and $R > 0$ and every $0 < r < R$,
$$|\partial^* E \cap B_r| \leq (1 + h)\eta(f, r).$$
Then for every $0 < \rho < r < R$, with $\alpha = \log(r/\rho)$,
\begin{align*}
&\left|r^{1 - d} |\partial^* E \cap B_r| - \rho^{1 - d} |\partial^* E \cap B_\rho|\right|^2 \\
&\qquad \leq d\omega_d(1 + h)(1 + \alpha + h\alpha^2) \left(r^{1 - d} |\partial^* E \cap B_r| - \rho^{1 - d} |\partial^* E \cap B_\rho| + d\omega_d h \alpha\right).
\end{align*}
\end{lemma}
\begin{proof}
Let $f = 1_E$.
We recall from Proposition \ref{estimates on good set} that $\eta(f, s) \leq |\partial B_s|$, so that we can define
\begin{align*}
\psi(s) &= \int_{B_s} |df|~\vol - \eta(f, s)\\
&\leq h\eta(f, s) \leq h|\partial B_s| \leq d\omega_d hs^{d - 1}.
\end{align*}
From \cite[Proposition 5.12]{Giusti77} we have
$$\left|r^{1 - d} \int_{B_r} df ~\vol - \rho^{1 - d} \int_{B_\rho} df ~\vol\right|^2 \leq PQ$$
where
\begin{align*}
P &= 2r^{1 - d}(1 + (d - 1)\alpha) \int_{B_r} |df| + 2(d - 1)^2 \int_\rho^r \log \frac{s}{\rho} \psi(s) \frac{ds}{s^d},\\
Q &= r^{1 - d} \int_{B_r} |df| - \rho^{1 - d} \int_{B_\rho} |df| + (d - 1)\int_\rho^r \psi(s) \frac{ds}{s^d}.
\end{align*}
Evidently $\log(s/\rho) \leq \alpha$ so we estimate
$$\frac{1}{\alpha} \int_\rho^r \log \frac{s}{\rho} \psi(s) \frac{ds}{s^d} \leq \int_\rho^r \psi(s) \frac{ds}{s^d} \leq d\omega_d h \int_\rho^r \frac{ds}{s} = d\omega_d h\alpha.$$
Thus
$$\int_{B_r} |df|~\vol \leq (1 + h)\eta(f, r) \leq (1 + h)|\partial B_r| \leq d\omega_d(1 + h) r^{d - 1}.$$
These estimates immediately give
\begin{align*}
P &\leq d\omega_d(1 + h)(1 + \alpha + h\alpha^2),\\
Q &\leq r^{1 - d} \int_{B_r} |df|~\vol - \rho^{1 - d} \int_{B_\rho} |df|~\vol + d\omega_d h\alpha. \qedhere
\end{align*}
\end{proof}

\begin{lemma}\label{approximate monotonicity 2}
Let $E$ be a Caccioppoli set in $\RR^d$ such that for some $0 \leq h < 1$ and $R > 0$ and every $0 < r < R$,
$$|\partial^* E \cap B_r| \leq (1 + h)\eta(f, r).$$
Then for every $0 < r < R$,
$$\partial_r \left(r^{1 - d} |\partial^* E \cap B_r| + d\omega_d h \log r\right) \geq 0.$$
\end{lemma}
\begin{proof}
Dividing by both sides of Lemma \ref{approximate monotonicity} by $d\omega_d(1 + h)(1 + \alpha + h\alpha^2)$, which is positive since $0 \leq h < 1$ and $\alpha > 0$, we get
$$r^{1 - d} |\partial^* E \cap B_r| - \rho^{1 - d} |\partial^* E \cap B_\rho| + d\omega_d h \alpha \geq 0.$$
Writing out $\alpha = \log r - \log \rho$ we now see the claim.
\end{proof}

For the proofs of the next few inequalities we follow \cite{Mooney11}, but modify the arguments to account for the presence of a nonzero Ricci-Taylor error $c \geq 0$.
We write $\lesssim_d$ to emphasize that the implied constant only depends on the dimension\footnote{Presumably, one could compute the implied constants as explicit functions of the sharp Sobolev constants on $\RR^d$, but we do not need this.}, not the constants in (\ref{definition of c-zeta}), so these inequalities will be suitably scale-invariant.

\begin{lemma}[isoperimetric inequality]\label{isoperimetric inequality}
Suppose that $E$ is a Caccioppoli set in $M$ and $p \in \partial^* E$.
Then for every $r > 0$ smaller than the strong injectivity radius of $p$,
$$|E \cap B(p, r)|^{d - 1} \lesssim_d |\partial^*(E \cap B(p, r))|^d.$$
\end{lemma}
\begin{proof}
Let $f = 1_{E \cap B(p, r)}$, $q = d/(d-1)$.
We apply (\ref{definition of c-zeta}) and the $BV$ Sobolev inequality \cite[\S5.6.1]{evans1991measure}:
\begin{align*}
|E \cap B(p, r)|^{1/q} &= ||f||_{L^q(\vol)} \leq 2||f||_{L^q(\vol')} \\
&\lesssim_d \int_M |df| ~\vol' \leq \int_M |df| ~\vol\\
&= |\partial^*(E \cap B(p, r))|. \qedhere
\end{align*}
\end{proof}

\begin{lemma}[density estimate]\label{uniform density estimate}
Suppose that $E$ is a Caccioppoli set in $M$, $p \in \partial^* E$, and $r_0 > 0$ is smaller than the strong injectivity radius of $p$.
If for every $r \in (0, r_0)$,
\begin{equation}\label{density estimate hypothesis}
|\partial^* E \cap B(p, r)| \leq 2\eta(1_E, B(p, r)),
\end{equation}
then for every such $r$,
\begin{align*}|E \cap B(p, r)|, ~|(M \setminus E) \cap B(p, r)|, ~|\partial^* E \cap B(p, r)| &\gtrsim_d r^{d - 1}.\end{align*}
\end{lemma}
\begin{proof}
Let $q = d/(d-1)$. By Lemmata \ref{isoperimetric inequality} and \ref{estimates on good set} and (\ref{density estimate hypothesis}),
\begin{align*}
|E \cap B(p, r)|^{1/q} &\lesssim_d |\partial^*(E \cap B(p, r))| \\
&\leq |\partial^* E \cap B(p, r)| + |E \cap \partial B(p, r)|\\
&\leq 2\eta(1_E, B(p, r)) +  |E \cap \partial B(p, r)|\\
&\leq 3 |E \cap \partial B(p, r)|.
\end{align*}
Let $F(r) = |E \cap B(p, r)|$, $G = F^{1/d}$.
Since $|E \cap \partial B(p, r)| = \partial_r F(r)$,
$$F(r)^{1/q} \lesssim_d F'(r) \leq F'(r).$$
This gives the estimate on $|E \cap B(p, r)|$.
Since $M \setminus E$ shares its perimeter with $E$, the same argument with $M \setminus E$ replacing $E$ gives the analogous bound on $|(M \setminus E) \cap B(p, r)|$.

Let $f = 1_{E \cap B(p, r)}$ and let
$$[f] = \frac{1}{|B(p, r)|} \int_{B(p, r)} f ~\vol'.$$
By (\ref{definition of c-zeta}),
$$|\partial^* (E \cap B(p, r))| \geq 0.5 |\partial^* (E \cap B(p, r))|' = 0.5 \int_{B(p, r)} |df| ~\vol'.$$
By the $BV$ Poincar\'e inequality \cite[\S5.6.1]{evans1991measure}, it follows that
\begin{align*}
|\partial^* (E \cap B(p, r))| &\gtrsim_d ||f - [f]||_{L^1(B(p, r), \vol')} \\
&= ||1 - [f]||_{L^1(B(p, r) \cap E, \vol')} + ||[f]||_{L^1(B(p, r) \setminus E, \vol')}.
\end{align*}
From (\ref{definition of c-zeta}) we have
$$|E \cap B(p, r)| \leq 2\omega_{d - 1} r^{d - 1} [f] \leq 4|E \cap B(p, r)|$$
so
\begin{align*}
|\partial^* (E \cap B(p, r))| &\gtrsim_d ||1 - [f]||_{L^1(B(p, r) \cap E, \vol')} + ||[f]||_{L^1(B(p, r) \setminus E, \vol')}\\
&\geq 0.5 (1 - [f]) |B(p, r) \cap E| + 0.5 [f] |B(p, r) \setminus E|\\
&\geq 0.5 r^{1-d}|E \cap B(p, r)||E \setminus B(p, r)| \gtrsim_d r^{1 - d}. \qedhere
\end{align*}
\end{proof}

Now let us study the blowup of $M$ at a point $p$ on the reduced boundary of a set $U$ of least perimeter.
Our main result is that the tangent cone to $U$ at $p$ is analytic if $d \leq 7$.

We will be interested in objects that depend on a small parameter $t > 0$, but may depend on a choice of subsequence of $t_n \to 0$.
Thus we will suppress all subindices and implicitly pass to subsequences whenever referring to a limit $t \to 0$.\footnote{This can be made rigorous as follows. Fix a nonprincipal ultrafilter $\mathbf p$ on $2^{\aleph_0}$, and \emph{define} any limit as $t \to 0$ to be a $\mathbf p$-limit. See \cite{Tao07} for details.}

\begin{proposition}\label{blowup theorem}
Fix $p \in M$ with $\zeta \in (0, 1)$ chosen to be at most the strong injectivity radius of $p$.
Suppose that $U$ is an open set with least perimeter in $B(p, \zeta)$ and $p \in \partial^* U$.
Let $A = {\exp_p}^* U$, $A_t = \{v \in T_pM: tv \in A\}$, and $u_t = 1_{A_t}$.
Then:
\begin{enumerate}
\item The net $(u_t)$ has approximately least gradient with respect to the flat metric on $T_pM$ and in fact satisfies the estimate
\begin{equation}\label{approximately least gradient target}|
\partial^* A_t \cap V|' \leq (1 + ct^2)\eta'(A_t, V) \leq \eta'(U_t, V) + ct^2|\partial^* V|'
\end{equation}
for every Caccioppoli set $V \Subset T_pM$.
\item There exists an indicator function $u_0 \in BV_l(B'(r_0))$ of least gradient such that $u_t \to u_0$ in $L^1_l$, almost everywhere, and in total variation on sets with no singularities.
\item Let $C = \{u_0 = 1\}$. Then $0 \in \partial C$ and, if $d \leq 7$, then $\partial C$ is analytic\footnote{If
we had a suitably strong monotonicity formula, it would follow from the proof of \cite[Theorem 9.3]{Giusti77} that $\partial C$ is a minimal cone, so that by Theorem \ref{minimal cones in R8}, $\partial C$ is a hyperplane.
However, we note carefully that we have not proven that yet, as the results of Section \ref{inequalities} are not strong enough.}.
\end{enumerate}
\end{proposition}
\begin{proof}
Suppose that (\ref{approximately least gradient target}) holds.
Then $(u_t)$ clearly is a net of indicator functions of least gradient, so the existence of $u_0$ and $C$ is immediate from Corollary \ref{compactness}.
The analyticity of $\partial C$ then follows from \cite[Corollary 9.5]{Giusti77}, since $T_pM$ is flat, $d \leq 7$, and $C$ has least perimeter in $T_pM \cong \RR^d$.

Therefore we just have to prove the estimate (\ref{approximately least gradient target}), and that $0 \in \partial C$.

Let $B'_r$ denote the ball (with respect to the flat metric) of radius $r > 0$ centered at $0$ in $T_pM$; then ${\exp_p}_* B'_r = B(p, r)$ if $r < \zeta$.
Moreover, $A_t$ enjoys the scaling invariance
\begin{equation}\label{scaling of psi}
|\partial^* A_t \cap B'_r|' = t^{1 - d} |\partial^* A \cap B'_r|'.
\end{equation}
To prove (\ref{scaling of psi}), let $\mathscr H^{d - 1}$ denote $(d-1)$-dimensional Hausdorff measure with respect to the euclidean metric.
Then by \cite[Theorem 4.4]{Giusti77},
$$|\partial^* A \cap E|' = \mathscr H^{d - 1}(\partial^* A \cap E),$$
and $\mathscr H^{d - 1}$ scales as desired.

Let $V \Subset B'_\zeta$ be an arbitrary open Caccioppoli set.
We write $V_t = \{x \in \RR^d: tx \in V\}$, so that $V_{1/t} \subseteq B'_{t\zeta}$ and so by (\ref{definition of c-zeta}), $\vol/\vol' \approx 1 \pm ct^2$ on $V_t$.

\begin{claim}\label{blowup claim 1}
$(U_t)$ satisfies the estimate (\ref{approximately least gradient target}).
\end{claim}

We must show that for every $w \in BV_c(V)$\footnote{The definition of bounded variation only depends on the equivalence class of a measure with respect to mutual absolute continuity, and in particular does not depend on a choice of volume form.},
\begin{equation}\label{approximately least gradient target 2}
|\partial^* A_t \cap V|' \leq (1 + ct^2) \int_V |du_t + dw| ~\vol'.
\end{equation}
Indeed, if (\ref{approximately least gradient target 2}) holds, then we can take the infimum over all $w$ to get
$$|\partial^*AU_t \cap V|' \leq (1 + ct^2) \eta'(u_t, V)$$
and then complete the proof using Lemma \ref{estimates on good set} to bound $ct^2 \eta'(u_t, V) \leq ct^2 |\partial^* V|$.

We now prove (\ref{approximately least gradient target 2}).
Using (\ref{scaling of psi}, \ref{definition of c-zeta}),
$$t^{d - 1} | \partial^* A_t \cap V|' = |\partial^* A \cap V_{1/t}|' \leq (1 + ct^2) |\partial^* A \cap V_{1/t}|.$$
For a function $a$ we let $a_t$ denote the rescaling $a_t(x) = a(tx)$.
Since $U$ has least perimeter,
$$|\partial^* A \cap V_{1/t}| \leq \int_{V_{1/t}} |d(u + w_{1/t})| ~\vol = \int_{V_{1/t}} |d(u + w_{1/t})| \frac{\vol}{\vol'} ~\vol'.$$
Thus
$$|\partial^* A \cap V_{1/t}| \leq (1 + ct^2) \int_{V_{1/t}} |d(u + w_{1/t})| ~\vol'$$
and so
$$|\partial^* A_t \cap V|' \leq (t^{1 - d} + ct^{3 - d}) \int_{V_{1/t}} |d(u + w_{1/t})| ~\vol'.$$
Rescaling using (\ref{scaling of psi}), we obtain (\ref{approximately least gradient target 2}) and hence Claim \ref{blowup claim 1}.

\begin{claim}
$0 \in \partial C$.
\end{claim}

Owing to (\ref{approximately least gradient target}, \ref{definition of c-zeta}), if $r,t$ are small enough then
$$|\partial^* A_t \cap B'_r| \leq 1.5|\partial^* A_t \cap B'_r|' \leq 1.5(1 + ct^2) \eta'(u_t, B'_r) \leq 2 \eta(u_t, B'_r)$$
and hence by Lemma \ref{uniform density estimate},
$$|\partial C \cap B'(r)| \gtrsim r^{d - 1}.$$
Since $\partial C$ is a smooth $(d-1)$-dimensional hypersurface which is closed as a subset of $T_pM$, this is only possible if $0 \in \partial C$.
\end{proof}

One can show that just to show the existence of a (possibly singular) tangent cone, one only needs that $U$ is a Caccioppoli set.
This just requires a slight modification of our argument to show that $(u_t)$ is bounded in $BV_l$, as in the proof of \cite[Theorem 9.3]{Giusti77}.
However, we will never need this fact.

\section{de Giorgi's lemma}\label{de Giorgi section}
Let us now Proposition \ref{imprecise induction 2}.
This proposition is a variant of de Giorgi's lemma \cite{deGiorgi61}, which, along with a standard argument, will show that the reduced boundary $\partial^* U$ is as smooth as possible.

Throughout this section, let $g$ be a metric on $\RR^d$ such that
\begin{equation}
\text{lines through the origin are geodesics}.
\end{equation}

\begin{notation}
To facilitate geometry on $\RR^d$, we write $B(p, r)$ for a ball with respect to the metric $g$ and $B'(p, r)$ with respect to the euclidean metric.
We write $B_r = B(0, r)$ and $B'_r = B'(0, r)$.
We write
$$\Lambda(E, \rho) = |\partial^* E \cap B_\rho| - \left|\int_{B_\rho} d1_E ~\vol\right|$$
for the unnormalized excess of a Caccioppoli set $E \subseteq \RR^d$ in $B_\rho$.
We also write $\eta(E, \rho) = \eta(1_E, B_\rho)$, and write $\eta'$ and $\Lambda'$ for their euclidean versions, with $B_\rho$ replaced by $B'_\rho$ and $\vol$ replcaed by $\vol'$.
\end{notation}

\begin{theorem}[de Giorgi's lemma, euclidean version]
Let $g$ be the euclidean metric.
For every $d \geq 2$ and $0 < \alpha < 1$, there exists $\sigma > 0$ such that for every Caccioppoli set $E \subseteq \RR^d$ every $\rho > 0$ such that $E$ has least perimeter in $B_\rho$ and $\Lambda(E, \rho) < \sigma \rho^{d - 1}$,
\begin{equation}\label{classic DG}
\Lambda(E, \alpha \rho) < \alpha^d \Lambda(E, \rho).
\end{equation}
\end{theorem}

We recall the proof of de Giorgi's lemma; for the details, see \cite[\S4-5]{Miranda66}.
One first proves an analogous estimate to (\ref{classic DG}), but for the Laplace equation rather than the MSE \cite[Lemma 4.1]{Miranda66}:

\begin{lemma}\label{harmonic DG}
Let $g$ be the euclidean metric.
Let $[f]$ denote the mean of $f: B_\rho \to \RR$ with $B_\rho \subset \RR^{d - 1}$.
If $\Delta f = 0$, then
$$\int_{B_{\alpha\rho}} |df|^2 - [df]^2 ~\vol \leq \alpha^{d + 2} \int_{B_\rho} |df|^2 - [df]^2 ~\vol.$$
\end{lemma}

Since $\rho^{1 - d}\Lambda(E, \rho)$ is small, $E$ is approximately flat; moreover $E$ has least perimeter.
Therefore, after a mollification, $E$ can be locally written as the graph of a function $f: \RR^{d - 1} \to \RR$ such that $|df|$ is small and $f$ approximately solves the MSE.
The Lagrangian density $\mathscr L$ for the MSE satisfies (\ref{Lagrangian density}), so the linearization of $\mathscr L$ about $|df| = 0$ is $1 + |df|^2/2 ~\vol$, whose integral is the Dirichlet energy of $f$.
So we can approximate $f$ by a harmonic function using Lemma \ref{harmonic DG}.

The key lemma \cite[Teorema 4.4]{Miranda66} in the above argument, which transfers the estimates on the Laplace equation to smooth solutions to the MSE, is:

\begin{lemma}\label{Giusti64}
Let $(L_j)$ be a sequence of Caccioppoli sets in $\RR^d$ with $C^1$ boundary, $\beta_j > 0$, $\rho > 0$, and $\Lambda'(L_j, \rho) \leq \beta_j$.
Suppose that the conormal $1$-forms $\normal_j$ at $0$ satisfy
$$\lim_{j \to \infty} \inf_{\partial^* L_j \cap B_\rho} (\normal_j)_d = 1$$
and
$$|\partial^* L_j \cap B_\rho|' - \eta(L_j, \rho) \ll \beta_j.$$
Then for every $0 < \alpha < 1$,
$$\limsup_{j \to \infty} \frac{\Lambda'(L_j, \rho)}{\beta_j} \leq \alpha^{d + 1}.$$
\end{lemma}

While it would be interesting to develop an analogous estimate to Lemma \ref{harmonic DG} for the Laplace-Beltrami equation, we bypass this step and instead use Lemma \ref{harmonic DG} via Lemma \ref{Giusti64}.
As usual we will be interested not in a single set of least perimeter, but a sequence of sets of approximately least perimeter.

\subsection{An approximation lemma}
A set of approximately least perimeter may be rough, so before applying Lemma \ref{Giusti64} we need to mollify it.
To this end we use a modification of the arguments of \cite[Chapter 7]{Giusti77} and introduce the convolution kernel
$$\chi_\varepsilon(x) = C(\varepsilon) \varepsilon^{-d}\left(1 - \frac{|x|}{\varepsilon}\right) \vee 0.$$
Here $C(\varepsilon) = C(d, \vol, \varepsilon) > 0$ is the constant that enforces $||\chi_\varepsilon||_{L^1(\RR^d, \vol)} = 1$, and $|x| = |x|'$ if $\varepsilon < \zeta$, which is the only case that we will be interested in.

\begin{notation}
If $f$ is a Radon measure, we write $f^{(\varepsilon)} = f * \chi_\varepsilon$.
\end{notation}

\begin{lemma}\label{convolution fax}
If $2\varepsilon < \zeta$ and $K = (d+1)/\omega_d$, then:
\begin{enumerate}
\item $\chi_\varepsilon$ is radial, differentiable almost everywhere, and supported on $\overline B_\varepsilon$.
\item $K \leq 2C(\varepsilon) \leq 4K$, and $C(\varepsilon) \to K$ as $\varepsilon \to 0$.
\item If $f$ is an indicator function, then $f^{(\varepsilon)} \in C^1$.
\item For every $\delta \in (0, 1)$, every ball $V = B(\xi, \delta\varepsilon) \subset B_\zeta$, every $x \in B_1$, and every $y \in V$,
\begin{align*}
0.5\frac{K}{\varepsilon^d}\left(1 - \delta - \frac{|x - \xi|}{\varepsilon}\right) \leq \chi_\varepsilon(x - y) \leq 2\frac{K}{\varepsilon^d}\left(1 + \delta - \frac{|x - \xi|}{\varepsilon}\right).
\end{align*}
\end{enumerate}
\end{lemma}
\begin{proof}
We prove:
\begin{enumerate}
\item This is clear.
\item We first observe that for every $\delta > 0$, $K = C(d, \vol', \delta)$.
But since $\varepsilon < \zeta$, $\vol$ is well-approximated by $\vol'$ in the sense of (\ref{definition of c-zeta}), and $\vol/\vol'(0) = 1$.
Since, for every Radon measure $\mu$, $C(d, \mu, \delta)$ is defined to be an integral against $\mu$, this gives the result.
\item Suppose that there is a Borel set $E \subseteq B_\zeta$ such that $f = 1_E$. Then the identity $df^{(\varepsilon)} = f * d\chi_\varepsilon$ and the fact that $\chi_\varepsilon$ is differentiable almost everywhere and supported on $\overline B_\varepsilon$ implies
\begin{align*}
|df^{(\varepsilon)}(x) - df^{(\varepsilon)}(y)| &\lesssim_{\varepsilon,\vol,d} \int_{B_\varepsilon} |1_{E - x} - 1_{E - y}| ~\vol \\
&\leq |(E - x) \Delta (E - y)|
\end{align*}
where $\Delta$ denotes symmetric difference; this quantity vanishes in the limit $y \to x$ since $\overline E$ is compact. Therefore $f^{(\varepsilon)} \in C^1$.
\item We show the bound
$$\chi_\varepsilon(x - y) \leq 2\frac{K}{\varepsilon^d}\left(1 + \delta - \frac{|x - \xi|}{\varepsilon}\right)$$
as the other bound is similar.
We seek to minimize $|x - y|$ subject to $|y - \xi|' \leq \delta\varepsilon$.
On the other hand, if $|x - y| > \zeta$ then $\chi_\varepsilon(x - y) = 0$ and there is nothing to show, so we may assume $|x - y| = |x - y|'$.
This is minimized when $|x - y|' = |x - \xi| + \delta\varepsilon = |x - \xi|' + \delta\varepsilon$.
It remains to use the estimate $C(\varepsilon) \leq 2K$ to complete the proof.
 \qedhere
\end{enumerate}
\end{proof}

\begin{lemma}\label{big mollify}
Let $\zeta$ be the strong injectivity radius and $c$ the Ricci-Taylor error of $g$ at $0$.
Then, there exists\footnote{the implied constants in Big-O notation are here allowed to depend on $d$}
\begin{equation}\label{estimates on Gamma}
\Gamma = \Gamma(c, d, \zeta) \geq \min(\zeta^5, \zeta, O(1)c^{-O(1)}, O(1)) > 0
\end{equation}
such that every $\gamma \in (0, \Gamma)$ and every $\ell \geq 1$ have the following property:
For every set $U$ of least perimeter with respect to $g$ and $t \in (\gamma^{0.8}, \gamma^{0.25})$, let $E = U_t$ and suppose that
\begin{equation}\label{hypothesis on Giusti73}
|\partial^* E \cap B_1| \leq \gamma + \int_{B_1} \partial_d 1_E~\vol.
\end{equation}
Let $\sigma = \gamma^{0.5/(d-1)}$, $I = (d^2 \gamma^2, 1 - d^2\gamma^2) \Subset (0, 1)$, and $\varepsilon = \gamma^\ell$.
Then for every $x \in B_{1 - 2\sigma}$ such that ${1_E}^{(\varepsilon)}(x) \in I$,
$$(|d1_E| - \partial_d 1_E)^{(\varepsilon)}(x) < \gamma^{0.24} |d1_E|^{(\varepsilon)}(x).$$
Here $|d1_E|$ and $\partial_d 1_E$ are defined with respect to $\vol$.
\end{lemma}
\begin{proof}
Since $\gamma < 1$ and $\ell \geq 1 > 0.5/(d-1)$, $\varepsilon < \sigma$.
Let $r = |x| < 1 - 2\sigma$, $0 < \delta < 1/2$, $f = |d1_E| - \partial_d 1_E$, and
$$\alpha = \log \frac{\sigma}{2\delta\varepsilon}.$$

Here is the strategy of the proof:
We want to estimate $f^{(\varepsilon)}$.
To do this, we will estimate both $(f1_{B_{\varepsilon(1 - 2\delta)}})^{(\varepsilon)}$ and $(f1_{B_\varepsilon \setminus \overline{B_{\varepsilon(1 - 2\delta)}}})^{(\varepsilon)}$.
As we go through the proof, we will need to make stronger and stronger assumptions on $\Gamma$, while carefully ensuring that our assumptions are not so strong that we violate the lower bound (\ref{estimates on Gamma}) on $\Gamma$.

\begin{assumption}\label{BMA1}
$\Gamma \leq \zeta$.
\end{assumption}

Since $\zeta$ is at least the (weak) injectivity radius at $0$, for every $\rho \in (0, \gamma]$, $B_\rho = B_\rho'$\footnote{Note carefully that this only applies to balls centered at $0$.}.

We now cover the domain of integration of the convolution by euclidean balls\footnote{as opposed to $g$-balls} on which we can easily apply the monotonicity formula.
We can use a greedy algorithm to find a finite set $\mathcal V$ of disjoint euclidean balls $V$ of radius $\delta\varepsilon$ with $\cent V \in \partial^* E \cap B_{\varepsilon(1 - 2\delta)}$ such that
\begin{equation}\label{2dilates cover}
\partial^* E \cap B_{\varepsilon(1 - 2\delta)} \subseteq \bigcup_{V \in \mathcal V} 2V.
\end{equation}
Since $\mathcal V$ is disjoint and $\chi_\varepsilon = 0$ outside of $B_\varepsilon$,
$$|d1_E(x)|^{(\varepsilon)} \geq \sum_{V \in \mathcal V} \int_V \chi_\varepsilon(x - y) |d1_E(y)| ~\vol(y).$$
By (\ref{2dilates cover}),
$$(f1_{B_{\varepsilon(1 - 2\delta)}})^{(\varepsilon)} \leq \sum_{V \in \mathcal V} \int_{2V} \chi_\varepsilon(x - y)f(y) ~\vol(y).$$

Our next assumption is necessary to show that $E$ has approximately least perimeter with respect to the euclidean metric.
This will allow us to apply our logarithmic monotonicity formula, and density estimates.

\begin{assumption}\label{BMA2}
$\Gamma \leq \min(1/3, 1/c, \zeta^5)$.
\end{assumption}

\begin{claim}\label{BMC1}
For every $V \in \mathcal V$ and ball $W \subseteq 2V$,
\begin{equation}\label{approximately least perimeter in 2V}
|\partial^*E \cap W|' \leq (1 + \gamma)\eta(E, W) < (2 \wedge (1 + 3\gamma))\eta'(U, W).
\end{equation}
\end{claim}
\begin{proof}[Proof of claim]
Since $\gamma < \zeta^5$,
\begin{align*}
\varepsilon(1 - 2\delta) &< \varepsilon = \gamma^\ell \leq \gamma = \gamma^{0.8} \gamma^{0.2} < t\zeta.
\end{align*}
Moreover, since $U$ has least perimeter with respect to $g$, $E = U_t$ has least perimeter with respect to $\widetilde g = t^2g$.
Rescaling the volume form appropriately, we conclude that
$$\widetilde \vol \approx (1 + cr^2) ~\widetilde \vol'$$
on $B_r$ provided that $r < t\zeta$; in particular this holds if $r = \varepsilon(1 - 2\delta)$.
Since $E$ has least perimeter with respect to $\widetilde g$,
\begin{align*}
|\partial^* E \cap W|' &\leq (1 + c\varepsilon^2(1 - 2\delta)^2) |\partial^* E \cap W| \leq (1 + c\gamma^2) \eta(E, W);
\end{align*}
here all volume forms are the rescaled versions.
Since $\gamma < c$, we conclude
$$|\partial^* E \cap W|' \leq (1 + \gamma) \eta(E, W),$$
and a similar argument shows $(1 + \gamma) \eta(E, W) \leq (1 + \gamma)^2 \eta'(E, W)$.
Moreover, since $\gamma < 1/3$, $(1 + \gamma)^2 < 1 + 3\gamma$ and $(1 + \gamma)^2 < 2$.
\end{proof}

\begin{claim}\label{BMC1.5}
For every $V \in \mathcal V$,
\begin{equation}\label{Giusti73 upper bound}
\int_V \chi_\varepsilon(x - y) |d1_E(y)| ~\vol(y) \gtrsim_d \frac{\delta^{d - 1}}{\varepsilon}(1 - \delta - |x - \xi|\varepsilon^{-1}).
\end{equation}
\end{claim}
\begin{proof}[Proof of claim]
By Lemma \ref{convolution fax} and the fact that $\delta \varepsilon < \gamma < \zeta$,
\begin{equation}\label{preliminary lower bound}
\int_V \chi_\varepsilon(x - y) |d1_E(y)| ~\vol(y) \geq \frac{K}{2\varepsilon^d}(1 - \delta - |x - \cent V|\varepsilon^{-1}) |\partial^* E \cap V|.
\end{equation}
By Claim \ref{BMC1} and Lemma \ref{uniform density estimate},
$$|\partial^* E \cap V| \gtrsim_d (\delta\varepsilon)^{d - 1}.$$
Combining this estimate with (\ref{preliminary lower bound}) gives (\ref{Giusti73 upper bound}).
\end{proof}

\begin{claim}\label{BMC2}
Let
\begin{equation}\label{definition of beta}
\frac{\beta}{2K} = 3cd\omega_d\alpha\gamma + \gamma^{0.5} + d\omega_d(1 + c\gamma^{0.5})\sqrt{(1 + \alpha + c\gamma^{0.5}\alpha^2)(1 + c\alpha)}\gamma^{0.25}.
\end{equation}
Then
\begin{equation}\label{kernel eliminated}
\int_{2V} \chi_\varepsilon(x - y)f(y) ~\vol(y) \leq \frac{\delta^{d - 1}}{\varepsilon}(1 + 2\delta + |x - \xi|\varepsilon^{-1})\beta.
\end{equation}
\end{claim}
\begin{proof}[Proof of claim]
We first bound $\int_{2V} \chi_\varepsilon(x - y)f(y) ~\vol(y)$ from above by
\begin{equation}\label{eliminate the kernel}
\int_{2V} \chi_\varepsilon(x - y)f(y) ~\vol(y) \leq \frac{2K}{\varepsilon^d}(1 + 2\delta - |x - \xi|\varepsilon^{-1}) \int_{2V} f ~\vol
\end{equation}
using Lemma \ref{convolution fax}.
By Claim \ref{BMC1} and Lemma \ref{approximate monotonicity 2},
$$\int_{2V} f~\vol \leq (2\delta\varepsilon)^{d - 1} \left(\sigma^{1 - d} \int_{B(\xi, \sigma)} (f + \partial_d1_E) ~\vol - (2\delta\varepsilon)^{1 - d} \int_V \partial_d1_E ~\vol + 3cd\omega_d\gamma\alpha\right).$$
But $\sigma^{1 - d} = \gamma^{-0.5}$ and since $\xi \in B_{\varepsilon(1 - 2\delta)}$, (\ref{hypothesis on Giusti73}) gives
\begin{equation}\label{we get sqrt gamma}
\sigma^{1 - d} \int_{B(\xi, \sigma)} f ~\vol \leq \sigma^{1 - d} \int_{B_1} f ~\vol \leq \gamma^{0.5}.
\end{equation}
Meanwhile, by Lemma \ref{approximate monotonicity},
\begin{align*}
&\sigma^{1 - d} \int_{B(\xi, \sigma)} \partial_d1_E ~\vol - (2\delta\varepsilon)^{1 - d} \int_{2V} \partial_d1_E ~\vol\\
& \qquad \leq \left|\sigma^{1 - d} \int_{B(\xi, \sigma)} d1_E ~\vol - (2\delta\varepsilon)^{1 - d} \int_{2V} d1_E ~\vol\right|\\
& \qquad \leq d\omega_d (1 + ct^2)\sqrt{1 + \alpha + ct^2 \alpha^2} \\
&\qquad \qquad \sqrt{\sigma^{1 - d} \int_{B(\xi, \sigma)} |d1_E| ~\vol - (2\delta\varepsilon)^{1 - d} \int_{2V} |d1_E| ~\vol + d\omega_dct^2\alpha}.
\end{align*}
We now rewrite
\begin{align*}
&\sigma^{1 - d} \int_{B(\xi, \sigma)} |d1_E| ~\vol - (2\delta\varepsilon)^{1 - d} \int_{2V} |d1_E| ~\vol\\
&\qquad = \left(\sigma^{1 - d} \int_{B(\xi, \sigma)} f ~\vol\right) + \left(\sigma^{1 - d} \int_{B(x, \sigma)} \partial_d 1_E ~\vol\right) - \left((2\delta\varepsilon)^{1 - d} \int_{2V} |d1_E| ~\vol\right)\\
&\qquad=: P + Q - R.
\end{align*}
We use (\ref{we get sqrt gamma}) to dispose of $P$ as $P \leq \gamma^{0.5}$, and integrate $Q$ by parts to get, with $\normal$ the unit normal to $\partial B(x, \sigma)$,
$$Q = \sigma^{1 - d} \int_{\partial B(x, \sigma) \cap E} \normal_d ~\vol \leq \sigma^{1 - d} \int_{\partial B(x, \sigma)} ~\vol = \omega_{d - 1}.$$
On the other hand, $\xi \in \partial^* E$, so if we write $N$ for the disk bounded by the equator in $2V$, then
$$|2V \cap \partial^* E| \geq |2V \cap N| = \omega_{d - 1}(2\delta\varepsilon)^{d - 1}$$
or in other words $R \geq \omega_{d - 1}$.
This gives $P + Q - R \leq \gamma^{0.5}$, so
$$\sigma^{1 - d} \int_{B(\xi, \sigma)} |d1_E| ~\vol - (2\delta\varepsilon)^{1 - d} \int_{2V} |d1_E| ~\vol \leq (1 + c\alpha)\gamma^{0.5}.$$
Plugging everything into (\ref{eliminate the kernel}) gives (\ref{kernel eliminated}).
\end{proof}

Combining Claims \ref{BMC1.5} and \ref{BMC2},
$$\int_{2V} \chi_\varepsilon(x - y)f(y) ~\vol(y) \lesssim_d \beta \int_V \chi_\varepsilon(x - y)|d1_E(y)| ~\vol(y).$$
Since $\beta$ does not depend on $V$, we sum over $\mathcal V$ to obtain
\begin{equation}\label{Giusti73 inner estimate}
\int_{B_{\varepsilon(1 - 2\delta)}} \chi_\varepsilon(x - y)f(y) ~\vol(y) \lesssim_d \beta |d1_E|^{(\varepsilon)}(x).
\end{equation}

Now we estimate $\chi_\varepsilon(x - y)f(y)$ on $Z = B_\varepsilon \setminus B_{\varepsilon(1 - 2\delta)}$.
Since $\varphi(x) \in I$, we can reason identically to \cite[p92]{Giusti77} with \cite[(5.14)]{Giusti77} replaced with Lemma \ref{estimates on good set} and Claim \ref{BMC1} to deduce that
$$\int_Z \chi_\varepsilon(x - y)f(y) ~\vol(y) \lesssim_d (1 + c\gamma^{0.25}) \delta \gamma^{1 - d} |d1_E|^{(\varepsilon)}(x).$$
Adding this estimate to (\ref{Giusti73 inner estimate}), we find $K = K(d) > 0$ such that
$$f^{(\varepsilon)}(x) \leq K(\beta + \delta \gamma^{1 - d}(1 + c\gamma^{0.25}))|d1_E|^{(\varepsilon)}(x).$$
Thus we set
$$\lambda = K(\beta + \delta \gamma^{1 - d}(1 + c\gamma^{0.25})).$$

The proof will be complete if we select the parameters $\delta,\Gamma$ so that $\lambda < 1$ and (\ref{estimates on Gamma}) holds.
First we set $\delta = \gamma^d$, which is valid if we make the following assumption:
\begin{assumption}\label{BMA3}
$\Gamma \leq 2^{-1/d}$.
\end{assumption}

\begin{claim}\label{BMC3}
There exists $\Gamma_0 = \Gamma_0(c, d) \gtrsim_d c^{-O(1)} > 0$ such that if $\gamma < \Gamma_0$ then $\lambda < \gamma^{0.24}$.
\end{claim}
\begin{proof}[Proof of claim]
Since $\varepsilon < \sigma$, $\alpha \leq d\log(2/\gamma)$, so
\begin{align*}
\beta &\leq 3cd^2\omega_d\log(2/\gamma)\gamma +  \gamma^{0.5}\\
&\qquad+ d\omega_d(1 + c\gamma^{0.5})\sqrt{(1 + d\log(2/\gamma) + cd^2\gamma^{0.5}\log(4/\gamma))(1 + cd\log(2/\gamma))}\gamma^{0.25}\\
&\lesssim_{c, d} \gamma^{0.25}
\end{align*}
and hence
$$\lambda \lesssim_{c, d} \gamma^{0.25} \ll \gamma^{0.24}.$$
In particular we can find $\Gamma_0 = \Gamma_0(c, d) \gtrsim_d c^{-O(1)} > 0$ so that if $\Gamma \leq \Gamma_0$ then $\lambda < \gamma^{0.24}$.
\end{proof}

To finish the proof we combine Claim \ref{BMC3} with Assumptions \ref{BMA1}, \ref{BMA2}, and \ref{BMA3} to get
\begin{align*}\Gamma &= \min(1/3, 2^{-1/d}, 1/c, \zeta^5, \Gamma_0). \qedhere
\end{align*}
\end{proof}

We will not need it, but it is interesting to note that Assumption \ref{BMA2} can be weakened significantly if the metric $g$ is homogeneous in the sense that there is a Lie group $L$ which acts transitively on $(\RR^d, g)$ by isometries.
Indeed, let $\mathcal V$ be the set returned by the greedy algorithm used in the proof of Lemma \ref{big mollify}.
Then for every $V \in \mathcal V$, there is an isometry in $L$ which interchanges $\cent V$ and $0$, so we can assume that $\cent V = 0$.
One can then show that $E$ has approximately least perimeter with respect to the euclidean metric without using Assumption \ref{BMA2}.

\subsection{Mollification data sets}
\begin{definition}
A set $\mathcal G$ of metrics on $\RR^d$ is \dfn{uniform for mollification} if:
\begin{enumerate}
\item There exists $\zeta > 0$ which is smaller than the strict injectivity radius at $0$ of every metric in $\mathcal G$.
\item There exists $c > 0$ which is larger than the Ricci-Taylor error at $0$ of every metric in $\mathcal G$.
\item For every $g \in \mathcal G$, the standard basis of $\RR^d$ is $g$-orthogonal\footnote{but possibly not $g$-orthonormal}.
\end{enumerate}
\end{definition}

The point is that if $\mathcal G$ is uniform for mollification then the constant $\Gamma$ appearing in Lemma \ref{big mollify} can be chosen uniformly over $\mathcal G$.

\begin{example}
If $g$ is a metric on $M$ and we identify $M$ with $T_pM \cong \RR^d$ using geodesic normal coordinates,
\begin{equation}\label{condition for MDS}
\{\lambda g: \lambda \in [1, \infty)\} \text{ is uniform for mollification.}
\end{equation}
This is an immediate consequence of (\ref{scale invariance}).
\end{example}

\begin{definition}
A sequence of tuples $\mathcal A = (\mathcal A_j)_{j \geq j_0}$,
$$\mathcal A_j = (g_j, \gamma_j, t_j, U_j, E_j),$$
is a \dfn{mollification data set}
if $(g_j)$ is uniform for mollification, $\gamma_j \in (0, 1)$, $(\gamma_j) \in \ell^1$, $(t_j) \in [{\gamma_j}^{0.7}, {\gamma_j}^{0.3}]$, $U_j$ has $g_j$-least perimeter, $E_j = (U_j)_{t_j}$, and
$$\int_{B_1} |d1_{E_j}| - \partial_d 1_{E_j} ~\vol_{g_j} \leq \gamma_j.$$
We \dfn{truncate} a mollification data set by increasing the index $j_0$ by a finite amount.
\end{definition}

In the remainder of this subsection we allow implied constants to depend on the mollification data set $\mathcal A$, but they must be uniform in the parameter $j$.

\begin{lemma}\label{big mollify 2}
Let $\mathcal A$ be a mollification data set and let $r \in (0, 1)$.
For every sequence of Caccioppoli sets $(L_j)$ in $B_1$ such that
$$||1_{L_j} - 1_{E_j}||_{L^1(\partial B_r)} \ll \gamma_j$$
and $|\partial L_j| \lesssim 1$, one has
$$|\partial^* E_j \cap B_r| - |\partial L_j \cap B_r| \ll \gamma_j.$$
\end{lemma}
\begin{proof}
Since $U_j$ has $g_j$-least perimeter, (\ref{change of volume form}) implies
\begin{align*}
|\partial^* E_j \cap B_r| & = {t_j}^{1 - d}|\partial^* U_j \cap B_{t_jr}|\\
&\leq {t_j}^{1 - d} (1 + c(t_jr)^2) |\partial^* U_j \cap B_{t_jr}|_{g_j}\\
&= {t_j}^{1 - d} (1 + c(t_jr)^2) \eta_{g_j}(U_j, t_jr).
\end{align*}
By Lemma \ref{estimates on good set},
$$\eta_{g_j}(U_j, B_{t_jr}) \leq \eta_{g_j}(L_j, B_{t_jr}) + ||1_{U_j} - 1_{(L_j)_{t_j}}||_{L^1(\partial B_{t_jr}, \vol_j)}.$$
which rescales to
\begin{equation}\label{Giusti723 rescaled}
||1_{U_j} - 1_{(L_j)_{t_j}}||_{L^1(\partial B_{t_jr})} \ll \gamma_j {t_j}^{1 - d}.
\end{equation}
From Lemma \ref{flattening of eta}, (\ref{Giusti723 rescaled}), and $\eta(L_j, t_jr) \leq |\partial^* L_j \cap B_{t_jr}|$,
\begin{align*}
\eta_{g_j}(L_j, t_jr) + ||1_{U_j} - 1_{(L_j)_{t_j}}||_{L^1(\partial B_{t_jr}, \vol_j)} &\ll (1 + c(t_jr)^2) \eta(L_j, t_jr) + \gamma_j {t_j}^{1 - d})\\
&\leq (1 + c(t_jr)^2) |\partial^* L_j \cap B_{t_jr}| + \gamma_j {t_j}^{1 - d}.
\end{align*}
Moreover $t_j \leq {\gamma_j}^{0.6}$ implies
$$(1 + c(t_jr)^2) \leq (1 + c(t_jr)^2)^2 < 1 + 3c(t_jr)^2 = 1 + o(\gamma_j) < 2$$
if $j$ is large enough.
So now we recale everything and use $|\partial L_j| \lesssim 1$ to get
\begin{align*}
|\partial^* E_j \cap B_r| &\leq (1 + c(t_jr)^2)^2 |\partial L_j \cap B_r| + (1 + c(t_jr)^2) o(\gamma_j) \\
&< (1 + o(\gamma_j)) |\partial L_j \cap B_r| + 2o(\gamma_j)\\
&\leq |\partial^* L_j \cap B_r| + o(\gamma_j). \qedhere
\end{align*}
\end{proof}

\begin{lemma}\label{big mollify 3}
For every mollification data set $\mathcal A$ there exists a sequence of Caccioppoli sets $(L_j)$ with $C^1$ boundary and the following property:
Let $r \in (0, 1)$ be drawn uniformly at random, let $j$ be sufficiently large, and let $\normal_j$ be the conormal $1$-form to $\partial L_j$.
Then almost surely,
$$|\partial L_j \cap B_r| - \eta(L_j, B_r) \ll \gamma_j,$$
$$|\partial L_j \cap B_r| - |\partial^* U_j \cap B_r| \ll \gamma_j,$$
$$\left|\int_{B_r} d1_{L_j} - d1_{U_j} ~\vol\right| \ll \gamma_j,$$
and for every $\rho \in (0, r)$,
$$\lim_{j \to \infty} \inf_{\partial L_j \cap B_\rho} (\normal_j)_d = 1.$$
\end{lemma}
\begin{proof}
Since $(g_j)$ is uniform for mollification, we can take the Ricci-Taylor error $c$ and strong injectivity radius $\zeta$ to be uniform in $j$.
Let $\Gamma$ be the constant induced by $c,\zeta$ in Lemma \ref{big mollify}.
After truncating a finite amount of mollification data we may assume that for every $j$, $\gamma_j \in (0, \Gamma)$.
One can then copy the proofs of Lemma 7.2, Remark 7.4, and Lemma 7.5 of \cite{Giusti77} nearly verbatim, replacing \cite[Lemma 7.1]{Giusti77} with Lemma \ref{convolution fax} and \cite[Lemma 7.3]{Giusti77} with Lemma \ref{big mollify}.
Let us just record where we must modify the argument.

Since \cite[(5.14)]{Giusti77} is false in our setting, it must be replaced with, for every $r \in (0, 1)$,
\begin{equation}\label{Giusti514replace}
|\partial^* E_j \cap B_r| \lesssim_d r^{d - 1}
\end{equation}
where the implied constant only depends on $d$.
This is sufficient at any point in the proof where one uses \cite[(5.14)]{Giusti77}.
To prove (\ref{Giusti514replace}), we first recall from the definition of $\Gamma$ that $\zeta/t_j \geq 1 > r$.
Thus by definition of $\zeta$, the minimality of $U_j$, and Lemma \ref{estimates on good set},
\begin{align*}
|\partial^* E_j \cap B_r| &\leq 2|\partial^* U_j \cap B_r|_{g_j} \lesssim_d r^{d - 1}.
\end{align*}

In addition, \cite[(5.5)]{Giusti77} is false in our setting, and so we must give an alternative proof of \cite[(7.24)]{Giusti77}.
By \cite[(7.23)]{Giusti77}, however, we meet the hypotheses of Lemma \ref{big mollify 2}, which implies \cite[(7.24)]{Giusti77}, except possibly that $|\partial L_j \cap B_r|$ could be unbounded as $j \to \infty$.
To avoid this, we use \cite[(7.20-7.21)]{Giusti77} and (\ref{Giusti514replace}) to bound for every $j$ large enough
\begin{align*}
|\partial L_j \cap B_r| &\leq \int_{B_r} |d((1_{E_j})^{({\gamma_j}^4)})| ~\vol \leq 2|\partial^*E \cap B_r| \lesssim 1. \qedhere
\end{align*}
\end{proof}

\subsection{Induction on scales}
\begin{lemma}\label{induction 1}
For every mollification data set $\mathcal A$ and $\alpha \in (0, 1)$,
$$\limsup_{j \to \infty} \frac{\Lambda(E_j, \alpha)}{\gamma_j} \leq \alpha^{d + 1}.$$
\end{lemma}
\begin{proof}
Let $(L_j)$ be the sequence generated by Lemma \ref{big mollify 3}.
Then for almost every $r \in (0, 1)$ we can make the following argument:
By definition of $(L_j)$, $(L_j)$ and every $0 < \rho < r$ meet the hypotheses, and hence the conclusion, of Lemma \ref{Giusti64} with $\beta_j = \gamma_j$.
The conclusion of Lemma \ref{big mollify 3} implies that this happens iff $(E_j)$ and $\rho$ satisfy the conclusion of Lemma \ref{Giusti64} and now the claim follows if we take $\rho \to 1$.
\end{proof}

\begin{proposition}[de Giorgi inequality]\label{induction 2}
For every $p \in M$ and every $\alpha \in (0, 0.5)$ there exists $\sigma > 0$ such that for every set $U$ of least perimeter and every $\rho \in (0, 1)$ with unnormalized excess $\Lambda(U, \rho)$ at $p$ such that
$$\Lambda(U, \rho) < \sigma \rho^{d - 1},$$
then
$$\Lambda(U, \alpha\rho) < \alpha^d \Lambda(U, \rho).$$
Moreover, if $\alpha$ is fixed then $\sigma$ depends continuously on $p$.
\end{proposition}
\begin{proof}
We can work in geodesic normal coordinates and hence assume that $g$ is a metric on $\RR^d$ and $p = 0$.
Towards a contradiction suppose that there exist $\alpha \in (0, 1)$, $(\rho_j)$, $(\gamma_j) \in \ell^1$, and Caccioppoli sets $(U_j)$ such that $U_j$ has least perimeter in $B(p, \rho_j)$ and
$$\Lambda(U_j, \rho_j; g) = \gamma_j {\rho_j}^{d - 1}$$
but
$$\Lambda(U_j, \alpha \rho_j; g) > \alpha^d \gamma_j {\rho_j}^{d - 1}.$$

Define the metric $g_j = {\rho_j}^{-0.5} g$, so that $B(p, \rho_j; g) = B(p, 1; g_j)$ and $U_j$ has least perimeter in $B(p, 1; g_j)$.
In particular
$$\Lambda(U_j, 1; g_j) = \gamma_j$$
and
$$\Lambda(U_j, \alpha; g_j) > \alpha^d \gamma_j.$$
After deleting finitely many entries and possibly rescaling every metric by a constant, we may assume that $\gamma_j \in (0, 1)$ and that we can take $\zeta = 1$ for every $g_j$.
We may also rotate the sets $E_j$ and the metrics $g_j$ so that $\int_{B_1} d1_{E_j} ~$ lies on the $x_d$-axis.
TODO Clarify this

We define $t_j = \gamma_j^{0.6}$ and $E_j = (U_j)_{t_j}$ and claim that
$$\mathcal A_j = (g_j, 2\gamma_j, t_j, U_j, E_j)$$
is a mollification data set such that
\begin{equation}
\label{contradicting MDS}
\Lambda(E_j, \alpha) > 0.5\alpha^d\gamma_j.
\end{equation}
By (\ref{condition for MDS}) and the assumption that $\rho_j < 1$ (so ${\rho_j}^{-0.5} > 1$), $(g_j)$ is uniform for mollification.
Moreover
\begin{align*}
|\partial^* E_j \cap B_1| - \left|\int_{B_1} \partial_d 1_{E_j} ~\vol\right| &= |\partial^* E_j \cap B_1| - \left|\int_{B_1} d1_{E_j} ~\vol\right| \\
&\leq 2|\partial^* U_j \cap B_1|_j - 2\left|\int_{B_1} d1_{U_j} ~\vol_j\right| = \Lambda(U_j, 1; g_j) \\
&\leq 2\gamma_j.
\end{align*}
The proof of (\ref{contradicting MDS}) is similar, which proves our claim.

By Lemma \ref{induction 1} and (\ref{contradicting MDS}), if $j$ is large then
\begin{align*}
\Lambda(E_j, \alpha) &> 0.5 \alpha^d \gamma_j \geq \frac{\gamma_j}{4\alpha} \frac{\Lambda(E_j, \alpha)}{2\gamma_j} \\
&= \frac{\Lambda(E_j, \alpha)}{2\alpha}
\end{align*}
which is a contradiction if $2\alpha < 1$.

TODO Continuity of $\sigma$
\end{proof}

We may weaken the hypothesis of Proposition \ref{induction 2} to $\alpha \in (0, 1)$ by modifying the definition of strong injectivity radius suitably depending on $\alpha$.
We do not need this fact, however, so we omit the details.

Let us record the consequences of Proposition \ref{induction 2} that we need in the proof of Theorem \ref{main thm}.

\begin{corollary}\label{tangent cones 1}
Let $p \in M$, $\alpha \in (0, .5)$, and let $\sigma$ be as in Proposition \ref{induction 2}.
Let $U$ be a Caccioppoli subset of $M$ with $p \in \partial U$.
If $U$ has least perimeter in $B(p, \rho)$ and
$$\Lambda(U, \rho) < \sigma \rho^{d - 1},$$
then $\partial U \cap B(p, \rho(\alpha - \alpha^{d/(d-1)}))$ is a hypersurface which is as smooth as possible, and $p \in \partial^* U$.
\end{corollary}
\begin{proof}
In the proofs of Lemma 8.2 and Theorem 8.4 of \cite{Giusti77}, replace \cite[Theorem 8.1]{Giusti77} with Proposition \ref{induction 2}, \cite[(5.16)]{Giusti77} with Lemma \ref{uniform density estimate},
and \cite[Theorem 4.11]{Giusti77} and \cite{morrey2009multiple} with Proposition \ref{regularity of reduced boundary}.
TODO: Replace these citations with the citations used by \cite[Teorema 4]{Miranda67}.
Also check the continuity
\end{proof}

\section{Proof of main theorem}\label{proof of main thm}
We are finally ready to prove Theorem \ref{main thm} by applying Proposition \ref{blowup theorem} and Corollary \ref{tangent cones 1}.
At this point we just need to give an argument analogous to the proof of \cite[Teorema 4]{Miranda67} to patch everything together.

To this end, let $u$ be a function of least gradient on $M$.
Let $U$ be a superlevel set of $u$.
By Corollary \ref{level sets are minimal}, $U$ has least perimeter, so by Proposition \ref{blowup theorem} and the fact that $d \leq 7$, for every $p \in M$ there exists a hyperplane $C$ in $T_pM$ after passing to a subsequence we obtain the convergence $1_{U_t} \to 1_C$ in $L^1_{loc}$
and in total variation along all hypersurfaces in $T_pM$ with no singularities for $(u_n)$.
Let $g'$ denote the flat metric on $T_pM$.

Let $\normal^{(t)}$ be the conormal $1$-form to $U_t$ and $\normal^{(0)}$ the conormal $1$-form to $C$.
Let $x_t \in \partial U_t$ satisfy $x_t \to 0$.
By Lemma \ref{reduced boundary is metricfree}, these notions are preserved by exponential pushforward.
Moreover, since $C$ is a hyperplane, $\Lambda(C, r; g') = 0$ whenever $r > 0$.
By Proposition \ref{Miranda convergence} and Lemma \ref{probabilistic method}, for almost every $r > 0$,
$$\lim_{t \to 0} \int_{B(x_t, r)} |d1_{U_t}| ~\vol = \int_{B(0, r)} |d1_C| ~\vol$$
and hence, since $\Lambda(C, r; g') = 0$,
$$\lim_{t \to 0} \left|\int_{B(x_t, r)} d1_{U_t} ~\vol\right| = \left|\int_{B(0, r)} d1_C~\vol~\right|.$$
In particular,
\begin{equation}\label{excesses die}
\lim_{t \to 0} \Lambda(U_t, B(x_t, r); g') = 0.
\end{equation}
TODO: Check these estimates carefully against Miranda. Also $C$ is NOT A HYPLERPLANE!!

The constant $\sigma$ in Proposition \ref{induction 2}, selected for the metric $g$ and $\alpha = 0.25$, can be chosen uniformly in a neighborhood of $p$, since it is continuous in $p$.
By (\ref{excesses die}), once $\sigma$ has been chosen uniformly, then we can also select $\delta > 0$ so that if $t < \delta$ then
$$\Lambda(U_t, B(x_t, r); g) < \sigma r^{d - 1}.$$
Therefore by Lemma \ref{tangent cones 1}, $\partial U \cap B(p, 0.25 - 0.25^{d/(d-1)}))$ is as smooth as possible.
Since $p$ was arbitrary, $N = \partial U$ is as smooth as possible.
By Proposition \ref{area estimates}, $N$ is the locally finite union of smooth hypersurfaces, which completes the proof.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix \section{The coarea formula} \label{coarea section}
We start by recalling the Bochner integral.
Let $F \to M$ be a vector bundle and let $(\Omega, P)$ be a measure space.
The space $\mathcal R(U, F)$ of $F$-valued Radon measures on $U$ equipped with the weak topology of measures can be viewed as a separable Fr\'echet space with seminorms $\langle \cdot, f_j\rangle$ where $(f_j)$ is a countable basis for a dense subspace of $C_c(U, F)$.

\begin{theorem}[Pettis]
Let
$$f: \Omega \to \mathcal R(U, F).$$
Then $f$ is measurable iff for every $X \in \mathcal D(M, F')$, $\omega \mapsto \langle f(\omega), X\rangle$ is measurable.
In this case,
$$\left\langle \int_\Omega f ~dP, X\right\rangle = \int_\Omega \langle f(\omega), X\rangle ~dP(\omega).$$
\end{theorem}
\begin{proof}
See \cite[Chapter V]{yosida2012functional}.
\end{proof}

Throughout this section we consider the superlevel sets $E_y = \{u \geq y\}$ of a function $u \in BV_l(M)$.
We first observe that for every $X \in \mathcal D(M, TM)$,
$$\langle d1_{E_y}, X\rangle = -\int_{E_y} \mathcal L_XdV$$
is measurable in $y$, since $E_y$ is monotone in $y$.
So by Pettis' theorem, $d1_{E_y}$ is measurable in $y$ with respect to the weak topology of measures.

\begin{lemma}[coarea formula for measures]\label{Coarea1}
If $u \in BV_l(M)$ and $E_y = \{u \geq y\}$, then
$$du = \int_{-\infty}^\infty d1_{E_y} ~dy.$$
\end{lemma}
\begin{proof}
Fix a test field $X \in \mathcal D(M, TM)$.
Breaking up $u$ into positive and negative parts, and exploiting the identity
$$\int_{-\infty}^\infty d1_{E_y} ~dy = \int_{-\infty}^0 d1_{E_y} ~dy + \int_0^\infty d1_{E_y} ~dy,$$
it is no loss to assume that $u \geq 0$.

We must show
\begin{equation}
\label{gradient is integral of fibers}
\langle du, X\rangle = \int_{-\infty}^\infty \langle d1_{E_y}, X\rangle ~dy.
\end{equation}
Since $u \geq 0$,
\begin{align*}
\langle du, X\rangle &= -\int_M u~\mathcal L_XdV = -\int_M \int_0^{u(x)} dy ~\mathcal L_XdV\\
&= -\int_0^\infty \int_{E_y} ~\mathcal L_XdV ~dy = \int_0^\infty \langle d1_{E_y}, X\rangle ~dy.
\end{align*}
by Fubini's theorem.
If $y < 0$ then $1_{E_y} = 1$ so $d1_{E_y} = 0$, so we conclude (\ref{gradient is integral of fibers}).
\end{proof}

Throughout the rest of this section we will be interested in the trivial line bundle
$$p: L \to M.$$
Let $\eta = dV \wedge dy$ be the natural volume form on $L$ and let $h$ be the natural metric.

If $W$ is a vector field on $L$, we will write $W_1$ for the projection of $W$ onto $M$ and $W_2$ for its projection onto $\RR$.
Then Cartan's magic formula implies that if $W_2$ is constant, then
\begin{equation}
\label{Lie derivative computation}
\mathcal L_W\eta = \mathcal L_WdV \wedge dy.
\end{equation}

\begin{lemma}\label{coarea converse}
Let $u \in BV_l(M)$ and $E_y = \{u \geq y\}$.
Suppose that $W$ is a test field on $L$ which depends on a parameter $n \in \NN$, such that $W_2 = 0$ and for every $y \in \RR$, $X = W_1(\cdot, y)$ is a maximizing sequence for $\langle d 1_{E_y}, X\rangle$ subject to $X \prec U$.
Then
$$\int_{-\infty}^\infty \langle d 1_{E_y}, W(y)\rangle ~dy \leq |d u|_{TV}(U).$$
\end{lemma}
\begin{proof}
Let
$$E = \{(x, y) \in L: x \in E_y\}$$
be the undergraph of $u$.
By Fubini's theorem and (\ref{Lie derivative computation}),
\begin{align*}
\int_{-\infty}^\infty \langle d 1_{E_y}, W(y)\rangle ~dy &= -\int_{-\infty}^\infty \int_{E_y} \mathcal L_WdV ~dy = -\iint_E ~\mathcal L_W\eta = \langle d 1_E, W\rangle.
\end{align*}

Let $(u_m)$ be a mollification of $u$, so that $u_m \to u$ in the weak topology of distributions.
Then if $\chi$ is a cutoff, $\langle u_m, \chi\rangle \to \langle u, \chi\rangle$; taking a sequence of $\chi$ which increase to the indicator function of a compact set $K$, we conclude that $u_m \to u$ in $L^1(K)$, and hence $u_m \to u$ in $L^1_l$.

Let $E^{(m)}$ be the undergraph of $u_m$, and $E^{(m)}_y = \{u_m \geq y\}$.
Then for every test function $v$,
\begin{align*}
\langle 1_{E^{(m)}} - 1_E, v\rangle &= \int_{E^{(m)} \Delta E} v ~dV \leq |(E^{(m)} \Delta E) \cap (\supp v \times \RR)| \cdot ||v||_{L^\infty}\\
&\leq ||v||_{L^\infty} \int_{\supp v} |u_m - u| ~dV \to 0
\end{align*}
so $1_{E^{(m)}} \to 1_E$ in the weak topology of distributions.
Therefore
$$\lim_{m \to \infty} \langle d 1_E^{(m)}, W\rangle = \langle d 1_E, W\rangle.$$

Since $u_m$ is smooth, its graph $F_m = \partial E^{(m)}$ is a smooth manifold.
Let $\nu_m$ be the upwards unit normal field of $F_m$ and let $dV_m$ be the volume form on $F_m$ induced by $\eta$.
Then
$$\langle d 1_{E^{(m)}}, W\rangle = -\iint_{E^{(m)}} \mathcal L_W\eta = -\int_{F_m} h(\nu_n, W) ~dV_m.$$
Let $q_m = p|F_m$ and $Y_m$ be the vector field $(Y_m)_1 = -d u_m$, $(Y_m)_2 = 1$.
Since $F_m$ is a graph, $q_m: F_m \to M$ is a diffeomorphism, $(q_m)_*\nu_m = Y_m/|Y_m|$, and $(q_m)_*dV_m = |Y_m|~dV$. Therefore
$$\int_{F_m} h(\nu_n, W) ~dV_m = \int_M h(Y_m, W) ~dV = \int_M g(d u_m, W_1) ~dV = \langle d u_m, W_1\rangle.$$
Thus
$$|\langle d 1_E, W\rangle| = \lim_{m \to \infty} |\langle d u_m, W_1\rangle| \leq |d u|_{TV}(U),$$
which completes the proof.
\end{proof}

\begin{proposition}[coarea formula for $BV_l$ functions]\label{Coarea2}
If $u \in BV_l(M)$ and $E_y = \{u \geq y\}$ then
$$|du|_{TV} = \int_{-\infty}^\infty |d1_{E_y}|_{TV} ~dy.$$
\end{proposition}
\begin{proof}
By Lemma \ref{Coarea1} and the triangle inequality, one has
$$|du|_{TV} \leq \int_{-\infty}^\infty |d1_{E_y}|_{TV} ~dy.$$
So we just need to prove the converse.

Let $U \Subset M$.
Suppose that for every $y \in \RR$, $X = X^{(n)}_y$ is a maximizing sequence for $\langle d1_{E_y}, X\rangle$ subject to $X \prec U$.
Since $d 1_{E_y}$ is measurable in $y$ with respect to the weak topology of measures, for every $n$, $X^{(n)}_y(x)$ can be chosen to be measurable in $(x, y)$; indeed, we can take $X^{(n)}_y$ to be a smooth approximation to the Radon measure $d 1_{E_y}/|d 1_{E_y}|_{TV}$ in the weak topology of distributions, which is a product of the measurable functions $y \mapsto d 1_{E_y}$ and $y \mapsto |d 1_{E_y}|_{TV}^{-1}$.

By an approximation argument, we can find test fields $W^{(n)}$ on $L$ such that $W^{(n)}_2 = 0$ and for every $y \in \RR$, $X = W^{(n)}_1(\cdot, y)$ is a maximizing sequence for $\langle d 1_{E_y}, X\rangle$ subject to $X \prec U$.
Let us now suppress the $n$ and write $W(y) = W^{(n)}(\cdot, y)$.

By Lemma \ref{coarea converse}, since $W$ has compact support, the integrand $\langle d 1_{E_y}, W(y)\rangle$ is uniformly bounded in $y$.
Therefore, by Fatou's lemma,
\begin{align*}
\int_{-\infty}^\infty |d 1_{E_y}| ~dy &= \int_{-\infty}^\infty \lim_{n \to \infty} \langle d 1_{E_y}, W(y)\rangle ~dy \leq \liminf_{n \to \infty} \int_{-\infty}^\infty \langle d 1_{E_y}, W(y)\rangle ~dy \\
&\leq |d u|_{TV}(U),
\end{align*}
which completes the proof.
\end{proof}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Application to minimal currents} \label{chains of geodesics}
In this section we prove Theorem \ref{main crly} and Corollaries \ref{main crly 2} and \ref{main crly 3}.
Thus we always asssume that $M$ is a surface.

\begin{definition}
A \dfn{geodesic chain} is an $1$-chain $\sum_i \theta_i \gamma_i$, with $\theta_i \in \ZZ$ and $\gamma_i$ geodesics.
\end{definition}

We now rapidly review the notion of integral currents; see \cite[Chapter 4]{federer2014geometric} for a more detailed exposition.

\begin{definition}
Let $\mathscr H^k$ be the $k$-dimensional Hausdorff measure.
A \dfn{$\ZZ$-rectifiable $k$-current} $T$ is defined by the data of a $\mathscr H^k$-rectifiable Borel set $E$, a section $\tau$ of the approximate sphere bundle $SE \to E$, and a Borel function $\theta: E \to \ZZ$, so that $T$ acts on test $k$-forms $\omega$ by
$$\langle T, \omega\rangle = \int_E (\theta\tau, \omega) ~d\mathscr H^k.$$
The \dfn{mass} of a $\ZZ$-rectifiable $k$-current $T$ satisfies
$$||T|| = ||\theta||_{L^1(E, \mathscr H^k)},$$
and the \dfn{boundary} $\partial T$ is the $k-1$-current defined by $\langle \partial T, \omega\rangle = \langle T, d\omega\rangle$.
A $\ZZ$-rectifiable current $T$ is \dfn{integral} if $\partial T$ is $\ZZ$-rectifiable and $||T|| + ||\partial T|| < \infty$.
An integral current $T$ is \dfn{minimal} if it has minimal mass among those integral currents $S$ with $\partial T = \partial S$.
\end{definition}

Since every geodesic chain $T = \sum_i \theta_i \gamma_i$ can be viewed as an integral $1$-current according to the formula
$$\langle T, \omega\rangle = \sum_i \theta_i \int_{\gamma_i} \omega,$$
we will call a geodesic chain \dfn{minimal} if it is minimal in the sense of currents.

\begin{example}\label{least gradient gives minimal chains}
Suppose that $u$ is a function of least gradient and $y \in \RR$.
Then Theorem \ref{main thm} says that for every $y \in \RR$, the set $\partial \{u > y\}$ is (the support of) a minimal geodesic chain.
\end{example}

\begin{lemma}\label{minimal geodesic chains are separated}
Suppose that $T = \sum_i \theta_i \gamma_i$ is a minimal geodesic chain, where $\theta_i \in \NN$ and the $\gamma_i$ are not geodesically collinear.
If $M$ is complete, then there exists $\varepsilon > 0$ such that the distance between any the images of any two $\gamma_i$ is $\geq \varepsilon$.
\end{lemma}
\begin{proof}
Suppose this is false.
We can reduce to the case $T = \gamma_0 + \gamma_1$ by reordering the geodesics and suppressing multiplicity.
Since $T$ has compact support in $M$, we can extend $\gamma_0, \gamma_1$ from $(0, 1)$ to $[0, 1]$, so that by our contradiction hypothesis and a compactness argument, there exist $t_0, t_1 \in [0, 1]$ and $p \in M$ such that $\gamma_i(t_i) = p$.
Furthermore, since the $\gamma_i$ are geodesics, we can find geodesic normal coordinates $(U, x)$ centered on $p$ so that $U \Subset M$ is convex and the image $K_i$ of $\gamma_i$ in $\overline U$ is $\{x_i = 0\}$ or $\{x_i = 0\} \cap \{x_{1-i} \leq 0\}$.

First suppose that exactly one of the $i$ has $K_i = \{x_i = 0\}$, say $i = 0$.
By the Hopf-Rinow theorem, for any $a,b \in M$, there is a length-minimizing geodesic connecting them, which we denote by $[a, b]$.
We write $a_i$ for the start of $\gamma_i$ and $b_i$ for its end.
Define $\rho_0 = [a_0, b_1]$, $\rho_1 = [a_1, b_0]$, and $S = \rho_0 + \rho_1$; then $\partial S = \partial T$, but the rectifiable curve $\gamma_1 + [b_1, b_0]$ has been replaced by $\rho_1$.
Since $\gamma_1$ and $[b_1, b_0]$ are transverse at $p$, they are not geodesically collinear and therefore the triangle inequality implies that
$$||\rho_1|| < ||\gamma_1 + [b_1, b_0]||.$$
But $T = \rho_0 + \gamma_1 + [b_1, b_0]$, so $T$ is not minimal, a contradiction.

Otherwise, either both $i$ satisfy $K_i = \{x_i = 0\}$ or $K_i = \{x_i = 0\} \cap \{x_{1-i} \leq 0\}$.
In the former case we define $V = \bigcup_{\sigma \in \{+, -\}} \bigcap_{i \in \{0, 1\}}\{\sigma x_i > 0\}$, and in the latter we define $V = \bigcap_{i \in \{0, 1\}} \{x_i < 0\}$.
Then $V$ is an open subset of $U$ and $\partial^* V \cap U = K_0 \cup K_1$.
If $V$ does not have least perimeter, then we can replace $K_0 \cup K_1$ with the reduced boundary in $U$ of a set with less perimeter than $V$ and thus reduce the mass of $T$, a contradiction to the minimality of $T$.
So $V$ has least perimeter, and by Theorem \ref{main thm}, $\partial^* V$ is as smooth as possible, which is a contradiction due to the singularity at $p$.
\end{proof}

\begin{proof}[Proof of Corollary \ref{main crly 2}]
Let $T$ be a minimal $1$-current.
By \cite[\S4.2.25]{federer2014geometric}, there exist minimal $1$-currents $T_i$, $i \in I$,
and functions $f_i: [0, ||T_i||] \to M$ such that $T = \sum_i T_i$, $(f_i)_*([0, ||T_i||]) = T_i$, and the Lipschitz seminorm of $f_i$ is $1$.
Since $T_i$ is rectifiable, $T_i$ represents a rectifiable curve $\gamma_i$.
Moreover, $||\partial T|| < \infty$, yet $||\partial T||$ is equal to the number of points of $\partial T$ counted with multiplicity, and every $T_i$ either has two points on its boundary or zero.
However, in the latter case, $\gamma_i$ is a Jordan curve and we can replace $T$ with $T - T_i$ to decrease the mass of $T$ without changing $\partial T$.
In that case, $T$ is not minimal, a contradiction; so $||\partial T_i|| = 2$.
Therefore, unless we have $\partial(T_i + T_j) = 0$ (in which case $T_i + T_j$ is a Jordan curve and we can remove both curves without changing $\partial T$), $||\partial T||$ is at least $\card I$, which therefore must be finite.

We now claim that $\partial U_i$ is the locally finite union of geodesics.
After replacing $M$ with a tubular neighborhood of $\supp T_i$ and applying the Jordan curve theorem, we may assume that there is an open Caccioppoli set $U_i$ such that $\supp T_i = \partial U_i$.
Since $T_i$ minimizes mass, and $||T_i|| = |\partial^* U_i|$, $U_i$ has least perimeter.
Therefore $\partial U_i$ is the locally finite union of geodesics by Theorem \ref{main thm}.

Returning to the case of $M$ a complete surface, the claim now follows from Lemma \ref{minimal geodesic chains are separated}.
\end{proof}

We will be interested in geodesic chains that arise from Example \ref{least gradient gives minimal chains}, so we now study geodesic chains of this form.

\begin{lemma}\label{geodesic foliation is closed}
For every function $u$ of least gradient, the set $\bigcup_{y \in \RR} \partial \{u > y\}$ is the closure of $\{du \neq 0\}$.
\end{lemma}
\begin{proof}
Let $A = \bigcup_y \partial \{u > y\}$, $B$ the interior of $\{du = 0\}$, and $x \in M$.
Then $x \in B$ iff $u = u(x)$ near $x$, but that happens iff for every $y < u(x)$, $x$ is interior to $\{u > y\}$ and for every $y \geq u(x)$, $x$ is exterior to $\{u > y\}$.
This happens iff for every $y \in \RR$, $x$ is either interior or exterior to $\{u > y\}$, thus $x \notin A_y$, which happens iff $x \notin A$.
Thus $\{A, B\}$ is a partition of $M$.
\end{proof}

\begin{notation}
Suppose that $u$ has least perimeter and $x \in M$.
We write $y(x)$ to denote some $y$ such that $x \in \partial \{u > y\}$, if such a $y$ exists.
\end{notation}
We write $\normal_y$ for the conormal $1$-form to $A_y = \partial \{u > y\}$, and set $X_x = {\normal_{y(x)}}^\sharp$, where $y(x)$ is chosen so that $x \in A_{y(x)}$.

\begin{lemma}\label{existence of normal field}
Suppose that $u$ has least perimeter and $A = \bigcup_y \partial \{u > y\}$.
Then there is a nonvanishing smooth vector field $X$ defined on an open neighborhood of $A$, such that $X$ restricts to a section of the sphere bundle of $A$, and for every $y \in \RR$, $X$ is normal to $y$.
\end{lemma}
\begin{proof}
Let $\normal_y$ be the conormal $1$-form to $A_y = \partial \{u > y\}$, and set $X_x = {\normal_{y(x)}}^\sharp$.
Since $A_y$ is smooth, $A_y$ is a reduced boundary and so $\normal_y$ has unit length; therefore $X$ is a section of the sphere bundle $SM \to M$ defined on $A = \bigcup_y A_y$.
On any curve $A_y$, $X$ is smooth, and by definition of the reduced boundary, we have $X|\grad u| = \grad u$.

TODO: Deal with the critical points of $u$ in $A_y$. We don't need $X$ smooth, just Lipschitz.

We can then extend $X$ to a nonvanishing vector field on an open neighborhood of $A$.
\end{proof}

\begin{proof}[Proof of Corollary \ref{main crly 3}]
Suppose that $u$ is a set of least perimeter such that the set of critical points of $u$ is nowhere-dense.
By Lemma \ref{geodesic foliation is closed}, the set $A = \bigcup_y \partial \{u > y\}$ is dense in $M$ and closed, and therefore $A = M$.
On the other hand, by Lemma \ref{existence of normal field}, there is a nonvanishing smooth vector field on $A$; therefore the tangent bundle of $M$ is trivial.
\end{proof}

\begin{proof}[Proof of Theorem \ref{main crly}]
Let $A,X$ be as in Lemma \ref{existence of normal field}.
We write $\gamma_x$ for the integral curve of $X$ with $\gamma_x(0) = x$, so $\gamma_x$ is normal to every curve $A_y$ which it crosses through.
Now consider a vector $v(0) \in T_x A_{y(x)}$, which is an infinitesimal generator of a geodesic $F_x(0, \cdot)$ which parametrizes $A_{y(x)}$ near $x$ and satisfies $F_x(0, 0) = x$.
Parallel transporting $v(0)$ along $\gamma_x$ using the Levi-Civita connection, we obtain tangent vectors $v(t) \in T_{\gamma_x(t)} A_{y(\gamma_x(t))}$ which are infinitesimal generators of geodesics $F_x(t, \cdot)$ that parametrize $A_{y(\gamma_x(t))}$ near $\gamma_x(t)$ and satisfies $F_x(t, 0) = \gamma_x(t)$.
Then $F_x$ defines a foliated chart $F_x: I \times J \to A$ centered on $x$ such that for every $t \in I$ such that $F_x(t, 0)$, $F_x(t, \cdot)$ is a geodesic.
Moreover, if $x_1, x_2$ have overlapping charts $F_{x_i}$, then it follows from the definitions that $F_{x_1},F_{x_2}$ are coherently foliated.
By Lemma \ref{geodesic foliation is closed}, $A$ is closed, so $(A, (F_x)_{x \in A})$ defines a geodesic lamination $\mathscr F$.

Now we extend off $U$ into $\overline U$.
According to Lemma \ref{minimal geodesic chains are separated}, if $U$ is the domain of $u$ then we can extend $\mathscr F$ to $\overline U$ provided that $M$ is complete.
\end{proof}



\printbibliography


\end{document}
